---
title: "ISCN visualizations and summary"
author: "K Todd-Brown (ktoddbrown@gmail.com)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
library(SoilDataR)
library(ggplot2)
library(reshape2)
library(ggmap)
library(maps)
library(mapdata)
library(plyr)
library(fiftystater)
```

```{r loadData}
ISCN3 <- processData_ISCN3(dir='~/Documents/CurrentProjects/soilDataRecovery/data/ISCN3')
```


#Map of sites
```{r mapSites}

mapWorld <- borders("world", colour="gray80", fill="gray80") # create a layer of borders
#ggplot() + mapWorld
ggplot(unique(ISCN3$field[, c('lat', 'long')])) +
  mapWorld + 
  geom_hex(aes(x=long, y=lat), bins=200) + 
  scale_fill_gradient(trans='log10')
```
```{r mapSiteByArea}
noLatLon <- unique(subset(ISCN3$field, is.na(lat+long))$fieldID)

stateCounts <- ddply(subset(ISCN3$fieldNotes, measurement == 'state' & fieldID %in% noLatLon), c('value'), summarize, count=length(unique(fieldID)))
names(stateCounts)[1] <- 'state'
stateCounts$state <- tolower(stateCounts$state)
stateCounts <- subset(stateCounts, state %in% fifty_states$id)


ggplot(stateCounts, aes(map_id = state)) +
  geom_map(aes(fill=count), map=fifty_states) +
  expand_limits(x = fifty_states$long, y = fifty_states$lat) +
  coord_map() +
  scale_x_continuous(breaks = NULL) + 
  scale_y_continuous(breaks = NULL) +
  scale_fill_gradient(trans='log10') +
  labs(x = "", y = "") +
  theme(legend.position = "bottom", 
        panel.background = element_blank())

countryCounts <- ddply(subset(ISCN3$fieldNotes, measurement == 'country' & fieldID %in% noLatLon), c('value'), summarize, count=length(unique(fieldID)))
names(countryCounts)[1] <- 'region'
countryCounts$region <- as.character(countryCounts$region)
countryCounts$region[grepl('United States', countryCounts$region)] <- 'USA'

map.world <- map_data(map="world")

missingCountries <- data.frame(region=setdiff(map.world$region, countryCounts$value), count=NA)
countryCounts <- rbind(countryCounts, missingCountries)
#map.world <- merge(map.world, countryCounts, all=TRUE)
ggplot(countryCounts, aes(map_id = region)) +
  geom_map(aes(fill=count), map=map.world) +
  expand_limits(x = map.world$long, y = map.world$lat) +
  coord_cartesian(xlim=c(-180,180), ylim=c(-90,90)) +
  scale_x_continuous(breaks = NULL) + 
  scale_y_continuous(breaks = NULL) +
  scale_fill_gradient(trans='log10') +
  labs(x = "", y = "") +
  theme(legend.position = "bottom", 
        panel.background = element_blank())



```

`r nrow(subset(subset(ISCN3$fieldNotes, measurement == 'country' & fieldID %in% noLatLon),
            value=='Unknown'))` ites with no known country.

##Total country counts
```{r totCountryCount}
countryCounts <- ddply(subset(ISCN3$fieldNotes, measurement == 'country'), c('value'), summarize, count=length(unique(fieldID)))
names(countryCounts)[1] <- 'region'
countryCounts$region <- as.character(countryCounts$region)
countryCounts$region[grepl('United States', countryCounts$region)] <- 'USA'


map.world <- map_data(map="world")

missingCountries <- data.frame(region=setdiff(map.world$region, countryCounts$region), count=NA)
setdiff(countryCounts$region, map.world$region)
countryCounts <- rbind(countryCounts, missingCountries)
#map.world <- merge(map.world, countryCounts, all=TRUE)
ggplot(countryCounts, aes(map_id = region)) +
  geom_map(aes(fill=count), map=map.world) +
  expand_limits(x = map.world$long, y = map.world$lat) +
  coord_cartesian(xlim=c(-180,180), ylim=c(-90,90)) +
  scale_x_continuous(breaks = NULL) + 
  scale_y_continuous(breaks = NULL) +
  scale_fill_gradient(trans='log10') +
  labs(x = "", y = "") +
  theme(legend.position = "bottom", 
        panel.background = element_blank())
```

#Histograms of measurements
```{r types}
unique(ISCN3$measurement$type)
```
Most of the data is from NRCS (soil characterization database) Sept 2014 version of data base. 

All measurements are on oven dry soils and in mass percent where percent noted.

##Bulk density

+ bd_samp is the bulk density of the <2mm (fine earth) fraction, in which the mass is expressed on an oven-dry (105 deg) basis. This is the best form of Db for the most uses.
+ bd_tot is the whole soil bulk density, includes fine earth fraction and rocks.
+ bd_whole is the bulk density of the whole soil (coarse frags, if present, included), expressed on a 1/3 bar moisture content (a little drier than typical field moist, but not oven-dry). It is collected by the clod method. For a variety of reasons, including the presence of water, possible coarse frags, and the tendency of clods to cleave along pore spaces, this will overestimate Db relative to bd_samp. The SSL standard method here is 4A1d if you want to look it up.
+ bd_other is, for data contributed by NRCS SSL, the bulk density of the fine earth fraction, but expressed on a field moist condition. It is collected by a volumetric core, but given the presence of some water, will probably usually be an overestimate relative to the bd_samp. This corresponds to the SSL standard method 4A3a. For datasets contributed by sources other than NRCS, bd_other is anyone's guess. Hopefully specified in the associated bd_method variable, or the metadata sheet for that dataset.

Reasonable BD minimum 0.03 or lower and maximum 2.7 g cm-3. 
```{r BD}
plot.df <- merge(subset(ISCN3$sample, grepl('^bd_', measurementID) & 
                          value > 0 & value < 10), ISCN3$measurement)
ggplot(plot.df) +
  geom_histogram(aes(x=value, y=..density..)) + 
  scale_x_log10() + 
  facet_wrap(type~unit)

summary(plot.df)
```

Number of samples for BD `r length(plot.df$value)`.

##C OC LOI SOC

+c_tot carbon concentration in a dry combustion analysis, sometimes include inorganic carbon
+oc organic carbon concentration from either acidified (to remove carbonates) and then ran on dry combustion analyizer, or (more common) wet oxidation 'Walkly-Black' dicromate oxidation, chemical digest for organic carbon; old methods 1930-1990s tends to over estimate it in certain soils. pH below 7 c_tot generally == oc otherwise you might have carbonates
+loi loss on ignition uncorrected for C:organics, generally close to half for O-horizon, C about a third of the LOI for lower horizons.

+soc calculated soil carbon stock of layer/profile

```{r C_OC_LOI_SOC}
ggplot(merge(subset(ISCN3$sample, grepl('^(c_|oc|loi|soc)', measurementID) & value > 0),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  scale_x_log10() + 
  facet_wrap(type~unit)

summary(subset(ISCN3$sample, grepl('^(c_|oc|loi)', measurementID) & (value < 0)))
```

## Nitrogen
+ n_tot most N is organic mass percent dry weight, probably estimated by dry combustion (Dumas methods), sometimes also Kjeldahl digestion converts N to nitrate and then measures nitrate.

```{r N}
ggplot(merge(subset(ISCN3$sample, grepl('^n_', measurementID) & value > 0),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  scale_x_log10() + 
  facet_wrap(type~unit)
```

## pH

+ ph_cacl calsium cloride slury in the NRCS documentation
+ ph_h2o water slury in the NRCS documentation
+ ph_other non standard catch all

```{r pH}
ggplot(merge(subset(ISCN3$sample, grepl('^ph_', measurementID) & value > 1.5),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  #scale_x_log10() + 
  facet_wrap(type~unit)

summary(subset(ISCN3$sample, grepl('^ph_', measurementID) & (value <= 1.5)))
```

## CaCO3 
Percent inorganic carbon in a sample.
Standard methods from NRCS SSL (SoiL Survey Manual, R Burt) manual
```{r CaCO3}

ggplot(merge(subset(ISCN3$sample, grepl('^caco3', measurementID) & value > 0),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  scale_x_log10() + 
  facet_wrap(type~unit)

summary(subset(ISCN3$sample, grepl('^caco3', measurementID) & (value <= 0)))

```

##Sand, clay slit
Partical size analysis. 
Methods might not be completely annotated.

```{r Texture}

ggplot(merge(subset(ISCN3$sample, grepl('tot_psa', measurementID) & value >= 0),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  #scale_x_log10() + 
  facet_wrap(type~unit)

summary(subset(ISCN3$sample, grepl('tot_psa', measurementID) & (value < 0)))

```

##Cation exchange capasity
We don't specify extrat solution, protocol in SSL. 
Only for NRCS data in vs ISCN3.

```{r cat_exch}

ggplot(merge(subset(ISCN3$sample, grepl('(cat_exch)', measurementID)& (value > 0)),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  scale_x_log10() + 
  facet_wrap(type~unit)

summary(subset(ISCN3$sample, grepl('(cat_exch)', measurementID) & (value <= 0)))

subset(ISCN3$measurement,grepl('(cat_exch)', measurementID))
```

## Percent coarse fragement content
By weight coarse fragment (>2mm) content. Inherited SSL nominclature
```{r wpg}
ggplot(merge(subset(ISCN3$sample, grepl('wpg2', measurementID) & value > 0),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  scale_x_log10() + 
  facet_wrap(type~unit)

#summary(summary(subset(ISCN3$sample, grepl('wpg2', measurementID) & (value <= 0))))
```

## Metal extracts
bc_* - base cation exchange
We don't specify extrat solution, protocol in SSL. National Corroporative Soil Survey data base details in Soil Survey Lab methods (SSL).
_dith vs _ox - differnt organic extractent. Might not be comparable across extractents. Selection is reflectent of different mineralogies.
Only for NRCS data in vs ISCN3.
```{r metals_percent}
ggplot(merge(subset(ISCN3$sample, 
                    grepl('^(al|fe|mn|ca|k|mg|na|p)_', measurementID) &
                      !grepl('ext', measurementID)),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  scale_x_log10() + 
  facet_wrap(type~unit, scales='free')

summary(subset(ISCN3$sample, grepl('^(al|fe|mn|ca|k|mg|na|p)_', measurementID) &
                      !grepl('ext', measurementID) & (value > 100)))
```

```{r metals_ext}
ggplot(merge(subset(ISCN3$sample, grepl('ext', measurementID) & value > 0),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  scale_x_log10() + 
  facet_wrap(type~unit, scales='free')

summary(subset(ISCN3$sample, grepl('ext', measurementID) & (value <= 0)))
```

## Cation exchange
NRCS variables
bs - sum of sum of all base cation exchange
bs_sum - alternative 2 calculation of sum of all base cation exchange

base_sum - 

cec_sum - cation exchange by different method
ecec - effective cation exchange capasity (standarized soil pH)
h_ext - exchangable acidity

```{r base_cec}
#base_sum     cec_sum      ecec         bs           bs_sum 
ggplot(merge(subset(ISCN3$sample, grepl('(base_sum|cec_sum|ecec)_', measurementID) & 
                      value > 0),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  scale_x_log10() + 
  facet_wrap(type~unit, scales='free')

summary(subset(ISCN3$sample, grepl('(base_sum|cec_sum|ecec)_', measurementID) & (value <= 0)))
```

```{r bs}
#bs           bs_sum 
ggplot(merge(subset(ISCN3$sample, grepl('^bs_', measurementID)),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  #scale_x_log10() + 
  facet_wrap(type~unit, scales='free')

#summary(subset(ISCN3$sample, grepl('^bs_', measurementID)))
```



```{r isotope}
#15n          13c          14c
ggplot(merge(subset(ISCN3$sample, grepl('^1[345]', measurementID)),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  #scale_x_log10() + 
  facet_wrap(~type, scales='free')

summary(subset(ISCN3$sample, grepl('^1[345]', measurementID)))
```

#Locator_parent_alias 
ties back to old site name in NRCS
