---
title: "ISCN 3.3 data"
author: "K Todd-Brown (ktoddbrown@ufl.edu)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Purpose: *This vignette describes the ISCN3 data contributions.*

```{r setup, echo=FALSE}
library(SOCDRaH2)
library(ggplot2)
library(knitr)
library(kableExtra)
dataDir <- '~/Documents/Datasets/ISCN'

knitr::opts_chunk$set(eval=FALSE)
```

```{r fakeSetUp, eval=FALSE}
library(SOCDRaH2)
library(ggplot2)
dataDir <- 'Your/Data/Dir'
```

# Accessing the data

```{r load, eval=FALSE}
ISCN3.ls <- ISCN3_3(data_dir = dataDir)
```

# Creating a table to show summary statistics of the number of unique sites each dataset has

```{r sites}
knitr::kable(profile_raw %>%
  select(dataset_name_sub, site_name) %>%
  unique()%>%
  group_by(dataset_name_sub) %>% 
  tally(name = 'site_count'), 
  caption = 'Site counts for each contributing dataset submision (dataname_sub) in the profile data table.')%>%
  kable_styling()
```

# Dataset description table

```{r}
study <- dataset_raw %>% 
  select(dataset_name, `dataset_type (dataset_type)`, dataset_description) 
knitr::kable(study,
  caption = 'Site counts for each contributing dataset submision (dataname_sub) in the profile data table.')%>%
  kable_styling()
```

# Description of the tables and columns

Create a table with the table name, column name, and long description similar to what is on the EDI page.

# New section for each data contribution

Include both a description of the data sets including counts of layers and sites, histograms and tables of variables, and site map.
List data contributions from largest to smallest.

# Citations and bibliography

Create a table with each dataset and citation, repeat the dataset name as needed for each citation.

```{r}
#need to gap fill these or find the gap filled ones
study <- citation_raw %>% 
  select(dataset_name, citation) 
knitr::kable(study,
  caption = 'Citations for each contributing dataset submision (dataname_sub).')%>%
  kable_styling()
```

