---
title: "ISCN 3.3 data"
author: "K Todd-Brown (ktoddbrown@ufl.edu)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Purpose: *This vignette describes the ISCN3 data contributions.*

```{r setup, echo=FALSE}
library(SOCDRaH2)
library(ggplot2)
dataDir <- '~/Documents/Datasets/ISCN'
```

```{r fakeSetUp, eval=FALSE}
library(SOCDRaH2)
library(ggplot2)
dataDir <- 'Your/Data/Dir'
```

# Accessing the data

```{r load}
ISCN3.ls <- ISCN3(dataDir = dataDir, orginalFormat = TRUE)
```

# Description of the tables and columns

# New section for each data contribution

Include both a description of the data sets including counts of layers and sites, histograms and tables of variables, and site map.
List data contributions from largest to smallest.

# Citations and bibliography

Create a table with each dataset and citation, repeat the dataset name as needed for each citation.
