---
title: "ISCN 3.3 database"
author: "K Todd-Brown (ktoddbrown@ufl.edu)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: ISCN3_bibliography.bib 
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Purpose: *This vignette describes the ISCN3 data contributions.*

```{r setup, echo=FALSE}
library(tidyverse)
library(SOCDRaH2)
library(ggplot2)
dataDir <- '~/Desktop/datasets/ISCN'

knitr::opts_chunk$set(eval=FALSE)
```

The ISCN database is a compilation of research and survey layer-level soil data and observations.
The most recent archive of ISCN3 is here [https://portal-s.edirepository.org/nis/mapbrowse?scope=edi&identifier=360&revision=4](https://portal-s.edirepository.org/nis/mapbrowse?scope=edi&identifier=360&revision=4)

# Setup

```{r fakeSetUp, eval=FALSE}
library(SOCDRaH2)
library(ggplot2)
dataDir <- 'Desktop/datasets/ISCN'
```

# Accessing the data


# Description of the tables and columns

Create a table with the table name, column name, and long description similar to what is on the EDI page.

# New section for each data contribution

Include both a description of the data sets including counts of layers and sites, histograms and tables of variables, and site map.
List data contributions from largest to smallest.

ISCN is large and will take a minute to download and load.
ISCNv3.3 does not add any new data to ISCN3 however it correct several known data short comings.

```{r load, message=FALSE, warning=FALSE}
ISCN3.ls <- ISCN3_3(data_dir = dataDir) #this will throw an NA coercion warning that you can ignore
```

ISCN profile (2-D) and layer (3-D) level data comes from the following contributing data sets.

```{r sites}
profile_summary <- ISCN3.ls$profile %>%
  select(dataset_name_sub, site_name) %>%
  unique()%>%
  group_by(dataset_name_sub) %>% 
  tally(name = 'profile_site_count')

layer_summary <- ISCN3.ls$layer %>%
  select(dataset_name_sub, site_name) %>%
  unique()%>%
  group_by(dataset_name_sub) %>% 
  tally(name = 'layer_site_count')

table_summary <- ISCN3.ls$study %>% 
  select(dataset_name, `dataset_type (dataset_type)`, dataset_description) %>%
  unique() %>%
  right_join(profile_summary %>%
              full_join(layer_summary, by = 'dataset_name_sub'), by=c('dataset_name' = 'dataset_name_sub'))

knitr::kable(table_summary %>%
    arrange(layer_site_count), 
  caption = 'Site counts for each contributing dataset submision (dataname_sub) in the profile and layer data table.')%>%
  kable_styling()
```


# Data Contributions

## University of Michigan Biological Station Forest Accelerated Succession ExperimenT (`UMBS_FASET`)

```{r}
#temp.outputs <- plotISCN3(ISCN3 = ISCN3.ls, 'UMBS_FASET')
```

The University of Michigan Biological Station Forest Accelerated Succession Experiment (`UMBS_FASET`) data set is uncited and unpublished.

```{r}
study <- ISCN3.ls$study %>% 
  filter(dataset_name == "UMBS_FASET") %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == "UMBS_FASET")%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == "UMBS_FASET") %>%
  select(where(~any(!is.na(.))))

locations <- profile %>% 
  select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
  bind_rows(layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)) %>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
knitr::kable(study,
  caption = 'Citations for the UMBS_FASET data set.') %>%
  kable_styling()
```

This data set is located in the United States and has `r nrow(layer)` layer level observations and `r nrow(profile)` profile-level observations of `r nrow(locations)` unique latitude-longitude location.

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

#remove data from the layer that appears in the profile that is not the top/bottom/soc
#select all numerical data, pivot and histogram
#select all factor data and tally
```


### Fixed issues

Below is a list of know issues that were addressed between ISCN 3 and ISCN 3.3

  + The UMBS FASET data set contains `r nrow(filter(ISCN3.ls$layer, dataset_name_sub == "UMBS_FASET"))` layer-level rows and `r nrow(filter(ISCN3.ls$profile, dataset_name_sub == "UMBS_FASET"))` profile-level information rows after cleaning.
  + `ISCN SOC computation` soil organic carbon calculations were removed and duplicate rows deleted.
  + Values in the `hzn` column of the `layer` table were reassigned to an NA character from "?", a nonspecific character.
  + Values in  `country (country)` column were reassigned to "United States" in the `layer` and `profile` tables because it was the known location of the collection state.
  
### Future issues
  + Citation practices were initially incompatible with CC-BY license for ISCN3; contacting data providers about modifying release for ISCN4
  + Check and see if data is still being collected at this research site
  
### Figrues

The soil data in this data set were sampled from Michigan (see figure below)

```{r}
print(temp.outputs$layerMap) 
```

The profile level map shows

```{r}
print(temp.outputs$profileMap) 
```

A summary of the data set and its contacts are shown below.

```{r}
temp.outputs$summaryStudyTable%>%
  kable_styling()
```

The layer data shows 

```{r}
temp.outputs$layerTable%>%
  kable_styling()
```

The profile data shows 

```{r}
temp.outputs$profileTable%>%
  kable_styling()
```

The histograms for the data set show 

```{r}
print(temp.outputs$profileValueHistograms) 
```

The layer depth plots show 

```{r}
print(temp.outputs$depthValuePlots) 
```

### Changes from ISCN3 to ISCN3.3

    - ISCN SOC computations were removed to elminate gapfilled values.
    - Layer level horizon information (`hzn`) were reassigned to `NA` from `"?"`.
    - Layer and profile level country identifiers (`country (country)`) were reassigned to `"United States"` from an empty string. 
    - Citations were also updated.

A citation was pulled regarding the unpublished dataset and placed in the bibliography but should be revisted in ISCN 4.

# Heckman lithosequence

```{r warning=FALSE, message=FALSE}
datasetName <- 'Heckman lithosequence'

study <- ISCN3.ls$study %>% 
  filter(dataset_name == datasetName) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

#comparison for pre ISCN soc stock correction
# dataset_profile_temp <- profile_raw  %>%
#   filter(dataset_name_sub == datasetName) %>%
#   mutate(`country (country)` = 'United States') %>%
#   standardCast()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == datasetName)%>%
  select(where(~any(!is.na(.))))

#comparison before SOC correction
# dataset_layer_temp <- layer_raw %>%
#   filter(dataset_name_sub == datasetName) %>%
#   standardCast() %>%
#   mutate(`country (country)` = 'United States')

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == datasetName) %>%
  select(where(~any(!is.na(.))))

locations <- profile %>% 
  select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
  bind_rows(layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)) %>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))
```

The `Heckman lithosequence` data set in ISCN3 contains `r nrow(dataset_layer)` layer-level information and `r nrow(dataset_profile)` profile-level information after cleaning for ISCN3.5.
In ISCN3 the country was not set to United states originally dispite the state being identified as Arizona, ISCN3.5 now identifies this as `'United States'`.
ISCN3 also filled in the soil carbon stocks where bulk density and organic carbon was provided, this was removed for ISCN3.5.

```{r message=FALSE, warning=FALSE}
knitr::kable(t(study))
```


There are the following factors in the profile:

```{r}
knitr::kable(summary(profile %>% select_if(is.factor)))
```

And the following factors in the layers:
```{r}
knitr::kable(summary(layer %>% select_if(is.factor)))
```


## Location

```{r}
ggplot(data =  map_data('state') %>%
         full_join(profile %>% 
                     mutate(region = 
                              tolower(`state (state_province)`), by = 'region') %>%
                     group_by(region) %>% 
                     tally, by='region')) + 
  geom_polygon(aes(x=long, y = lat, group=group, fill = n), 
               color = 'black') +
  coord_fixed(1.3) + 
  theme_nothing() 

ggplot(data =  map_data('state') %>% 
         filter(region %in% c("arizona", "utah", "new mexico", "colorado"))) + 
  geom_polygon(aes(x=long, y = lat, group = group), 
               fill = 'grey', color = 'black') + 
  geom_point(data=profile, 
             aes(x = `long (dec. deg)`, y = `lat (dec. deg)`),
             shape = 'x', color = 'red') +
  coord_fixed(1.3) +
  theme_nothing() 

```

```{r eval=FALSE}
#this is useful to see for the analysis but we don't want it in the report
layer %>%
  pivot_longer(cols = intersect(names(.), type_cols$num_cols), values_drop_na = TRUE) %>%
  group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) %>%
  bind_rows(
    layer %>%
      pivot_longer(cols = intersect(names(.), type_cols$factor_cols), values_drop_na = TRUE) %>%
      group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) ) %>%
  arrange(n) %>%
  knitr::kable()
```



## Profile histograms

```{r}
ggplot(profile %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
  geom_histogram(aes(x=measurement)) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```

## Depth plots

```{r}
ggplot(layer %>% 
         pivot_longer(cols=c('layer_top (cm)', 'layer_bot (cm)'),
                       values_to='depth') %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
         geom_line(aes(x=depth, y= measurement, group = profile_name)) +
  facet_wrap(~type, scales='free')
```

## TODO

- Follow up with Heckman to get key for `root_quant_size` and `color`.

## Citations

Please see @Heckman2009 for additional details and if you are using ISCN3 please cite.

## Myers Smith

```{r}
temp.outputs <- plotISCN3(ISCN3 = ISCN3.ls, 'Myers-Smith')
```

The Myers Smith data set is published and cited.
See @MyersSmithHarden2007 @MyersSmithMcGuire2007 for additional details

```{r}
study <- ISCN3.ls$study %>% 
  select(dataset_name, citation) %>%
  filter(dataset_name == "Myers-Smith")
knitr::kable(study,
  caption = 'Citations for the Myers Smith data set.')%>%
  kable_styling()
```

The Myers Smith data set contains `r nrow(filter(ISCN3.ls$layer, dataset_name_sub == "Myers-Smith"))` layer-level rows and `r nrow(filter(ISCN3.ls$profile, dataset_name_sub == "Myers-Smith"))` profile-level information rows after cleaning.

### Figrues

The soil data in this data set were sampled from Alaska (see figure below)

```{r}
print(temp.outputs$layerMap) 
```

The profile level map shows

```{r}
print(temp.outputs$profileMap) 
```

A summary of the data set and its contacts are shown below.

```{r}
temp.outputs$summaryStudyTable%>%
  kable_styling()
```

The layer data shows 

```{r}
temp.outputs$layerTable%>%
  kable_styling()
```

The profile data shows 

```{r}
temp.outputs$profileTable%>%
  kable_styling()
```

The histograms for the data set show 

```{r}
print(temp.outputs$profileValueHistograms) 
```

The layer depth plots show 

```{r}
print(temp.outputs$depthValuePlots) 
```

### Changes from ISCN3 to ISCN3.3

    - ISCN SOC computations were removed to elminate gapfilled values.
    - Citations were also updated.


## Oak Ridge National Lab_Loblolly DWJ

```{r}
temp.outputs <- plotISCN3(ISCN3 = ISCN3.ls, 'Oak Ridge National Lab_Loblolly_DWJ')
```

The Oak Ridge National Lab Loblolly DWJ data set is published and cited.
See @Parr2006 for additional details. 

```{r}
study <- ISCN3.ls$study %>% 
  select(dataset_name, citation) %>%
  filter(dataset_name == "Oak Ridge National Lab_Loblolly_DWJ")
knitr::kable(study,
  caption = 'Citations for the Oak Ridge National Lab Loblolly DWJ data set.')%>%
  kable_styling()
```

The Oak Ridge National Lab Loblolly DWJ data set contains `r nrow(filter(ISCN3.ls$layer, dataset_name_sub == "Oak Ridge National Lab_Loblolly_DWJ"))` layer-level rows and `r nrow(filter(ISCN3.ls$profile, dataset_name_sub == "Oak Ridge National Lab_Loblolly_DWJ"))` profile-level information rows after cleaning. 
The data set does not contain profile level information therefore the profile information graphs are omitted.

The soil data in this data set were sampled from Tennessee (see figure below)

```{r}
print(temp.outputs$layerMap) 
```

A summary of the data set and its contacts are shown below.

```{r}
temp.outputs$summaryStudyTable%>%
  kable_styling()
```

The layer data shows 

```{r}
temp.outputs$layerTable%>%
  kable_styling()
```

The histograms for the data set show 

```{r}
print(temp.outputs$profileValueHistograms) 
```

The layer depth plots show 

```{r}
print(temp.outputs$depthValuePlots) 
```

### Changes from ISCN3 to ISCN3.3

    - ISCN SOC computations were removed to elminate gapfilled values.
    - Layer and profile level country identifiers (`country (country)`) were reassigned to `"United States"` from an empty string. 
    - Layer level horizon information (`hzn`) were reassigned to `NA` from `"?"`.
    - gapfilled bd with sampling were reasigned to "NA" based on method notes.
    - Citations were also updated.
    

## Oak Ridge National Lab TDE

```{r}
temp.outputs <- plotISCN3(ISCN3 = ISCN3.ls, 'Oak Ridge National Lab_TDE')
```

The Oak Ridge National Lab TDE data set is published and cited.
See @Froberg2008, https://search.dataone.org/view/doi%3A10.3334%2FCDIAC%2FVRC.002, and 
http://cdiac.ess-dive.lbl.gov/epubs/ndp/ndp078a/ndp078a.html for additional details

```{r}
study <- ISCN3.ls$study %>% 
  select(dataset_name, citation) %>%
  filter(dataset_name == "Oak Ridge National Lab_TDE")
knitr::kable(study,
  caption = 'Citations for the Oak Ridge National Lab TDE data set.')%>%
  kable_styling()
```

The Oak Ridge National Lab TDE data set contains `r nrow(filter(ISCN3.ls$layer, dataset_name_sub == "Oak Ridge National Lab_TDE"))` layer-level rows and `r nrow(filter(ISCN3.ls$profile, dataset_name_sub == "Oak Ridge National Lab_TDE"))` profile-level information rows after cleaning. 
The data set does not contain profile level information therefore the profile information graphs are omitted.

### Figrues

The soil data in this data set were sampled from Tennessee (see figure below)

```{r}
print(temp.outputs$layerMap) 
```

A summary of the data set and its contacts are shown below.

```{r}
temp.outputs$summaryStudyTable%>%
  kable_styling()
```


The layer data shows 

```{r}
temp.outputs$layerTable%>%
  kable_styling()
```

The histograms for the data set show 

```{r}
print(temp.outputs$profileValueHistograms) 
```

The layer depth plots show 

```{r}
print(temp.outputs$depthValuePlots) 
```

### Changes from ISCN3 to ISCN3.3

    - Empty columns were removed
    - Cryptic `site_note` information in the profile level was also replaced with a more formal citation.
    - Citations were also updated.
    

# Citations and bibliography

Create a table with each dataset and citation, repeat the dataset name as needed for each citation.

```{r}
#need to gap fill these or find the gap filled ones
study <- ISCN3.ls$study %>% 
  select(dataset_name, citation) 
knitr::kable(study,
  caption = 'Citations for each contributing dataset submision (dataname_sub).')%>%
  kable_styling()
```

