---
title: "Example of merging data files"
author: "Kathe Todd-Brown <ktoddbrown@ufl.edu>"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
---

```{r setup}
library(tidyverse)

```

```{r downloadData}
#download and simplify the files to only look at bulk density

#ISCN3Datafile <- '~/Documents/Datasets/ISCN_new/ISCN3_layer.csv'
ISCN.df <- read_delim(file = 'https://portal.edirepository.org/nis/dataviewer?packageid=edi.1160.1&entityid=4af719a84f8981fcc63f1f92760cb253',
                    col_types = cols(.default =col_character()),
                    delim = ';',
                    n_max = 1e4)  %>%
  select("dataset_name_sub", "site_name", "profile_name", "layer_name", #ID columns
         "bd_method", "bd_samp (g cm-3)", "bd_tot (g cm-3)", "bd_whole (g cm-3)", "bd_other (g cm-3)", "bdNRCS_prep_code")# bulk density columns

#Download url
#co_soils_visit <- read.csv('https://apps.fs.usda.gov/fia/datamart/CSV/CO_SOILS_VISIT.csv')
#co_soils_lab <- read.csv('https://apps.fs.usda.gov/fia/datamart/CSV/CO_SOILS_LAB.csv')

FIA.df <- read_csv(file = 'https://apps.fs.usda.gov/fia/datamart/CSV/CO_SOILS_LAB.csv',
                    col_types = cols(.default =col_character())) %>%
  select("CN", "STATECD", "COUNTYCD", "PLOT", "SMPLNNBR", "VSTNBR", "LAYER_TYPE", #ID columns
         "BULK_DENSITY") # bulk density columns
```

# Method 1: no metadata

Assign everything in line via scripts and use the code as documentation.

```{r eval=FALSE}
##RESCUE OLD CODE Do not refer to the dataStructure nor dataDescriptions here
#Transform the data into a set of id-variable-value-unit-method 


data1.df <- FIA.df %>%
  mutate(study_id = 'FIADB_RMRS', table_id = 'Soils_Lab') %>%
  #pivot the values longer
  pivot_longer(cols = valueCols$column_id, names_to = 'column_id', 
               values_to = 'value', values_drop_na = TRUE) %>%
  left_join(FIA.dataDescription, by = join_by(study_id, table_id, column_id))%>%
  mutate(variable = 'bulk_density')
  


ISCN.dataDescription <- ISCN.dataDescription %>%
  pivot_wider(names_from = is_type, values_from=with_entry)

valueRows <- ISCN.dataStructure %>%
  filter(is_type %in% c('value'))

temp2 <- ISCN.dataStructure %>%
              full_join(ISCN.dataDescription)


##CODE BELOW HERE BROKEN##
#Need to figure out how to assign all the units/methods from
#...the dataDescription to each colume and then 
#...link everything up with the variables in dataStructure
temp1 <- ISCN.df %>%
  mutate(study_id = 'ISCN3', table_id = 'layer') %>%
  pivot_longer(cols = ISCN.dataStructure$column_id,
               names_to = 'column_id',
               values_to = 'with_entry',
               values_drop_na = TRUE) %>%
  left_join(ISCN.dataStructure %>%
              full_join(ISCN.dataDescription)) %>%
  select(-column_id) %>%
  pivot_wider(names_from = is_type,
              values_from = with_entry,
              values_fn = list) %>%
  left_join(ISCN.dataDescription) 
  
  
  left_join(ISCN.dataDescription, by = c('study_id', 'table_id', 'column_id')) %>%
  mutate(method = paste0('bdNRCS_prep_code:', bdNRCS_prep_code, "; ", 
                         'bd_method:', bd_method,  "; ", 
                         'metadata_method:', method)) %>%
  select(-c('bd_method', 'bdNRCS_prep_code', 'control_vocabulary')) %>%
  mutate(variable = 'bulk_density')
  
dataFull.df <- data1.df %>%
  bind_rows(data2.df)

```

# Method 2

Leveraging meta data annotations more generally to document transformations.

```{r}

#Transform the data into a set of id-variable-value-unit-method 

FIA.dataStructure <- tribble(~'study_id', ~'table_id', ~'column_id', ~'is_type', ~of_variable,
  'FIADB_RMRS', 'Soils_Lab', 'BULK_DENSITY', 'value', 'Bulk_density'
)

FIA.dataDescription <- tribble(~'study_id', ~'table_id', ~'column_id', ~'is_type', ~'with_entry',
         'FIADB_RMRS', 'Soils_Lab',	'BULK_DENSITY', 'unit', 'g cm^-3',
'FIADB_RMRS','Soils_Lab', 'BULK_DENSITY', 'method','mass per volume of oven-dry soil inclusive of coarse fraction') 


FIA_long.df <- FIA.df %>%
  mutate(study_id = 'FIADB_RMRS', table_id = 'Soils_Lab') %>%
  #pivot the values longer
  pivot_longer(cols = FIA.dataStructure$column_id, names_to = 'column_id', 
               values_to = 'with_entry', values_drop_na = TRUE) %>%
  left_join(FIA.dataStructure, by = join_by(study_id, table_id, column_id)) %>%
  pivot_wider(names_from = is_type, values_from = with_entry) %>%
  left_join(FIA.dataDescription %>%
              pivot_wider(names_from = 'is_type', values_from = 'with_entry'), 
            by = join_by(study_id, table_id, column_id))
  
```

```{r}
ISCN.dataDescription <- tribble(~'study_id', ~'table_id', ~'column_id',~'is_type', ~'with_entry',
'ISCN3',	'layer',	'bd_method',	'description',	"Method used to determine bulk density",
'ISCN3',	'layer',	'bd_samp (g cm-3)',	'description',	'Grams of oven-dried soil per cubic centimeter, with soil particles greater than 2 mm and roots greater that 1 cm diameter removed',
'ISCN3',	'layer',	'bd_samp (g cm-3)',	'unit',	'g cm-3',
'ISCN3',	'layer',	'bd_samp (g cm-3)',	'method',	'Fine earth bulk density',
'ISCN3',	'layer',	'bd_tot (g cm-3)',	'description',	'Grams of oven-dried soil per cubic centimeter, with soil particles greater than 2 mm and roots greater that 1 cm diameter included',
'ISCN3',	'layer',	'bd_tot (g cm-3)',	'unit',	'g cm-3',
'ISCN3',	'layer',	'bd_tot (g cm-3)',	'method',	'Whole soil bulk density',
'ISCN3',	'layer',	'bd_whole (g cm-3)',	'description',	'Grams of oven-dried soil per cubic centimeter, with the content of soil particles >2mm and roots >1cm diameter estimated and subtracted',
'ISCN3',	'layer',	'bd_whole (g cm-3)',	'unit',	'g cm-3',
'ISCN3',	'layer',	'bd_whole (g cm-3)',	'method',	'Fine earth (estimated coarse fraction correction) bulk density',
'ISCN3',	'layer',	'bd_other (g cm-3)',	'description',	'Grams of oven-dried soil per cubic centimeter, method used in the associated Bulk Density Method, soil particle fraction used',
'ISCN3',	'layer',	'bd_other (g cm-3)',	'unit',	'g cm-3',
'ISCN3',	'layer',	'bdNRCS_prep_code',	'description',	'NRCS bulk density method code',
'ISCN3',	'layer',	'bdNRCS_prep_code',	'control_vocabulary',	'NRCS_bulk_density_code')

ISCN.dataStructure <- tribble(~'study_id', ~'table_id', ~'column_id', ~'is_type', ~'of_variable',
  'ISCN3',	'layer',	'bd_method',	'method',	'bulk_density_sample',
  'ISCN3',	'layer',	'bd_method',	'method',	'bulk_density_total',
  'ISCN3',	'layer',	'bd_method',	'method',	'bulk_density_whole',
  'ISCN3',	'layer',	'bd_method',	'method',	'bulk_density_other',
'ISCN3',	'layer',	'bd_samp (g cm-3)',	'value',	'bulk_density_sample',
'ISCN3',	'layer',	'bd_tot (g cm-3)',	'value',	'bulk_density_total',
'ISCN3',	'layer',	'bd_whole (g cm-3)',	'value',	'bulk_density_whole',
'ISCN3',	'layer',	'bd_other (g cm-3)',	'value',	'bulk_density_other',
'ISCN3',	'layer',	'bdNRCS_prep_code',	'method',	'bulk_density_sample',
'ISCN3',	'layer',	'bdNRCS_prep_code',	'method',	'bulk_density_total',
'ISCN3',	'layer',	'bdNRCS_prep_code',	'method',	'bulk_density_whole',
'ISCN3',	'layer',	'bdNRCS_prep_code',	'method',	'bulk_density_other'
)

fullMeta <- ISCN.dataStructure  %>% select(-is_type) %>%
  full_join(ISCN.dataDescription, multiple = 'all') %>%
  full_join(ISCN.df %>%
              mutate(study_id = 'ISCN3', table_id = 'layer') %>%
              pivot_longer(cols = ISCN.dataDescription$column_id, names_to = 'column_id', 
                           values_to = 'with_entry', values_drop_na = TRUE)%>% 
              select(study_id, table_id, column_id, 
                     dataset_name_sub, site_name, profile_name, layer_name) %>%
              unique())

ISCN_long.df <- ISCN.df %>%
  mutate(study_id = 'ISCN3', table_id = 'layer')  %>%
  pivot_longer(cols = ISCN.dataDescription$column_id, names_to = 'column_id', 
               values_to = 'with_entry', values_drop_na = TRUE) %>%
  #workflow 1
  left_join(ISCN.dataStructure, by = join_by(study_id, table_id, column_id), multiple = 'all') %>%
  bind_rows(fullMeta) %>%
  select(-column_id) %>%
  ##pull out the tuple
  pivot_wider(names_from=is_type, values_from='with_entry', values_fn = function(...) paste(..., sep = ',', collapse = ';')) %>%
  filter(!is.na(value))
  
   

ISCN_long.df <- ISCN.df %>%
  mutate(study_id = 'ISCN3', table_id = 'layer')  %>%
  pivot_longer(cols = ISCN.dataDescription$column_id, names_to = 'column_id', 
               values_to = 'with_entry', values_drop_na = TRUE) %>%
  #work flow 2
  left_join(ISCN.dataStructure, 
            by = join_by(study_id, table_id, column_id), multiple = 'all') %>%
  left_join(ISCN.dataDescription, 
            by = join_by(study_id, table_id, column_id),
            suffix = c('.data', '.meta'), multiple = 'all') %>%
  reframe(
    with_entry = c(with_entry.meta, with_entry.data),
    is_type = c(is_type.meta, is_type.data),
    .by = c(dataset_name_sub, site_name, profile_name, layer_name, 
           study_id, table_id, of_variable)
  ) %>%
  unique() %>%
  ##pull out the tuple
  pivot_wider(names_from=is_type, 
              values_from='with_entry', 
              values_fn = function(...) paste(..., sep = ',', collapse = ';')) %>%
  filter(!is.na(value))



```

# Adding Identifying Columns


```{r}

#Transform the data into a set of id-variable-value-unit-method 

FIA.dataStructure <- tribble(~'study_id', ~'table_id', ~'column_id', ~'is_type', ~of_variable,
  'FIADB_RMRS', 'Soils_Lab', 'BULK_DENSITY', 'value', 'Bulk_density',
  'FIADB_RMRS', 'Soils_Lab', 'CN', 'identifier', 'code', 
 'FIADB_RMRS', 'Soils_Lab', 'STATECD', 'identifier', 'state', 
 'FIADB_RMRS', 'Soils_Lab', 'COUNTYCD', 'identifier', 'county', 
 'FIADB_RMRS', 'Soils_Lab', 'PLOT', 'identifier', 'plot', 
 'FIADB_RMRS', 'Soils_Lab', 'SMPLNNBR', 'identifier', 'sample_number', 
 'FIADB_RMRS', 'Soils_Lab', 'VSTNBR', 'identifier', 'station_number'
)

FIA.dataDescription <- tribble(~'study_id', ~'table_id', ~'column_id', ~'is_type', ~'with_entry',
         'FIADB_RMRS', 'Soils_Lab',	'BULK_DENSITY', 'unit', 'g cm-3',
'FIADB_RMRS','Soils_Lab', 'BULK_DENSITY', 'method','mass per volume of oven-dry soil inclusive of coarse fraction') 


FIA_long.df <- FIA.df %>%
  mutate(study_id = 'FIADB_RMRS', 
         table_id = 'Soils_Lab',
         row_num = 1:n()) %>%
  pivot_longer(cols = any_of(FIA.dataStructure$column_id), names_to = 'column_id', 
               values_to = 'with_entry', values_drop_na = TRUE) %>%
  left_join(FIA.dataStructure, 
            by = join_by(study_id, table_id, column_id), multiple = 'all') %>%
  left_join(FIA.dataDescription, 
            by = join_by(study_id, table_id, column_id),
            suffix = c('.data', '.meta'), multiple = 'all') %>%
  reframe(
    with_entry =c(with_entry.meta, with_entry.data),
    is_type =  c(is_type.meta, is_type.data),
    .by = c(study_id, table_id, row_num, of_variable)
  ) %>%
  filter(!is.na(with_entry)) %>%
  unique() 
  
```

```{r}

ISCN.dataDescription <- tribble(~'study_id', ~'table_id', ~'column_id',~'is_type', ~'with_entry',
'ISCN3',	'layer',	'bd_method',	'description',	"Method used to determine bulk density",
'ISCN3',	'layer',	'bd_samp (g cm-3)',	'description',	'Grams of oven-dried soil per cubic centimeter, with soil particles greater than 2 mm and roots greater that 1 cm diameter removed',
'ISCN3',	'layer',	'bd_samp (g cm-3)',	'unit',	'g cm-3',
'ISCN3',	'layer',	'bd_samp (g cm-3)',	'method',	'Fine earth bulk density',
'ISCN3',	'layer',	'bd_tot (g cm-3)',	'description',	'Grams of oven-dried soil per cubic centimeter, with soil particles greater than 2 mm and roots greater that 1 cm diameter included',
'ISCN3',	'layer',	'bd_tot (g cm-3)',	'unit',	'g cm-3',
'ISCN3',	'layer',	'bd_tot (g cm-3)',	'method',	'Whole soil bulk density',
'ISCN3',	'layer',	'bd_whole (g cm-3)',	'description',	'Grams of oven-dried soil per cubic centimeter, with the content of soil particles >2mm and roots >1cm diameter estimated and subtracted',
'ISCN3',	'layer',	'bd_whole (g cm-3)',	'unit',	'g cm-3',
'ISCN3',	'layer',	'bd_whole (g cm-3)',	'method',	'Fine earth (estimated coarse fraction correction) bulk density',
'ISCN3',	'layer',	'bd_other (g cm-3)',	'description',	'Grams of oven-dried soil per cubic centimeter, method used in the associated Bulk Density Method, soil particle fraction used',
'ISCN3',	'layer',	'bd_other (g cm-3)',	'unit',	'g cm-3',
'ISCN3',	'layer',	'bdNRCS_prep_code',	'description',	'NRCS bulk density method code',
'ISCN3',	'layer',	'bdNRCS_prep_code',	'control_vocabulary',	'NRCS_bulk_density_code')

ISCN.dataStructure <- tribble(~'study_id', ~'table_id', ~'column_id', ~'is_type', ~'of_variable',
  'ISCN3',	'layer',	'bd_method',	'method',	'bulk_density_sample',
  'ISCN3',	'layer',	'bd_method',	'method',	'bulk_density_total',
  'ISCN3',	'layer',	'bd_method',	'method',	'bulk_density_whole',
  'ISCN3',	'layer',	'bd_method',	'method',	'bulk_density_other',
'ISCN3',	'layer',	'bd_samp (g cm-3)',	'value',	'bulk_density_sample',
'ISCN3',	'layer',	'bd_tot (g cm-3)',	'value',	'bulk_density_total',
'ISCN3',	'layer',	'bd_whole (g cm-3)',	'value',	'bulk_density_whole',
'ISCN3',	'layer',	'bd_other (g cm-3)',	'value',	'bulk_density_other',
'ISCN3',	'layer',	'bdNRCS_prep_code',	'method',	'bulk_density_sample',
'ISCN3',	'layer',	'bdNRCS_prep_code',	'method',	'bulk_density_total',
'ISCN3',	'layer',	'bdNRCS_prep_code',	'method',	'bulk_density_whole',
'ISCN3',	'layer',	'bdNRCS_prep_code',	'method',	'bulk_density_other',
'ISCN3', 'layer',	'dataset_name_sub',	'identifier',	'data_source',
'ISCN3',	'layer',	'dataset_name_soc',	'identifier',	'soil_organic_carbon',
'ISCN3',	'layer',	'site_name',	'identifier',	'site',
'ISCN3',	'layer',	'profile_name',	'identifier',	'profile',
'ISCN3',	'layer',	'layer_name',	'identifier',	'layer'
)


ISCN_long.df <- ISCN.df %>%
  mutate(study_id = 'ISCN3',
         table_id = 'layer',
         row_num = 1:n())  %>%
  pivot_longer(cols = any_of(ISCN.dataStructure$column_id), names_to = 'column_id', 
               values_to = 'with_entry', values_drop_na = TRUE) %>%
  left_join(ISCN.dataStructure, 
            by = join_by(study_id, table_id, column_id), multiple = 'all') %>%
  left_join(ISCN.dataDescription, 
            by = join_by(study_id, table_id, column_id),
            suffix = c('.data', '.meta'), multiple = 'all') %>%
  reframe(
    with_entry =c(with_entry.meta, with_entry.data),
    is_type =  c(is_type.meta, is_type.data),
    .by = c(study_id, table_id, row_num, of_variable)
  ) %>%
  filter(!is.na(with_entry)) %>%
  unique() 

ISCN_wide_values.df <- ISCN_long.df %>%
  filter(is_type %in% c('identifier', 'value')) %>%
  select(-is_type) %>%
  ##pull out the tuple
  pivot_wider(names_from=of_variable, 
              values_from=with_entry)



```

# Alternative anootation structure


Leveraging meta data annotations more generally to document transformations with the data structures and description tables combined.
We want to be able to specify things like the download url or doi for the table/study with the url/doi being a `of_variable` instead of a `is_type`

```{r}

#Transform the data into a set of id-variable-value-unit-method 

FIA.dataAnnotaiton <- tribble(
  ~'study_id', ~'table_id', ~'column_id',  ~'is_type', ~'of_variable', ~'with_entry',
  'FIADB_RMRS', 'Soils_Lab', NA, 'vaule', 'download_url', 'https://apps.fs.usda.gov/fia/datamart/CSV/CO_SOILS_LAB.csv',
  'FIADB_RMRS', 'Soils_Lab', 'BULK_DENSITY', 'value', 'Bulk_density', NA,
  'FIADB_RMRS', 'Soils_Lab', 'BULK_DENSITY', 'unit', 'Bulk_density', 'g cm^-3',
'FIADB_RMRS','Soils_Lab',    'BULK_DENSITY',  'method', 'Bulk_density', 'mass per volume of oven-dry soil inclusive of coarse fraction'
)


FIA_long.df <- FIA.df %>%
  mutate(study_id = 'FIADB_RMRS', 
         table_id = 'Soils_Lab',
         row_num = 1:n()) %>%
  pivot_longer(cols = -c(study_id, table_id, row_num), names_to = 'column_id', 
               values_to = 'with_entry', values_drop_na = TRUE) %>%
  full_join(FIA.dataAnnotaiton, 
            by = join_by(study_id, table_id, column_id),
            suffix = c('.data', ''), multiple = 'all') %>%
  mutate(
    with_entry = if_else(is.na(with_entry), with_entry.data, with_entry)) %>%
  select(-with_entry.data)
  
```

```{r}

ISCN.dataAnnotation <- ISCN.dataStructure  %>% select(-is_type) %>%
  full_join(ISCN.dataDescription, multiple = 'all') %>%
  bind_rows(ISCN.dataStructure) %>%
  arrange(column_id)

ISCN.dataAnnotation <- tribble(~"study_id",~"table_id",~"column_id",~"of_variable",~"is_type",~"with_entry",
"ISCN3","layer","bdNRCS_prep_code","bulk_density_sample","description","NRCS bulk density method code",
"ISCN3","layer","bdNRCS_prep_code","bulk_density_sample","control_vocabulary","NRCS_bulk_density_code",
"ISCN3","layer","bdNRCS_prep_code","bulk_density_total","description","NRCS bulk density method code",
"ISCN3","layer","bdNRCS_prep_code","bulk_density_total","control_vocabulary","NRCS_bulk_density_code",
"ISCN3","layer","bdNRCS_prep_code","bulk_density_whole","description","NRCS bulk density method code",
"ISCN3","layer","bdNRCS_prep_code","bulk_density_whole","control_vocabulary","NRCS_bulk_density_code",
"ISCN3","layer","bdNRCS_prep_code","bulk_density_other","description","NRCS bulk density method code",
"ISCN3","layer","bdNRCS_prep_code","bulk_density_other","control_vocabulary","NRCS_bulk_density_code",
"ISCN3","layer","bdNRCS_prep_code","bulk_density_sample","method",NA,
"ISCN3","layer","bdNRCS_prep_code","bulk_density_total","method",NA,
"ISCN3","layer","bdNRCS_prep_code","bulk_density_whole","method",NA,
"ISCN3","layer","bdNRCS_prep_code","bulk_density_other","method",NA,
"ISCN3","layer","bd_method","bulk_density_sample","description","Method used to determine bulk density",
"ISCN3","layer","bd_method","bulk_density_total","description","Method used to determine bulk density",
"ISCN3","layer","bd_method","bulk_density_whole","description","Method used to determine bulk density",
"ISCN3","layer","bd_method","bulk_density_other","description","Method used to determine bulk density",
"ISCN3","layer","bd_method","bulk_density_sample","method",NA,
"ISCN3","layer","bd_method","bulk_density_total","method",NA,
"ISCN3","layer","bd_method","bulk_density_whole","method",NA,
"ISCN3","layer","bd_method","bulk_density_other","method",NA,
"ISCN3","layer","bd_other (g cm-3)","bulk_density_other","description","Grams of oven-dried soil per cubic centimeter, method used in the associated Bulk Density Method, soil particle fraction used",
"ISCN3","layer","bd_other (g cm-3)","bulk_density_other","unit","g cm-3",
"ISCN3","layer","bd_other (g cm-3)","bulk_density_other","value",NA,
"ISCN3","layer","bd_samp (g cm-3)","bulk_density_sample","description","Grams of oven-dried soil per cubic centimeter, with soil particles greater than 2 mm and roots greater that 1 cm diameter removed",
"ISCN3","layer","bd_samp (g cm-3)","bulk_density_sample","unit","g cm-3",
"ISCN3","layer","bd_samp (g cm-3)","bulk_density_sample","method","Fine earth bulk density",
"ISCN3","layer","bd_samp (g cm-3)","bulk_density_sample","value",NA,
"ISCN3","layer","bd_tot (g cm-3)","bulk_density_total","description","Grams of oven-dried soil per cubic centimeter, with soil particles greater than 2 mm and roots greater that 1 cm diameter included",
"ISCN3","layer","bd_tot (g cm-3)","bulk_density_total","unit","g cm-3",
"ISCN3","layer","bd_tot (g cm-3)","bulk_density_total","method","Whole soil bulk density",
"ISCN3","layer","bd_tot (g cm-3)","bulk_density_total","value",NA,
"ISCN3","layer","bd_whole (g cm-3)","bulk_density_whole","description","Grams of oven-dried soil per cubic centimeter, with the content of soil particles >2mm and roots >1cm diameter estimated and subtracted",
"ISCN3","layer","bd_whole (g cm-3)","bulk_density_whole","unit","g cm-3",
"ISCN3","layer","bd_whole (g cm-3)","bulk_density_whole","method","Fine earth (estimated coarse fraction correction) bulk density",
"ISCN3","layer","bd_whole (g cm-3)","bulk_density_whole","value",NA)


ISCN_long.df <- ISCN.df %>%
  mutate(study_id = 'ISCN3', table_id = 'layer', row_num = 1:n())  %>%
  pivot_longer(cols = -c(study_id, table_id, row_num), names_to = 'column_id', 
               values_to = 'with_entry', values_drop_na = TRUE) %>%
  full_join(ISCN.dataAnnotation, 
            by = join_by(study_id, table_id, column_id),
            suffix = c('.data', ''), multiple = 'all') %>%
  mutate(
    with_entry = if_else(is.na(with_entry), with_entry.data, with_entry)) %>%
  select(-with_entry.data)
   
```


# Pulling things together: 

```{r}
bigData <- FIA_long.df %>%
  select(study_id, table_id, row_num, of_variable, is_type, with_entry) %>%
  bind_rows(ISCN_long.df %>%
              select(study_id, table_id, row_num, of_variable, is_type, with_entry)) %>%
  unique() #remove duplicates from repeated methods applying to multiple colums

```

## Method 1: harmonizing data types

```{r}


newData.df <- bigData %>%
  mutate(of_variable = recode(of_variable, 
                           'Bulk_density' = 'bulk_density_whole'))

```

## Method 2: annotations harmonziation

```{r}

#bigData %>%
#  select(study_id, table_id, of_variable, is_type, with_entry) %>%
#  filter(is_type %in% c('unit', 'method')) %>%
#  unique() %>% write.csv(row.names = FALSE)

variableMap <- tribble(~"study_id", ~"table_id", ~"of_variable", ~'target_variable', 
"FIADB_RMRS","Soils_Lab","Bulk_density", 'bulk_density_whole_soil',
"ISCN3","layer","bulk_density_sample", 'bulk_density_fine_earth',
"ISCN3","layer","bulk_density_other",'bulk_density_unknown',
"ISCN3","layer","bulk_density_total", 'bulk_density_whole_soil',
"ISCN3","layer","bulk_density_whole", 'bulk_density_whole_soil',
 "FIADB_RMRS", "Soils_Lab", "code", "code", 
 "FIADB_RMRS", "Soils_Lab", "state", "state", 
 "FIADB_RMRS", "Soils_Lab", "county", "county", 
 "FIADB_RMRS", "Soils_Lab", "plot", "plot", 
 "FIADB_RMRS", "Soils_Lab", "sample_number", "sample_number", 
 "FIADB_RMRS", "Soils_Lab", "station_number", "station_number", 
 "ISCN3", "layer", "data_source", "data_source", 
 "ISCN3", "layer", "site", "site", 
 "ISCN3", "layer", "profile", "profile", 
 "ISCN3", "layer", "layer", "layer" )

newData.df <- bigData %>%
  left_join(variableMap, by = join_by(study_id, table_id, of_variable)) %>%
  select(-of_variable) %>%
  rename(of_variable = target_variable)
  
```


# Plot it
```{r}

data_wide_values.df <- newData.df %>%
  filter(is_type %in% c('identifier', 'value')) %>%
  select(-is_type) %>%
  ##pull out the tuple
  pivot_wider(names_from=of_variable, 
              values_from=with_entry) %>%
  mutate(across(.cols = starts_with('bulk_density'), as.numeric))

meta_data <- newData.df %>%
  filter(!is_type %in% c('identifier', 'value')) %>%
  select(study_id, table_id, of_variable, is_type, with_entry) %>%
  unique()

ggplot(data_wide_values.df) +
  geom_histogram(aes(x=bulk_density_whole_soil)) +
  facet_wrap(~study_id, scales='free_y') +
  scale_x_log10() +
  scale_y_log10()

ggplot(data_wide_values.df %>%
         pivot_longer(cols = c(bulk_density_whole_soil, bulk_density_fine_earth, bulk_density_unknown))) +
  geom_histogram(aes(x=value, fill = study_id)) +
  facet_grid(study_id~name, scales = 'free_y') +
  scale_x_log10() +
  scale_y_log10()

```
