---
title: "ISCN3 Rebuild"
author: "Kathe Todd-Brown [[your name here]]]"
date: "3/17/2022"
output: html_document
---

```{r setup}
library(SOCDRaH2)
library(tidyverse)
library(knitr)
library(ggplot2)
library(ggmap)
library(kableExtra)
dataDir <- '~/Documents/Datasets/ISCN'

#knitr::opts_chunk$set(eval=TRUE)
```

```{r load}
ISCN3.ls <- ISCN3_3(data_dir = dataDir) #this will throw an NA coercion warning that you can ignore
```

```{r summarys}
profile_summary <- ISCN3.ls$profile %>%
  select(dataset_name_sub, site_name) %>%
  unique()%>%
  group_by(dataset_name_sub) %>% 
  tally(name = 'profile_site_count')

layer_summary <- ISCN3.ls$layer %>%
  select(dataset_name_sub, site_name) %>%
  unique()%>%
  group_by(dataset_name_sub) %>% 
  tally(name = 'layer_site_count')

table_summary <- ISCN3.ls$study %>% 
  select(dataset_name, `dataset_type (dataset_type)`, dataset_description) %>%
  unique() %>%
  right_join(profile_summary %>%
              full_join(layer_summary, by = 'dataset_name_sub'), by=c('dataset_name' = 'dataset_name_sub'))

knitr::kable(table_summary %>%
    arrange(layer_site_count), 
  caption = 'Site counts for each contributing dataset submision (dataname_sub) in the profile and layer data table.')%>%
  kable_styling()
```

# Data Contributions


## TEMPLATE (`data_name_sub identifier`)

The TEMPLATE dataset is cited or uncited.
This data set is located in the United States and has XXX layer level observations and YYY profile-level observations of ZZZ unique latitude-longitude locations in the layer-level table and AAA in the profile-level table.

*To use this section remove the code chunk eval arguments and set the current_dataset to the right identifier.*

```{r eval=FALSE}
#subset the data
current_dataset <- "UMBS_FASET"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))

locations <- profile %>% 
  select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
  bind_rows(layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)) %>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

### Tables

```{r eval=FALSE}
study %>%
  knitr::kable(caption = paste('Citations for the ', current_dataset, ' data set.'))

layer %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Layer-level catagorical variable counts.')

profile %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Profile-level catagorical variable counts.')
```

### Figures


```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'), #pull out the numerics that aren't lat/long
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) + #make a histrogram
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & #pull out all the numerical colums
                        !starts_with('layer') & #as long as they aren't layer bounds
                        !ends_with('(dec. deg)'), #and aren't lat/lon
                      #create the paired measurement-variable tables
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',), #pull out the layers
                      #create a paired depth-variable
                      values_to='depth', names_to = 'layer_variable')) +
  #make nifty line plots that span the layer boundaries
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  facet_wrap(~variable,scales = 'free')


```

### Fixed issues

*Copy this from the data report*

Below is a list of know issues that were addressed between ISCN 3 and ISCN 3.3

  + In profile, the ISCN SOC to 1m computation rows were removed, getting rid of duplicate data
  
### Future issues

*Copy this from the data report*

None we are aware of

## University of Michigan Biological Station Forest Accelerated Succession ExperimenT (`UMBS_FASET`)

The University of Michigan Biological Station Forest Accelerated Succession Experiment (`UMBS_FASET`) data set is uncited and unpublished.

```{r}
study <- ISCN3.ls$study %>% 
  filter(dataset_name == "UMBS_FASET") %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == "UMBS_FASET")%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == "UMBS_FASET") %>%
  select(where(~any(!is.na(.))))

locations <- profile %>% 
  select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
  bind_rows(layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)) %>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

### Tables

```{r}
knitr::kable(study,
  caption = 'Citations for the UMBS_FASET data set.') %>%
  kable_styling()
```

This data set is located in the United States and has 405 layer-level observations and 137 profile-level observations of 1 unique latitude-longitude location.

```{r}
layer %>% select_if(is.factor) %>%
  mutate(across(.fns = droplevels)) %>%
  summary() %>%
  knitr::kable()
```

```{r}
profile %>% select_if(is.factor) %>%
  mutate(across(.fns = droplevels)) %>%
  summary() %>%
  knitr::kable()
```


### Figures


```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

#remove data from the layer that appears in the profile that is not the top/bottom/soc
#select all numerical data, pivot and histogram
#select all factor data and tally
```

```{r}
ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !starts_with('layer') &
                        !ends_with('(dec. deg)'),
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',),
                      values_to='depth', names_to = 'layer_variable')) +
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  #coord_flip() +
  facet_wrap(~variable,scales = 'free')


ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'),
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) +
  #coord_flip() +
  facet_wrap(~variable, scales = 'free')
```

### Fixed issues

Below is a list of known issues that were addressed between ISCN 3 and ISCN 3.3

  + The UMBS FASET data set is located in the United States and has 405 layer-level observations and 137 profile-level observations of 1 unique latitude-longitude location.
  + `ISCN SOC computation` soil organic carbon calculations were removed and duplicate rows deleted.
  + Values in the `hzn` column of the `layer` table were reassigned to an NA character from "?", a nonspecific character.
  + Values in  `country (country)` column were reassigned to "United States" in the `layer` and `profile` tables because it was the known location of the collection state.
  
### Future issues
  + Citation practices were initially incompatible with CC-BY license for ISCN3; contacting data providers about modifying release for ISCN4
  + Data is unpublished, and references "100 yrs of research." Revisit in ISCN 4. Check and see if data is still being collected at this research site



## Lu_LTER(`Lu_LTER`)

The Lu_LTER data set references @Ping2000, @Ping2002, @Ping2004, @Ping2005, and @Michaelson1996.

```{r}
current_dataset <- "Lu_LTER"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))

locations <- profile %>% 
  select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
  bind_rows(layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)) %>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

### Tables

```{r}
knitr::kable(study,
  caption = paste('Citations for the ', current_dataset, ' data set.')) %>%
  kable_styling()
```

This data set is located in the United States and has 103 layer level observations and 14 profile-level observations of 19 unique latitude-longitude locations.

```{r}
layer %>% select_if(is.factor) %>%
  mutate(across(.fns = droplevels)) %>%
  summary() %>%
  knitr::kable()
```

```{r}
profile %>% select_if(is.factor) %>%
  mutate(across(.fns = droplevels)) %>%
  summary() %>%
  knitr::kable()
```

### Figures


```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

#remove data from the layer that appears in the profile that is not the top/bottom/soc
#select all numerical data, pivot and histogram
#select all factor data and tally
```


```{r}
ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !starts_with('layer') &
                        !ends_with('(dec. deg)'),
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',),
                      values_to='depth', names_to = 'layer_variable')) +
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  #coord_flip() +
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'),
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) +
  #coord_flip() +

```

### Fixed issues

Below is a list of know issues that were addressed between ISCN 3 and ISCN 3.3

  + In profile, the ISCN SOC to 1m computation rows were removed, getting rid of duplicate data
  
### Future issues

None we are aware of


## Schuur (`Schuur`)

The Schuur data set references @Schuur2007.


```{r}
study <- ISCN3.ls$study %>% 
  filter(dataset_name == "Schuur") %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == "Schuur")%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == "Schuur") %>%
  select(where(~any(!is.na(.))))

locations <- profile %>% 
  select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
  bind_rows(layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)) %>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
knitr::kable(study,
  caption = 'Citations for the Schuur data set.') %>%
  kable_styling()
```

<<<<<<< HEAD
This data set is located in the United States and has `r nrow(layer)` layer level observations and `r nrow(profile)` profile-level observations of r nrow(locations)` unique latitude-longitude location.

```{r}
knitr::kable(profile %>% select_if(is.factor))
=======
The `Lu_PIMA` data set in ISCN3 contains 69 layer-level information and 2 profile-level information after cleaning for ISCN3.5.

```{r}
layer %>% select_if(is.factor) %>%
  mutate(across(.fns = droplevels)) %>%
  summary() %>%
  knitr::kable()
```

```{r}
profile %>% select_if(is.factor) %>%
  mutate(across(.fns = droplevels)) %>%
  summary() %>%
  knitr::kable()
>>>>>>> origin/MT_dev
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

#remove data from the layer that appears in the profile that is not the top/bottom/soc
#select all numerical data, pivot and histogram
#select all factor data and tally
```

<<<<<<< HEAD
```{r}
ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !starts_with('layer') &
                        !ends_with('(dec. deg)'),
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',),
                      values_to='depth', names_to = 'layer_variable')) +
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  #coord_flip() +
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'),
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) +
  #coord_flip() +
=======
### Figures
```{r}
ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & 
                      !starts_with('layer') &
                      !ends_with('(dec. deg)'),
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)', 'layer_bot (cm)'), values_to = 'depth', names_to = 'layer_variable')) +
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  #coord_flip() +
  facet_wrap(~variable, scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'),
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) +
  #coord_flip() +
>>>>>>> origin/MT_dev
  facet_wrap(~variable, scales = 'free')
```

<<<<<<< HEAD
### Fixed issues
=======


In profile, the ISCN SOC to 1m computation rows were removed, getting rid of duplicate data. Nonrelevant columns were removed from layer and profile and missing values were filled in. This dataset was set in Alaska, so the Alaska Location code from the Template was used instead of the generic.

### Future issues

X Rename header and file name to reflect dataset_name
X Pull citation and add it to the bib file, check that citation matches the acknowledgment for study
X Check that the clean profile is correct given the original information
X Check that the clean layer is correct given the original information
X Check that maps are reasonable for the location
X Summarize what the modifications were for this particular dataset
X Create an issue on github with the template below
X Commit and link the commit to the issue

## Lu_PIMA

The Lu_PIMA data set references @Ping2010.

The `Lu_PIMA` data set in ISCN3 contains 69 layer-level information and 2 profile-level information after cleaning for ISCN3.5.
Soil carbon columns were all calculated by ISCN3, thus these columns were removed in ISCN3.5 but no other manipulation was required.
>>>>>>> origin/MT_dev

Below is a list of know issues that were addressed between ISCN 3 and ISCN 3.3

  + The ggplot maps were readjusted to accurately show the datapoints and center around Alaska.
  + The layer tables were adjusted to reassign the value "NA" to a column with "?".
  + The clean layer and clean profile table were checked for accuracy of information.
  + A citation was added to the bibliography file.
  + Four columns were removed from the original dataset that remove the soc columns from both the dataset_layer table and dataset_profile table.
  
### Future issues
  + Cite a database referenced in the dataset_study table
      The url links provided lead to a 404 error page.
        The correct url for dataone: https://search.dataone.org/view/knb-lter-bnz.366.16
        The datasite download link provided: http://www.lter.uaf.edu/404.php
          This link leads to a 404 error, contacting the researchers is necessary.

## LuPima ('LuPima')

<<<<<<< HEAD
Please see @Ping2010 for additional details and if you are using ISCN3 please cite.
=======
The Schuur data set references @Schuur2007.
>>>>>>> origin/MT_dev

```{r}
study <- ISCN3.ls$study %>% 
  filter(dataset_name == "Lu_PIMA") %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == "Lu_PIMA")%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == "Lu_PIMA") %>%
  select(where(~any(!is.na(.))))

locations <- profile %>% 
  select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
  bind_rows(layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)) %>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
knitr::kable(study,
  caption = 'Citations for the Lu_PIMA data set.') %>%
  kable_styling()
```

<<<<<<< HEAD
The `Lu_PIMA` data set in ISCN3 contains `r nrow(dataset_layer)` layer-level information and `r nrow(dataset_profile)` profile-level information after cleaning for ISCN3.5.
=======
The `Schuur` data set in ISCN3 contains 153 layer-level information and 4 profile-level information after cleaning for ISCN3.5.
>>>>>>> origin/MT_dev

```{r}
layer %>% select_if(is.factor) %>%
  mutate(across(.fns = droplevels)) %>%
  summary() %>%
  knitr::kable()
```

```{r}
profile %>% select_if(is.factor) %>%
  mutate(across(.fns = droplevels)) %>%
  summary() %>%
  knitr::kable()
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

#remove data from the layer that appears in the profile that is not the top/bottom/soc
#select all numerical data, pivot and histogram
#select all factor data and tally
```

```{r}

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & 
                        !starts_with('layer') &
                        !ends_with('(dec. deg)'),
                      values_to = 'measurements' , names_to = 'variable') %>%
         pivot_longer(cols=c('layer_top (cm)',
                             'layer_bot (cm)'),
                      values_to='depth', names_to = 'layer_variable')) +
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  #coord_flip() +
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'),
                      values_to = 'measurements' , names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) +
  #coord_flip() +
  facet_wrap(~variable, scales = 'free')
```
### Fixed Issues

The `Lu_PIMA` data set in ISCN3 contains `r nrow(dataset_layer)` layer-level information and `r nrow(dataset_profile)` profile-level information after cleaning for ISCN3.5.
Soil carbon columns were all calculated by ISCN3, thus these columns were removed in ISCN3.5 but no other manipulation was required.

### Future Issues


=======
- [ ] Cite a database referenced in the dataset_study table
      The url links provided lead to a 404 error page.
        The correct url for dataone: https://search.dataone.org/view/knb-lter-bnz.366.16
        The datasite download link provided: http://www.lter.uaf.edu/404.php
          This link leads to a 404 error, contacting the researchers is necessary.


