---
title: "ISCN3 Rebuild"
author: "Kathe Todd-Brown, Michael"
date: "3/17/2022"
output: 
  html_document: 
    toc: yes
---

#knitr::opts_chunk$set(eval=TRUE)

Purpose: *This vignette describes the ISCN3 data contributions.*

```{r setup, message=FALSE, warning=FALSE}
library(SOCDRaH2)
library(tidyverse)
library(knitr)
library(ggplot2)
library(ggmap)
library(kableExtra)
dataDir <- '~/Documents/Datasets/ISCN'

knitr::opts_chunk$set(eval=TRUE, echo=FALSE)
```

```{r load}
ISCN3.ls <- ISCN3_3(data_dir = dataDir) #this will throw an NA coercion warning that you can ignore
```

```{r summaries}
profile_summary <- ISCN3.ls$profile %>%
  select(dataset_name_sub, site_name) %>%
  unique()%>%
  group_by(dataset_name_sub) %>% 
  tally(name = 'profile_site_count')

layer_summary <- ISCN3.ls$layer %>%
  select(dataset_name_sub, site_name) %>%
  unique()%>%
  group_by(dataset_name_sub) %>% 
  tally(name = 'layer_site_count')

table_summary <- ISCN3.ls$study %>% 
  select(dataset_name, `dataset_type (dataset_type)`, dataset_description) %>%
  unique() %>%
  right_join(profile_summary %>%
              full_join(layer_summary, by = 'dataset_name_sub'), by=c('dataset_name' = 'dataset_name_sub'))

knitr::kable(table_summary %>%
    arrange(layer_site_count), 
  caption = 'Site counts for each contributing dataset submision (dataname_sub) in the profile and layer data table.')%>%
  kable_styling()
```

# Data Contributions

## TEMPLATE (`data_name_sub identifier`)

The TEMPLATE dataset is cited or uncited.
This data set is located in the United States and has XXX layer-level observations and YYY profile-level observations of ZZZ unique latitude-longitude locations in the layer-level table and AAA in the profile-level table.

*To use this section remove the code chunk eval arguments and set the current_dataset to the right identifier.*

```{r eval=FALSE}
#subset the data
current_dataset <- "Template"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))

locations <- layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)%>%) %>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

if(any(ISCN3.ls$profile$dataset_name_sub == current_dataset)) {
  locations <- profile %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    bind_rows(locations) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))
}

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r eval=FALSE}
study %>%
  knitr::kable(caption = paste('Citations for the ', current_dataset, ' data set.'))

layer %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Layer-level catagorical variable counts.')

if(nrow(profile) != 0) {
  profile %>% select_if(is.factor) %>%
    pivot_longer(cols = everything(), names_to = 'column_name') %>%
    group_by_all() %>%
    tally() %>%
    knitr::kable(caption = 'Profile-level catagorical variable counts.')
}
```

```{r eval=FALSE}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'), #pull out the numerics that aren't lat/long
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) + #make a histrogram
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & #pull out all the numerical colums
                        !starts_with('layer') & #as long as they aren't layer bounds
                        !ends_with('(dec. deg)'), #and aren't lat/lon
                      #create the paired measurement-variable tables
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',), #pull out the layers
                      #create a paired depth-variable
                      values_to='depth', names_to = 'layer_variable')) +
  #make nifty line plots that span the layer boundaries
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  facet_wrap(~variable,scales = 'free')
```

### Fixed issues

*Copy this from the data report*
  
### Future issues

*Copy this from the data report*

None we are aware of

##  National Cooperative Soil Characterization Database (version date: 11 September 2014)(`NRCS Sept/2014`)

Please see [bibtex citation here] for additional details and if you are using ISCN3 please cite.
This data set is located in the United States and has 373326 layer-level observations and 26096 profile-level observations of 44896 unique latitude-longitude locations in the layer-level table and 13016 in the profile-level table.

```{r}
#subset the data
current_dataset <- "NRCS Sept/2014"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))

locations <- layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)%>% #) %>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

if(any(ISCN3.ls$profile$dataset_name_sub == current_dataset)) {
  locations <- profile %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    bind_rows(locations) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))
}

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
study %>%
  knitr::kable(caption = paste('Citations for the ', current_dataset, ' data set.'))

layer %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Layer-level catagorical variable counts.')

if(nrow(profile) != 0) {
  profile %>% select_if(is.factor) %>%
    pivot_longer(cols = everything(), names_to = 'column_name') %>%
    group_by_all() %>%
    tally() %>%
    knitr::kable(caption = 'Profile-level catagorical variable counts.')
}
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'), #pull out the numerics that aren't lat/long
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) + #make a histrogram
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & #pull out all the numerical colums
                        !starts_with('layer') & #as long as they aren't layer bounds
                        !ends_with('(dec. deg)'), #and aren't lat/lon
                      #create the paired measurement-variable tables
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',), #pull out the layers
                      #create a paired depth-variable
                      values_to='depth', names_to = 'layer_variable')) +
  #make nifty line plots that span the layer boundaries
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  facet_wrap(~variable,scales = 'free')
```

### Fixed issues

+ After cleaning for ISCN3.5 this dropped to 372639 layer-level rows and 60895 profile-level rows.
+ In the profile tables 14756 rows were removed from ISCN3 that match the count of the `ISCN SOC stock to 1m computation` flag in the `dataset_name_soc` column implying that when that 1m computation was + calculated, the other values in that row were duplicated.
+ In the layer tables no rows were removed, and the template brakes on this data set. We can't use the gap filling because the rows represent multiple organic carbon methods that we can not gap fill. 
+ Columns were with soil carbon density were removed in the profile and layer datasets since they were entirely filled in by ISCN.  

### Future issues

- We need to revisit datasets with rows removed and reaccess the function.
- Max/min bounds need to be worked out for this data.
- Should also add an evaluation of the number of georeferenced layers


## Worldwide soil carbon and nitrogen data(`Worldwide soil carbon and nitrogen data`)

Please see @Zinke1986 for additional details and if you are using ISCN3 please cite.
This data set is located in the United States and has 21972 layer-level observations and 7423 profile-level observations of 1931 unique latitude-longitude locations in the layer-level table and 1835 in the profile-level table.

```{r}
#subset the data
current_dataset <- "Worldwide soil carbon and nitrogen data"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))

locations <- layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)%>% #) %>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

if(any(ISCN3.ls$profile$dataset_name_sub == current_dataset)) {
  locations <- profile %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    bind_rows(locations) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))
}

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
study %>%
  knitr::kable(caption = paste('Citations for the ', current_dataset, ' data set.'))

layer %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Layer-level catagorical variable counts.')

if(nrow(profile) != 0) {
  profile %>% select_if(is.factor) %>%
    pivot_longer(cols = everything(), names_to = 'column_name') %>%
    group_by_all() %>%
    tally() %>%
    knitr::kable(caption = 'Profile-level catagorical variable counts.')
}
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'), #pull out the numerics that aren't lat/long
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) + #make a histrogram
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & #pull out all the numerical colums
                        !starts_with('layer') & #as long as they aren't layer bounds
                        !ends_with('(dec. deg)'), #and aren't lat/lon
                      #create the paired measurement-variable tables
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',), #pull out the layers
                      #create a paired depth-variable
                      values_to='depth', names_to = 'layer_variable')) +
  #make nifty line plots that span the layer boundaries
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  facet_wrap(~variable,scales = 'free')
```

### Fixed issues

+ Rows were removed from ISCN3 due to being duplicates (a unique function was applied).
+ Columns were removed from ISCN3 if they were SOC calculated.

### Future issues

- download and re-ingest from 10.3334/CDIAC/lue.ndp018


## USGS Site-specific soil C (`USGS_S3C`)

Please see @Buell2004 for additional details.
This data set is located in the United States and has 10434 layer-level observations and 1674 profile-level observations of 1069 unique latitude-longitude locations in the layer-level table and 749 in the profile-level table.

```{r}
#subset the data
current_dataset <- "USGS_S3C"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))

locations <- layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)%>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

if(any(ISCN3.ls$profile$dataset_name_sub == current_dataset)) {
  locations <- profile %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    bind_rows(locations) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))
}

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
study %>%
  knitr::kable(caption = paste('Citations for the ', current_dataset, ' data set.'))

layer %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Layer-level catagorical variable counts.')

if(nrow(profile) != 0) {
  profile %>% select_if(is.factor) %>%
    pivot_longer(cols = everything(), names_to = 'column_name') %>%
    group_by_all() %>%
    tally() %>%
    knitr::kable(caption = 'Profile-level catagorical variable counts.')
}
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'), #pull out the numerics that aren't lat/long
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) + #make a histrogram
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & #pull out all the numerical colums
                        !starts_with('layer') & #as long as they aren't layer bounds
                        !ends_with('(dec. deg)'), #and aren't lat/lon
                      #create the paired measurement-variable tables
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',), #pull out the layers
                      #create a paired depth-variable
                      values_to='depth', names_to = 'layer_variable')) +
  #make nifty line plots that span the layer boundaries
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  facet_wrap(~variable,scales = 'free')
```

### Fixed issues

+ Rows were removed from ISCN3 from the dataset_profile table because it removed duplicated entries.
+ Columns were removed from ISCN3 from the dataset_profile table because these were filled using ISCN soc stock correction.
+ Values in the country (country) column of both the dataset_layer and dataset_profile tables were reassigned to "United States" because this is the known location of the data points collected.
+ Ggplot maps were modified to ensure datasets accurately focused on the Mississippi River Valley region of the United States.
+ Bulk density values, pH values, and soil composition percentages were analyzed to determine the number of extraneous values.
+ 3 extraneous values were found in the ph_h2o column and 1 in bd_tot column of dataset_layer that will need to be revisited in ISCN 4.0 to fix formatting errors.
+ Bounds were checked for caco3 column in dataset_layer, and deemed to be appropriate for the dataset.
  
### Future issues

- download and re-ingest from @Buell2004
- Clean up pH values (3 extraneous), clean up bulk density total (1 extraneous)


## Alaska Deep Soil Carbon Project (`AK DSC Project SOC stock computation`)

Please see @Johnson2011 for additional details
This data set is located in the United States and has 9002 layer-level observations and 3159 profile-level observations of 743 unique latitude-longitude locations in the layer-level table and 743 in the profile-level table.

```{r}
#subset the data
current_dataset <- "AK DSC Project SOC stock computation"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))

locations <- layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)%>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

if(any(ISCN3.ls$profile$dataset_name_sub == current_dataset)) {
  locations <- profile %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    bind_rows(locations) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))
}

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
study %>%
  knitr::kable(caption = paste('Citations for the ', current_dataset, ' data set.'))

layer %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Layer-level catagorical variable counts.')

if(nrow(profile) != 0) {
  profile %>% select_if(is.factor) %>%
    pivot_longer(cols = everything(), names_to = 'column_name') %>%
    group_by_all() %>%
    tally() %>%
    knitr::kable(caption = 'Profile-level catagorical variable counts.')
}
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'), #pull out the numerics that aren't lat/long
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) + #make a histrogram
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & #pull out all the numerical colums
                        !starts_with('layer') & #as long as they aren't layer bounds
                        !ends_with('(dec. deg)'), #and aren't lat/lon
                      #create the paired measurement-variable tables
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',), #pull out the layers
                      #create a paired depth-variable
                      values_to='depth', names_to = 'layer_variable')) +
  #make nifty line plots that span the layer boundaries
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  facet_wrap(~variable,scales = 'free')
```

### Fixed issues

+ Rows were removed from ISCN3 from the dataset_layer table because they were generated using ISCN soc stock correction.
+ Columns were removed from ISCN3 from the dataset_layer table because they were generated using ISCN soc stock correction.
+ Rows were removed from ISCN3 from the dataset_profile layer table because they were generated using ISCN soc stock correction.
+ Columns were removed from ISCN3 from the dataset_profile layer table because of duplicated entries and using ISCN soc stock correction.
+ Ggplot maps were adjusted to focus on the location of Alaska, United States.
+ Longitude and bulk density values were deemed to be appropriate.
+ Dataset_layer values were checked and compared to original to ensure proper cleaning.
  
### Future issues

- download and re-ingest from 10.3334/CDIAC/lue.ndp018


## Jorgensen_NPS (`Jorgensen_NPS`)

Please see @Jorgenson2008 for additional details and if you are using ISCN3 please cite.
This data set is located in the United States and has 2531 layer-level observations and 0 profile-level observations of 530 unique latitude-longitude locations in the layer-level table and 0 in the profile-level table.

```{r}
#subset the data
current_dataset <- "Jorgensen_NPS"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))

locations <- layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)%>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

if(any(ISCN3.ls$profile$dataset_name_sub == current_dataset)) {
  locations <- profile %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    bind_rows(locations) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))
}

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
study %>%
  knitr::kable(caption = paste('Citations for the ', current_dataset, ' data set.'))

layer %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Layer-level catagorical variable counts.')

if(nrow(profile) != 0) {
  profile %>% select_if(is.factor) %>%
    pivot_longer(cols = everything(), names_to = 'column_name') %>%
    group_by_all() %>%
    tally() %>%
    knitr::kable(caption = 'Profile-level catagorical variable counts.')
}
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'), #pull out the numerics that aren't lat/long
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) + #make a histrogram
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & #pull out all the numerical colums
                        !starts_with('layer') & #as long as they aren't layer bounds
                        !ends_with('(dec. deg)'), #and aren't lat/lon
                      #create the paired measurement-variable tables
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',), #pull out the layers
                      #create a paired depth-variable
                      values_to='depth', names_to = 'layer_variable')) +
  #make nifty line plots that span the layer boundaries
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  facet_wrap(~variable,scales = 'free')
```

### Fixed issues

+ Rows weren't removed from ISCN3 because there were no duplicated entries within the layer-level and profile-level information.
+ Columns were removed from ISCN3 in places where information was marked as 'NA' throughout the entirety of the column.
+ Ggplot maps were modified to display a map centered around the Pacific Northwest and a second map focusing on the data collection site in Alaska.
  
### Future issues

- Go to https://www.nps.gov/im/vmi-wrst.htm and reingest the associated soil data.


## Northern Circumpolar Soil Carbon Database (`Northern Circumpolar Soil Carbon Database (NCSCD)`)

Please see @Hugelius2013 for additional details and if you are using ISCN3 please cite.
This data set is located in the United States and has 1907 layer-level observations and 1020 profile-level observations of 476 unique latitude-longitude locations in the layer-level table and 471 in the profile-level table.

```{r}
#subset the data
current_dataset <- "Northern Circumpolar Soil Carbon Database (NCSCD)"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))

locations <- layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)%>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

if(any(ISCN3.ls$profile$dataset_name_sub == current_dataset)) {
  locations <- profile %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    bind_rows(locations) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))
}

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
study %>%
  knitr::kable(caption = paste('Citations for the ', current_dataset, ' data set.'))

layer %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Layer-level catagorical variable counts.')

if(nrow(profile) != 0) {
  profile %>% select_if(is.factor) %>%
    pivot_longer(cols = everything(), names_to = 'column_name') %>%
    group_by_all() %>%
    tally() %>%
    knitr::kable(caption = 'Profile-level catagorical variable counts.')
}
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'), #pull out the numerics that aren't lat/long
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) + #make a histrogram
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & #pull out all the numerical colums
                        !starts_with('layer') & #as long as they aren't layer bounds
                        !ends_with('(dec. deg)'), #and aren't lat/lon
                      #create the paired measurement-variable tables
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',), #pull out the layers
                      #create a paired depth-variable
                      values_to='depth', names_to = 'layer_variable')) +
  #make nifty line plots that span the layer boundaries
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  facet_wrap(~variable,scales = 'free')
```

### Fixed issues

+ Rows were removed from ISCN3 because of duplicate data entries in both layer and profile.
+ Columns were removed from ISCN3 if they were SOC calculations. 
+ Special character '?' in hzn column in layer was reassigned to NA (so the column was then removed entirely.)
  
### Future issues

- check for updated database then download, and re-ingest from @Hugelius2013


## Jorgensen_ARCN (`Jorgensen_ARCN`)

Please see @Jorgenson2008 for additional details.
This data set is located in the United States and has 1108 layer-level observations and 0 profile-level observations of 316 unique latitude-longitude locations in the layer-level table and 0 in the profile-level table.

```{r}
#subset the data
current_dataset <- "Jorgensen_ARCN"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))

locations <- layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)%>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

if(any(ISCN3.ls$profile$dataset_name_sub == current_dataset)) {
  locations <- profile %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    bind_rows(locations) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))
}

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
study %>%
  knitr::kable(caption = paste('Citations for the ', current_dataset, ' data set.'))

layer %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Layer-level catagorical variable counts.')

if(nrow(profile) != 0) {
  profile %>% select_if(is.factor) %>%
    pivot_longer(cols = everything(), names_to = 'column_name') %>%
    group_by_all() %>%
    tally() %>%
    knitr::kable(caption = 'Profile-level catagorical variable counts.')
}
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'), #pull out the numerics that aren't lat/long
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) + #make a histrogram
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & #pull out all the numerical colums
                        !starts_with('layer') & #as long as they aren't layer bounds
                        !ends_with('(dec. deg)'), #and aren't lat/lon
                      #create the paired measurement-variable tables
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',), #pull out the layers
                      #create a paired depth-variable
                      values_to='depth', names_to = 'layer_variable')) +
  #make nifty line plots that span the layer boundaries
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  facet_wrap(~variable,scales = 'free')
```

### Fixed issues

+ Rows weren't removed from ISCN3 because there were no duplicated entries within the layer-level and profile-level information.
+ Columns were removed from ISCN3 in places where information was marked as 'NA' throughout the entirety of the column.
+ Ggplot maps were modified to display a map centered around the Pacific Northwest and a second map focusing on the data collection site in Alaska.
  
### Future issues

- Go to https://www.nps.gov/im/vmi-wrst.htm and reingest the associated soil data.


## Bockheim (`Bockheim`)

Please see Bockheim1998, Bockheim1999, Bockheim2001, Bockheim2003 and Bockheim2004 for additional details.
This data set is located in the United States and has 1611 layer-level observations and 46 profile-level observations of 127 unique latitude-longitude locations in the layer-level table and 24 in the profile-level table.

```{r}
#subset the data
current_dataset <- "Bockheim"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))

locations <- layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)%>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

if(any(ISCN3.ls$profile$dataset_name_sub == current_dataset)) {
  locations <- profile %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    bind_rows(locations) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))
}

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
study %>%
  knitr::kable(caption = paste('Citations for the ', current_dataset, ' data set.'))

layer %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Layer-level catagorical variable counts.')

if(nrow(profile) != 0) {
  profile %>% select_if(is.factor) %>%
    pivot_longer(cols = everything(), names_to = 'column_name') %>%
    group_by_all() %>%
    tally() %>%
    knitr::kable(caption = 'Profile-level catagorical variable counts.')
}
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'), #pull out the numerics that aren't lat/long
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) + #make a histrogram
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & #pull out all the numerical colums
                        !starts_with('layer') & #as long as they aren't layer bounds
                        !ends_with('(dec. deg)'), #and aren't lat/lon
                      #create the paired measurement-variable tables
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',), #pull out the layers
                      #create a paired depth-variable
                      values_to='depth', names_to = 'layer_variable')) +
  #make nifty line plots that span the layer boundaries
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  facet_wrap(~variable,scales = 'free')
```

### Fixed issues

+ Rows were removed from ISCN3 for being duplicates.
+ Columns were removed from ISCN3 for being SOC calculated.
  
### Future issues

- Check that maps are reasonable for the location (wrong because of longitudes)
- Check longitude that looks like it might be wrong (dropped the negative for some)


## Permafrost_RCN (`Permafrost_RCN`)

Please see @Harden2012 for additional details and if you are using ISCN3 please cite.
Cite the individual author if using their data, in addition cite Harden if using that data as part of the broader Permafrost RCN dataset.
This data set is located in the United States and has 3696 layer-level observations and 0 profile-level observations of ZZZ unique latitude-longitude locations in the layer-level table and AAA in the profile-level table.

```{r eval=FALSE}
#subset the data
current_dataset <- "Permafrost_RCN"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))

locations <- layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)%>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

if(any(ISCN3.ls$profile$dataset_name_sub == current_dataset)) {
  locations <- profile %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    bind_rows(locations) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))
}

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
study %>%
  knitr::kable(caption = paste('Citations for the ', current_dataset, ' data set.'))

layer %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Layer-level catagorical variable counts.')

if(nrow(profile) != 0) {
  profile %>% select_if(is.factor) %>%
    pivot_longer(cols = everything(), names_to = 'column_name') %>%
    group_by_all() %>%
    tally() %>%
    knitr::kable(caption = 'Profile-level catagorical variable counts.')
}
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'), #pull out the numerics that aren't lat/long
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) + #make a histrogram
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & #pull out all the numerical colums
                        !starts_with('layer') & #as long as they aren't layer bounds
                        !ends_with('(dec. deg)'), #and aren't lat/lon
                      #create the paired measurement-variable tables
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',), #pull out the layers
                      #create a paired depth-variable
                      values_to='depth', names_to = 'layer_variable')) +
  #make nifty line plots that span the layer boundaries
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  facet_wrap(~variable,scales = 'free')
```

### Fixed issues

+ Rows were removed from ISCN3 because of gap-filling.
+ Columns were removed from ISCN3 because they were entirely NA values, and the `elevation (m)` column in the profile-level data was innacurately filled with "10000."
  
### Future issues

- fill in the lat-long and other data from the profile level information
- contact Schuur and Harden, see if they published the unpublished data they cited
- address elevation being incorrect


## Lehmann_NE_US_soils

## USDA-FS_NRS_LandscapeCarbonInventory

## Jorgensen_YKDE

## Boby_Mack (`Boby_Mack`)

See @Boby2010 for additional details and if you are using ISCN3 please cite.
This data set is located in the United States and has XXX layer level observations and YYY profile-level observations of ZZZ unique latitude-longitude locations in the layer-level table and AAA in the profile-level table.

```{r}
#subset the data
current_dataset <- "Boby_Mack"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))

locations <- layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)%>% #) %>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

if (any(ISCN3.ls$profile$dataset_name_sub == current_dataset)){
locations <- profile %>% 
  select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
  bind_rows(locations) %>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))
}
#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
study %>%
  knitr::kable(caption = paste('Citations for the ', current_dataset, ' data set.'))

layer %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Layer-level catagorical variable counts.')

if (nrow(profile) != 0){
profile %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Profile-level catagorical variable counts.')
}
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'), #pull out the numerics that aren't lat/long
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) + #make a histrogram
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & #pull out all the numerical colums
                        !starts_with('layer') & #as long as they aren't layer bounds
                        !ends_with('(dec. deg)'), #and aren't lat/lon
                      #create the paired measurement-variable tables
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',), #pull out the layers
                      #create a paired depth-variable
                      values_to='depth', names_to = 'layer_variable')) +
  #make nifty line plots that span the layer boundaries
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  facet_wrap(~variable,scales = 'free')
```

### Fixed issues

Below is a list of know issues that were addressed between ISCN 3 and ISCN 3.3
  + Repeated information on soil carbon stocks was removed from the layer and profile information of the data set. 
  + There are no duplicate rows that needed to be removed, however, the soil carbon stock columns gap filled by ISCN were removed by ISCN3.5 cleaning. 
  + The site note in the data set profile was expanded to a formal citation and added to the ISCN bib file.
  + Note that data recorded as "Unknown" or "?" was left as is and NOT shifted to an NA value.
  + In profile, the ISCN SOC to 1m computation rows were removed, getting rid of duplicate data
  
### Future issues

None we are aware of


## Lehmann_Soil_CBC_1

## USGS_Harden


## Bonanza_LTER (`Bonanza LTER`)

The Bonanza_LTER dataset references @Cleve1993 and @Yarie2013.
This data set is located in the United States and has 583 layer level observations and 44 profile-level observations of ZZZ unique latitude-longitude locations in the layer-level table and AAA in the profile-level table.

```{r}
#subset the data
current_dataset <- "Bonanza LTER"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))


locations <- profile %>% 
  select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
  bind_rows(layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)) %>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```



```{r}
study %>%
  knitr::kable(caption = paste('Citations for the ', current_dataset, ' data set.'))

layer %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Layer-level catagorical variable counts.')

profile %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Profile-level catagorical variable counts.')
```




```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'), #pull out the numerics that aren't lat/long
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) + #make a histrogram
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & #pull out all the numerical colums
                        !starts_with('layer') & #as long as they aren't layer bounds
                        !ends_with('(dec. deg)'), #and aren't lat/lon
                      #create the paired measurement-variable tables
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',), #pull out the layers
                      #create a paired depth-variable
                      values_to='depth', names_to = 'layer_variable')) +
  #make nifty line plots that span the layer boundaries
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  facet_wrap(~variable,scales = 'free')
```

### Fixed issues

Below is a list of know issues that were addressed between ISCN 3 and ISCN 3.3

  + In the dataset_profile, SOC computed values were removed and duplicate entries were removed. 
  + In the dataset_layer, SOC computed values were removed.
  
### Future issues

  + Data citation present, import directly from EDI for ISCN4


## Lu_LTER(`Lu_LTER`)

The Lu_LTER data set references @Ping2000, @Ping2002, @Ping2004, @Ping2005, and @Michaelson1996.
The `Lu_LTER` data set in ISCN3 contains 103 layer-level observations and 14 profile-level observations after cleaning for ISCN3.5.

```{r}
current_dataset <- "Lu_LTER"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))

locations <- layer %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

if(any(ISCN3.ls$profile$dataset_name_sub == current_dataset)) {
  locations <- profile %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    bind_rows(locations) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))
}

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
study %>%
  knitr::kable(caption = paste('Citations for the ', current_dataset, ' data set.'))

layer %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Layer-level catagorical variable counts.')

if(nrow(profile) != 0) {
  profile %>% select_if(is.factor) %>%
    pivot_longer(cols = everything(), names_to = 'column_name') %>%
    group_by_all() %>%
    tally() %>%
    knitr::kable(caption = 'Profile-level catagorical variable counts.')
}
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'), #pull out the numerics that aren't lat/long
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) + #make a histrogram
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & #pull out all the numerical colums
                        !starts_with('layer') & #as long as they aren't layer bounds
                        !ends_with('(dec. deg)'), #and aren't lat/lon
                      #create the paired measurement-variable tables
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',), #pull out the layers
                      #create a paired depth-variable
                      values_to='depth', names_to = 'layer_variable')) +
  #make nifty line plots that span the layer boundaries
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  facet_wrap(~variable,scales = 'free')
```

### Fixed issues

Below is a list of know issues that were addressed between ISCN 3 and ISCN 3.3

  + In profile, the ISCN SOC to 1m computation rows were removed, as well as any resulting duplicate data.
  
### Future issues

None we are aware of


## HeckmanSwanstonBiscuitBurn

## Kane (`Kane`)

Please see @Kane2004, @Kane2005, @Kane2009, and @Valentine2004 for additional details.
This data set is located in the United States and has 574 layer level observations and 144 profile-level observations of ZZZ unique latitude-longitude locations in the layer-level table and AAA in the profile-level table.

```{r}
#subset the data
current_dataset <- "Kane"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))

locations <- layer %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    bind_rows(layer %>%
                select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

if(any(ISCN3.ls$profile$dataset_name_sub == current_dataset)) {
  locations <- profile %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    bind_rows(locations) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))
}
#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
study %>%
  knitr::kable(caption = paste('Citations for the ', current_dataset, ' data set.'))

layer %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Layer-level catagorical variable counts.')

if(nrow(profile) != 0) {
  profile %>% select_if(is.factor) %>%
    pivot_longer(cols = everything(), names_to = 'column_name') %>%
    group_by_all() %>%
    tally() %>%
    knitr::kable(caption = 'Profile-level catagorical variable counts.')
}
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'), #pull out the numerics that aren't lat/long
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) + #make a histrogram
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & #pull out all the numerical colums
                        !starts_with('layer') & #as long as they aren't layer bounds
                        !ends_with('(dec. deg)'), #and aren't lat/lon
                      #create the paired measurement-variable tables
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',), #pull out the layers
                      #create a paired depth-variable
                      values_to='depth', names_to = 'layer_variable')) +
  #make nifty line plots that span the layer boundaries
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  facet_wrap(~variable,scales = 'free')
```

### Fixed issues
Below is a list of know issues that were addressed between ISCN 3 and ISCN 3.3
  + Rows weren't removed from ISCN3 because there were no duplicated entries within the layer-level and profile-level information.
  + Columns were removed from ISCN3 where information was not available throughout the data set.
  + Unclear information on vegetation remains in the `vegclass_local` section of the layer-level and profile-level information.
  + This data set required the use of a map centered around Alaska. 
  + In profile, the ISCN SOC to 1m computation rows were removed, getting rid of duplicate data
  
### Future issues
  + Inquire about cryptic vegetation information located in the `vegclass_local` columns
  + Include the two data citations
    The two urls provided lead to 404 error pages.
    citation [2]: https://search.dataone.org/view/https%3A%2F%2Fpasta.lternet.edu%2Fpackage%2Fmetadata%2Feml%2Fknb-lter-bnz%2F133%2F19
                  or
                  https://search.dataone.org/view/https%3A%2F%2Fpasta.lternet.edu%2Fpackage%2Fmetadata%2Feml%2Fknb-lter-bnz%2F133%2F18
                    Looks like same datasets with two different DOI's - contact repository to see if they want this handled a certain way
    citation [4]: https://search.dataone.org/view/https%3A%2F%2Fpasta.lternet.edu%2Fpackage%2Fmetadata%2Feml%2Fknb-lter-bnz%2F132%2F18
          Download URL: https://pasta.lternet.edu/package/data/eml/knb-lter-bnz/132/18/e9bb75742dedc686e224c41ea6e33ad8   - the twelve pits data
                        AND
                        https://pasta.lternet.edu/package/data/eml/knb-lter-bnz/132/18/fc7153b558a9fa8d1f02ff61f8231fde   - biophysical data and GPS coordinates


## Lu_PIMA (`Lu_PIMA`)

The `Lu_PIMA` data set references @Ping2010.
This dataset contains 69 layer-level information and 0 profile-level observations of 11 unique latitude-longitude locations in the layer-level table.

```{r}
#subset the data
current_dataset <- "Lu_PIMA"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))

locations <- layer %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%

    bind_rows(layer %>%
                select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)) %>%

    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

if(any(ISCN3.ls$profile$dataset_name_sub == current_dataset)) {
  locations <- profile %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    bind_rows(locations) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))
}

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
study %>%
  knitr::kable(caption = paste('Citations for the ', current_dataset, ' data set.'))

layer %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Layer-level catagorical variable counts.')

if(nrow(profile) != 0) {
  profile %>% select_if(is.factor) %>%
    pivot_longer(cols = everything(), names_to = 'column_name') %>%
    group_by_all() %>%
    tally() %>%
    knitr::kable(caption = 'Profile-level catagorical variable counts.')
}
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'), #pull out the numerics that aren't lat/long
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) + #make a histrogram
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & #pull out all the numerical colums
                        !starts_with('layer') & #as long as they aren't layer bounds
                        !ends_with('(dec. deg)'), #and aren't lat/lon
                      #create the paired measurement-variable tables
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',), #pull out the layers
                      #create a paired depth-variable
                      values_to='depth', names_to = 'layer_variable')) +
  #make nifty line plots that span the layer boundaries
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  facet_wrap(~variable,scales = 'free')
```

### Fixed issues

  + In profile, the ISCN SOC to 1m computation rows were removed, getting rid of duplicate data.
  + The layer tables were adjusted to reassign the value "NA" to a column with "?".
  + The clean layer and clean profile table were checked for accuracy of information.
  + A citation was added to the bibliography file.
  + Four columns were removed from the original dataset that remove the soc columns from both the dataset_layer table and dataset_profile table.
  
### Future issues
  + Cite a database referenced in the dataset_study table
  + The url links provided lead to a 404 error page. The correct url for dataone: https://search.dataone.org/view/knb-lter-bnz.366.16
  + The datasite download link provided: http://www.lter.uaf.edu/404.php
          This link leads to a 404 error, contacting the researchers is necessary.

## USGS Harden Yazoo (`USGS Harden Yazoo`)

The USGS Harden Yazoo dataset references @Harden1999 and @Huntington1998.
This data set is located in the United States and has 234 layer level observations and 17 profile-level observations of 4 unique latitude-longitude locations in the layer-level table and AAA in the profile-level table.

```{r}
#subset the data
current_dataset <- "USGS Harden Yazoo"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))

locations <- layer %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    bind_rows(layer %>%
                select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

if(any(ISCN3.ls$profile$dataset_name_sub == current_dataset)) {
  locations <- profile %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    bind_rows(locations) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))
}

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
study %>%
  knitr::kable(caption = paste('Citations for the ', current_dataset, ' data set.'))

layer %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Layer-level catagorical variable counts.')

if(nrow(profile) != 0) {
  profile %>% select_if(is.factor) %>%
    pivot_longer(cols = everything(), names_to = 'column_name') %>%
    group_by_all() %>%
    tally() %>%
    knitr::kable(caption = 'Profile-level catagorical variable counts.')
}
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'), #pull out the numerics that aren't lat/long
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) + #make a histrogram
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & #pull out all the numerical colums
                        !starts_with('layer') & #as long as they aren't layer bounds
                        !ends_with('(dec. deg)'), #and aren't lat/lon
                      #create the paired measurement-variable tables
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',), #pull out the layers
                      #create a paired depth-variable
                      values_to='depth', names_to = 'layer_variable')) +
  #make nifty line plots that span the layer boundaries
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  facet_wrap(~variable,scales = 'free')
```

### Fixed issues

  + Half of the rows (234) of dataset_layer were removed from ISCN3 because they were repetitive entries from ISCN stock computation.
  + 3 columns of dataset_layer were removed from ISCN3 because they were determined using soc.
  + 17 rows of dataset_profile were removed from ISCN3 because they were filled using ISCN stock computation.
  + 5 columns of dataset_profile were removed from ISCN 3 because they were determined using soc.
  + Values in column  country (country) were reassigned to "united States" in the dataset_layer and dataset_profile tables because it was the known location of the collection state.
  + Values in the "hzn" column of the dataset_layer table were reassigned to an NA character from "?", a nonspecific character.
  + Ggplot maps were modified to ensure datasets accurately focused on Southeastern United States.
  + Cryptic site notes in dataset_profile table were resolved to contain full citations.
  
### Future issues

  + Fix remaining SOC column in Dataset_layer table?


## Vogel (`Vogel`)

The Vogel dataset references @Kane2009, @Vogel2007, and @Vogel2008.
This data set is located in the United States and has 230 layer level observations and YYY profile-level observations of ZZZ unique latitude-longitude locations in the layer-level table and AAA in the profile-level table.

```{r}
#subset the data
current_dataset <- "Vogel"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))

locations <- layer %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))


if (any(ISCN3.ls$profile$dataset_name_sub == current_dataset)) {
  locations <- profile %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    bind_rows(locations) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))
}

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```



```{r}
study %>%
  knitr::kable(caption = paste('Citations for the ', current_dataset, ' data set.'))

layer %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Layer-level catagorical variable counts.')


if (nrow(profile) != 0)
{
  profile %>% select_if(is.factor) %>%
    pivot_longer(cols = everything(), names_to = 'column_name') %>%
    group_by_all() %>%
    tally() %>%
    knitr::kable(caption = 'Profile-level catagorical variable counts.')
}
```




```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())


ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'), #pull out the numerics that aren't lat/long
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) + #make a histrogram
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & #pull out all the numerical colums
                        !starts_with('layer') & #as long as they aren't layer bounds
                        !ends_with('(dec. deg)'), #and aren't lat/lon
                      #create the paired measurement-variable tables
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',), #pull out the layers
                      #create a paired depth-variable
                      values_to='depth', names_to = 'layer_variable')) +
  #make nifty line plots that span the layer boundaries
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  facet_wrap(~variable,scales = 'free')
```

### Fixed issues

Below is a list of know issues that were addressed between ISCN 3 and ISCN 3.3

  + ISCN3 removed repeated information in citations, which was corrected and filled for ISCN3.5.
  + ISCN3 also filled in the soil carbon stocks where bulk density and organic carbon was provided, this was removed for ISCN3.5.
  + This data set was set in Alaska, so the Alaska Location code from the Template was used instead of the generic.
  
### Future issues

  + Find out where temperature and precipitation data is (currently not on ISCN)

## USGS Muhs (`USGS Muhs`)

The USGS Muhs dataset references @Muhs2003.
This data set is located in the United States and has 395 layer level observations and YYY profile-level observations of 4 unique latitude-longitude locations in the layer-level table and AAA in the profile-level table.

```{r}
#subset the data
current_dataset <- "USGS Muhs"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))

locations <- layer %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

if(any(ISCN3.ls$profile$dataset_name_sub == current_dataset)) {
  locations <- profile %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    bind_rows(locations) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))
}

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
study %>%
  knitr::kable(caption = paste('Citations for the ', current_dataset, ' data set.'))

layer %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Layer-level catagorical variable counts.')

if(nrow(profile) != 0) {
  profile %>% select_if(is.factor) %>%
    pivot_longer(cols = everything(), names_to = 'column_name') %>%
    group_by_all() %>%
    tally() %>%
    knitr::kable(caption = 'Profile-level catagorical variable counts.')
}
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'), #pull out the numerics that aren't lat/long
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) + #make a histrogram
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & #pull out all the numerical colums
                        !starts_with('layer') & #as long as they aren't layer bounds
                        !ends_with('(dec. deg)'), #and aren't lat/lon
                      #create the paired measurement-variable tables
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',), #pull out the layers
                      #create a paired depth-variable
                      values_to='depth', names_to = 'layer_variable')) +
  #make nifty line plots that span the layer boundaries
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  facet_wrap(~variable,scales = 'free')
```

### Fixed issues

  + Repeated information on soil carbon stocks was removed from the layer and profile information of the data set.
  + There are no duplicate rows that needed to be removed, however, the soil carbon stock columns gap filled by ISCN were removed by ISCN3.5 cleaning.
  + In addition, the site note in the data set profile was expanded to a formal citation and added to the ISCN bib file.
  
### Future issues

None we are aware of

## HeckmanLithosequence


## HeckmanLithosequence


## Schuur (`Schuur`)


The Schuur data set references @Schuur2007.
This data set is located in the United States and has 153 layer-level observations and 4 profile-level observations of 1 unique latitude-longitude location in the layer-level table and AAA in the profile-level table.

```{r}
current_dataset <- "Schuur"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))

locations <- layer %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

if(any(ISCN3.ls$profile$dataset_name_sub == current_dataset)) {
  locations <- profile %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    bind_rows(locations) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))
}

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
study %>%
  knitr::kable(caption = paste('Citations for the ', current_dataset, ' data set.'))

layer %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Layer-level catagorical variable counts.')

if(nrow(profile) != 0) {
  profile %>% select_if(is.factor) %>%
    pivot_longer(cols = everything(), names_to = 'column_name') %>%
    group_by_all() %>%
    tally() %>%
    knitr::kable(caption = 'Profile-level catagorical variable counts.')
}
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'), #pull out the numerics that aren't lat/long
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) + #make a histrogram
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & #pull out all the numerical colums
                        !starts_with('layer') & #as long as they aren't layer bounds
                        !ends_with('(dec. deg)'), #and aren't lat/lon
                      #create the paired measurement-variable tables
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',), #pull out the layers
                      #create a paired depth-variable
                      values_to='depth', names_to = 'layer_variable')) +
  #make nifty line plots that span the layer boundaries
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  facet_wrap(~variable,scales = 'free')
```

### Fixed issues

In profile, the ISCN SOC to 1m computation rows were removed, getting rid of duplicate data.

### Future issues

None that we are aware of


## Myers_Smith

## OakRidgeNationalLab_Loblolly_DWJ

## OakRidgeNationalLab_TDE


## University of Michigan Biological Station Forest Accelerated Succession ExperimenT (`UMBS_FASET`)

The University of Michigan Biological Station Forest Accelerated Succession Experiment (`UMBS_FASET`) data set is uncited and unpublished.

This data set is located in the United States and has 405 layer-level observations and 137 profile-level observations of 1 unique latitude-longitude location.

```{r}
#subset the data
current_dataset <- "UMBS_FASET"

study <- ISCN3.ls$study %>% 
  filter(dataset_name == current_dataset) %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == current_dataset)%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == current_dataset) %>%
  select(where(~any(!is.na(.))))

locations <- layer %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    bind_rows(layer %>%
                select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)) %>%

    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

if(any(ISCN3.ls$profile$dataset_name_sub == current_dataset)) {
  locations <- profile %>% 
    select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
    bind_rows(locations) %>%
    unique() %>%
    filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))
}

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
study %>%
  knitr::kable(caption = paste('Citations for the ', current_dataset, ' data set.'))

layer %>% select_if(is.factor) %>%
  pivot_longer(cols = everything(), names_to = 'column_name') %>%
  group_by_all() %>%
  tally() %>%
  knitr::kable(caption = 'Layer-level catagorical variable counts.')

if(nrow(profile) != 0) {
  profile %>% select_if(is.factor) %>%
    pivot_longer(cols = everything(), names_to = 'column_name') %>%
    group_by_all() %>%
    tally() %>%
    knitr::kable(caption = 'Profile-level catagorical variable counts.')
}
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'), #pull out the numerics that aren't lat/long
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) + #make a histrogram
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & #pull out all the numerical colums
                        !starts_with('layer') & #as long as they aren't layer bounds
                        !ends_with('(dec. deg)'), #and aren't lat/lon
                      #create the paired measurement-variable tables
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',), #pull out the layers
                      #create a paired depth-variable
                      values_to='depth', names_to = 'layer_variable')) +
  #make nifty line plots that span the layer boundaries
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  facet_wrap(~variable,scales = 'free')

#remove data from the layer that appears in the profile that is not the top/bottom/soc
#select all numerical data, pivot and histogram
#select all factor data and tally
```

### Fixed issues

Below is a list of known issues that were addressed between ISCN 3 and ISCN 3.3 for the University of Michigan Biological Station Forest Accelerated Succession Experiment (`UMBS_FASET`) data set.

  + The UMBS FASET data set is located in the United States and has 405 layer-level observations and 137 profile-level observations of 1 unique latitude-longitude location.
  + `ISCN SOC computation` soil organic carbon calculations were removed and duplicate rows deleted.
  + Values in the `hzn` column of the `layer` table were reassigned to an NA character from "?", a nonspecific character.
  + Values in  `country (country)` column were reassigned to "United States" in the `layer` and `profile` tables because it was the known location of the collection state.
  
### Future issues
  + Citation practices were initially incompatible with CC-BY license for ISCN3; contacting data providers about modifying release for ISCN4
  + Data is unpublished, and references "100 yrs of research." Revisit in ISCN 4. Check and see if data is still being collected at this research site
