title: "ISCN3 Rebuild"
author: "Kathe Todd-Brown"
date: "3/17/2022"
output: html_document
---

```{r setup}
library(tidyverse)
library(SOCDRaH2)
library(ggplot2)
library(kableExtra)
dataDir <- '~/Desktop/datasets/ISCN'

#knitr::opts_chunk$set(eval=FALSE)
```

```{r load}
ISCN3.ls <- ISCN3_3(data_dir = dataDir) #this will throw an NA coercion warning that you can ignore
```

```{r sites}

profile_summary <- ISCN3.ls$profile %>%
  select(dataset_name_sub, site_name) %>%
  unique()%>%
  group_by(dataset_name_sub) %>% 
  tally(name = 'profile_site_count')

layer_summary <- ISCN3.ls$layer %>%
  select(dataset_name_sub, site_name) %>%
  unique()%>%
  group_by(dataset_name_sub) %>% 
  tally(name = 'layer_site_count')

table_summary <- ISCN3.ls$study %>% 
  select(dataset_name, `dataset_type (dataset_type)`, dataset_description) %>%
  unique() %>%
  right_join(profile_summary %>%
              full_join(layer_summary, by = 'dataset_name_sub'), by=c('dataset_name' = 'dataset_name_sub'))

knitr::kable(table_summary %>%
    arrange(layer_site_count), 
  caption = 'Site counts for each contributing dataset submision (dataname_sub) in the profile and layer data table.')%>%
  kable_styling()
```

# Data Contributions

## University of Michigan Biological Station Forest Accelerated Succession ExperimenT (`UMBS_FASET`)

The University of Michigan Biological Station Forest Accelerated Succession Experiment (`UMBS_FASET`) data set is uncited and unpublished.

```{r}
study <- ISCN3.ls$study %>% 
  filter(dataset_name == "UMBS_FASET") %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == "UMBS_FASET")%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == "UMBS_FASET") %>%
  select(where(~any(!is.na(.))))

locations <- profile %>% 
  select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
  bind_rows(layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)) %>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
knitr::kable(study,
  caption = 'Citations for the UMBS_FASET data set.') %>%
  kable_styling()
```

This data set is located in the United States and has 405 layer-level observations and 137 profile-level observations of 1 unique latitude-longitude location.

```{r}
profile %>% select_if(is.factor) %>%
  mutate(across(.fns = droplevels)) %>%
  summary() %>%
  knitr::kable()
```

```{r}
layer %>% select_if(is.factor) %>%
  mutate(across(.fns = droplevels)) %>%
  summary() %>%
  knitr::kable()
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

#remove data from the layer that appears in the profile that is not the top/bottom/soc
#select all numerical data, pivot and histogram
#select all factor data and tally
```

```{r}
ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !starts_with('layer') &
                        !ends_with('(dec. deg)'),
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',),
                      values_to='depth', names_to = 'layer_variable')) +
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  #coord_flip() +
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'),
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) +
  #coord_flip() +
  facet_wrap(~variable, scales = 'free')
```

### Fixed issues

Below is a list of know issues that were addressed between ISCN 3 and ISCN 3.3

  + The UMBS FASET data set is located in the United States and has 405 layer-level observations and 137 profile-level observations of 1 unique latitude-longitude location.
  + `ISCN SOC computation` soil organic carbon calculations were removed and duplicate rows deleted.
  + Values in the `hzn` column of the `layer` table were reassigned to an NA character from "?", a nonspecific character.
  + Values in  `country (country)` column were reassigned to "United States" in the `layer` and `profile` tables because it was the known location of the collection state.
  
### Future issues
  + Citation practices were initially incompatible with CC-BY license for ISCN3; contacting data providers about modifying release for ISCN4
  + Data is unpublished, and references "100 yrs of research." Revisit in ISCN 4. Check and see if data is still being collected at this research site


## Lu_LTER (`Lu_LTER`)

The Lu_LTER data set references @Ping2000, @Ping2002, @Ping2004, @Ping2005, and @Michaelson1996.

```{r}
study <- ISCN3.ls$study %>% 
  filter(dataset_name == "Lu_LTER") %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == "Lu_LTER")%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == "Lu_LTER") %>%
  select(where(~any(!is.na(.))))

locations <- profile %>% 
  select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
  bind_rows(layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)) %>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
knitr::kable(study,
  caption = 'Citations for the Lu_LTER data set.') %>%
  kable_styling()
```

This data set is located in the United States and has 103 layer level observations and 14 profile-level observations of 19 unique latitude-longitude locations.

```{r}
profile %>% select_if(is.factor) %>%
  mutate(across(.fns = droplevels)) %>%
  summary() %>%
  knitr::kable()
```

```{r}
layer %>% select_if(is.factor) %>%
  mutate(across(.fns = droplevels)) %>%
  summary() %>%
  knitr::kable()
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

#remove data from the layer that appears in the profile that is not the top/bottom/soc
#select all numerical data, pivot and histogram
#select all factor data and tally
```

```{r}
ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !starts_with('layer') &
                        !ends_with('(dec. deg)'),
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',),
                      values_to='depth', names_to = 'layer_variable')) +
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  #coord_flip() +
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'),
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) +
  #coord_flip() +
  facet_wrap(~variable, scales = 'free')
```

### Fixed issues

Below is a list of know issues that were addressed between ISCN 3 and ISCN 3.3

  + In profile, the ISCN SOC to 1m computation rows were removed, getting rid of duplicate data
  + Nonrelevant columns were removed from layer and profile and missing values were filled in
  + This dataset was set in Alaska, so the Alaska Location code from the Template was used instead of the generic.
  
### Future issues
  + Rename header and file name to reflect dataset_name
  + Pull citation and add it to the bib file, check that citation matches the acknowledgment for study
  + Check that the clean profile is correct given the original information
  + Check that the clean layer is correct given the original information
  + Check that maps are reasonable for the location
  + Summarize what the modifications were for this particular dataset
  + Create an issue on github with the template below
  + Commit and link the commit to the issue

## Schuur (`Schuur`)

The Schuur data set references @Schuur2007.

```{r}
study <- ISCN3.ls$study %>% 
  filter(dataset_name == "Schuur") %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == "Schuur")%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == "Schuur") %>%
  select(where(~any(!is.na(.))))

locations <- profile %>% 
  select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
  bind_rows(layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)) %>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
knitr::kable(study,
  caption = 'Citations for the Schuur data set.') %>%
  kable_styling()
```

This data set is located in the United States and has 153 layer level observations and 4 profile-level observations of 1 unique latitude-longitude location.

```{r}
profile %>% select_if(is.factor) %>%
  mutate(across(.fns = droplevels)) %>%
  summary() %>%
  knitr::kable()
```

```{r}
layer %>% select_if(is.factor) %>%
  mutate(across(.fns = droplevels)) %>%
  summary() %>%
  knitr::kable()
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

#remove data from the layer that appears in the profile that is not the top/bottom/soc
#select all numerical data, pivot and histogram
#select all factor data and tally
```

```{r}
ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !starts_with('layer') &
                        !ends_with('(dec. deg)'),
                      values_to = 'measurements', names_to = 'variable') %>%
         pivot_longer(cols = c('layer_top (cm)',
                               'layer_bot (cm)',),
                      values_to='depth', names_to = 'layer_variable')) +
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  #coord_flip() +
  facet_wrap(~variable,scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'),
                      values_to = 'measurements', names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) +
  #coord_flip() +
  facet_wrap(~variable, scales = 'free')
```

### Fixed issues

Below is a list of know issues that were addressed between ISCN 3 and ISCN 3.3

  + The ggplot maps were readjusted to accurately show the datapoints and center around Alaska.
  + The layer tables were adjusted to reassign the value "NA" to a column with "?".
  + The clean layer and clean profile table were checked for accuracy of information.
  + A citation was added to the bibliography file.
  + Four columns were removed from the original dataset that remove the soc columns from both the dataset_layer table and dataset_profile table.
  
### Future issues
  + Cite a database referenced in the dataset_study table
      The url links provided lead to a 404 error page.
        The correct url for dataone: https://search.dataone.org/view/knb-lter-bnz.366.16
        The datasite download link provided: http://www.lter.uaf.edu/404.php
          This link leads to a 404 error, contacting the researchers is necessary.

## LuPima ('LuPima')

Please see @Ping2010 for additional details and if you are using ISCN3 please cite.

```{r}
study <- ISCN3.ls$study %>% 
  filter(dataset_name == "Lu_PIMA") %>%
  select(where(~any(!is.na(.)))) %>%
  t()

profile <- ISCN3.ls$profile %>%
  filter(dataset_name_sub == "Lu_PIMA")%>%
  select(where(~any(!is.na(.))))

layer <- ISCN3.ls$layer %>%
  filter(dataset_name_sub == "Lu_PIMA") %>%
  select(where(~any(!is.na(.))))

locations <- profile %>% 
  select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`) %>%
  bind_rows(layer %>%
              select(`lat (dec. deg)`, `long (dec. deg)`, `datum (datum)`)) %>%
  unique() %>%
  filter(is.finite(`lat (dec. deg)` + `long (dec. deg)`))

#Check to see if there is different/new information in the layer table that is not in the profile
#profile_info_layer <- layer %>%
#  select(any_of(setdiff(names(profile), c("layer_top (cm)", "layer_bot (cm)", "soc (g cm-2)")))) %>%
#  unique()
#
#series <- profile_info_layer %>% 
#  select(profile_name, soil_taxon, soil_series) %>% 
#  unique() %>%
#  anti_join(profile)
```

```{r}
knitr::kable(study,
  caption = 'Citations for the Lu_PIMA data set.') %>%
  kable_styling()
```

This data set is located in the United States and has 69 layer level observations and 2 profile-level observations of 11 unique latitude-longitude locations.

```{r}
profile %>% select_if(is.factor) %>%
  mutate(across(.fns = droplevels)) %>%
  summary() %>%
  knitr::kable()
```

```{r}
layer %>% select_if(is.factor) %>%
  mutate(across(.fns = droplevels)) %>%
  summary() %>%
  knitr::kable()
```

```{r}
ggplot(locations) +
  geom_polygon(data = map_data('world'), aes(x=long, y = lat, group = group)) +
  geom_point(aes(x=`long (dec. deg)`, y = `lat (dec. deg)`), color = 'red', size = 3) +
  theme(axis.title = element_blank())

#remove data from the layer that appears in the profile that is not the top/bottom/soc
#select all numerical data, pivot and histogram
#select all factor data and tally
```

```{r}
ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) &
                        !ends_with('(dec. deg)'),
                      values_to = 'measurements' , names_to = 'variable')) +
  geom_histogram(aes(x = measurements)) +
  #coord_flip() +
  facet_wrap(~variable, scales = 'free')

ggplot(data = layer %>%
         pivot_longer(cols = where(is.numeric) & 
                        !starts_with('layer') &
                        !ends_with('(dec. deg)'),
                      values_to = 'measurements' , names_to = 'variable') %>%
         pivot_longer(cols=c('layer_top (cm)',
                             'layer_bot (cm)'),
                      values_to='depth', names_to = 'layer_variable')) +
  geom_line(aes(x = depth, y = measurements,
                 group = paste(dataset_name_sub, site_name, profile_name))) +
  #coord_flip() +
  facet_wrap(~variable,scales = 'free')
```

### Fixed Issues

  + Soil carbon columns were all calculated by ISCN3, thus these columns were removed in ISCN3.5 but no other manipulation was required

### Future Issues

