# Heckman lithosequence

```{r warning=FALSE, message=FALSE}
datasetName <- 'Heckman lithosequence'

dataset_study <- citation_raw %>% 
  filter(dataset_name == datasetName) %>%
  select(where(function(xx){!(is.na(xx))})) %>%
  full_join(dataset_raw %>% 
              filter(dataset_name == datasetName) %>%
              select(where(function(xx){!(is.na(xx))})), suffix = c('_citation', '_dataset'))%>%
  standardCast()

#comparison for pre ISCN soc stock correction
# dataset_profile_temp <- profile_raw  %>%
#   filter(dataset_name_sub == datasetName) %>%
#   mutate(`country (country)` = 'United States') %>%
#   standardCast()

dataset_profile <- profile_raw  %>%
  filter(dataset_name_sub == datasetName) %>%
  mutate(`country (country)` = 'United States') %>%
  standardCast() %>% #comment out pipe here and take a look at the profile information to figure out how to remove ISCN soc computations
  filter(dataset_name_soc != 'ISCN SOC stock computation') %>% #in this case just remove the whole row, no other data is lost
  select(where(function(xx){!all(is.na(xx))})) #remove the extra soc column lingering from ISCN soc computation

#comparison before SOC correction
# dataset_layer_temp <- layer_raw %>%
#   filter(dataset_name_sub == datasetName) %>%
#   standardCast() %>%
#   mutate(`country (country)` = 'United States')

dataset_layer <- layer_raw %>%
  filter(dataset_name_sub == datasetName) %>%
  standardCast() %>%
  mutate(`country (country)` = 'United States') %>%
  arrange(dataset_name_sub, site_name, profile_name, layer_name, dataset_name_soc) %>% #sort the rows for the fill on the next line
  tidyr::fill(-starts_with('soc'), .direction='up') %>% #fill in the data that was not repeated for the rows not associated with ISCN SOC computations; be careful when reusing this code that you have the right direction set for the fill!
  filter(dataset_name_soc != 'ISCN SOC stock computation') %>% #remove the ISCN SOC stocks
  select(-dataset_name_soc) %>% #we don't care about the flag any more
  select(where(function(xx){!all(is.na(xx))}))
```

The `Heckman lithosequence` data set in ISCN3 contains `r nrow(dataset_layer)` layer-level information and `r nrow(dataset_profile)` profile-level information after cleaning for ISCN3.5.
In ISCN3 the country was not set to United states originally dispite the state being identified as Arizona, ISCN3.5 now identifies this as `'United States'`.
ISCN3 also filled in the soil carbon stocks where bulk density and organic carbon was provided, this was removed for ISCN3.5.

```{r message=FALSE, warning=FALSE}
knitr::kable(t(dataset_study))
```


There are the following factors in the profile:

```{r}
knitr::kable(summary(dataset_profile %>% select_if(is.factor)))
```

And the following factors in the layers:
```{r}
knitr::kable(summary(dataset_layer %>% select_if(is.factor)))
```


## Location

```{r}
ggplot(data =  map_data('state') %>%
         full_join(dataset_profile %>% 
                     mutate(region = 
                              tolower(`state (state_province)`), by = 'region') %>%
                     group_by(region) %>% 
                     tally, by='region')) + 
  geom_polygon(aes(x=long, y = lat, group=group, fill = n), 
               color = 'black') +
  coord_fixed(1.3) + 
  theme_nothing() 

ggplot(data =  map_data('state') %>% 
         filter(region %in% c("arizona", "utah", "new mexico", "colorado"))) + 
  geom_polygon(aes(x=long, y = lat, group = group), 
               fill = 'grey', color = 'black') + 
  geom_point(data=dataset_profile, 
             aes(x = `long (dec. deg)`, y = `lat (dec. deg)`),
             shape = 'x', color = 'red') +
  coord_fixed(1.3) +
  theme_nothing() 

```

```{r eval=FALSE}
#this is useful to see for the analysis but we don't want it in the report
dataset_layer %>%
  pivot_longer(cols = intersect(names(.), type_cols$num_cols), values_drop_na = TRUE) %>%
  group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) %>%
  bind_rows(
    dataset_layer %>%
      pivot_longer(cols = intersect(names(.), type_cols$factor_cols), values_drop_na = TRUE) %>%
      group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) ) %>%
  arrange(n) %>%
  knitr::kable()
```


## Depth plots

```{r}
ggplot(dataset_layer %>% 
         pivot_longer(cols=c('layer_top (cm)', 'layer_bot (cm)'),
                       values_to='depth') %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
         geom_line(aes(x=depth, y= measurement, group = profile_name)) +
  facet_wrap(~type, scales='free')
```

## Citations

Please see @Heckman2009 for additional details and if you are using ISCN3 please cite.
