# Heckman/Swanston Biscuit Burn

```{r}
## set the dataset name gotten from 
#unique(layer_raw$dataset_name_sub)
datasetName <- 'Heckman/Swanston Biscuit Burn'

## subset the citation and dataset description tables and merge them into a study table
dataset_study <- citation_raw %>% 
  filter(dataset_name == datasetName) %>% #pull out the info on the dataset we want here
  select(where(function(xx){!all(is.na(xx))})) %>% #pull out columns where values are not all "NA"
  full_join( #put the the citation and the dataset information together
    dataset_raw %>%
              filter(dataset_name == datasetName) %>% #subset the dataset table
              select(where(function(xx){!all(is.na(xx))})), #only pull columns that are not enitrely NA values
    suffix = c('_citation', '_dataset'))%>% #flag common columns that are not the id columns between the two tables
  standardCast() #go see index.Rmd for this function, mostly recasts the column to the right types
  
## subset the profile and dataset description tables and merge them into a study table
dataset_profile <- profile_raw  %>%
  filter(dataset_name_sub == datasetName) %>% #pull out the info on the dataset we want here
  mutate(`country (country)` = 'United States') %>% #hardcode country to correct one
  standardCast() %>% #recasting column to character type
  rename(dataset_name = 'dataset_name_sub') %>% #simplifying column name
  select(-dataset_name_soc, -starts_with('soc')) #removing ISCN flags that say nothing was done for the SOC fill
         
## subset the layer and dataset description tables and merge them into a study table
dataset_layer <- layer_raw %>%
  filter(dataset_name_sub == datasetName) %>% #pull out info for only this dataset
    standardCast() %>% 
  mutate(`country (country)` = 'United States') %>% #state is identified but not country, fill this in
  rename(dataset_name = 'dataset_name_sub') %>% #changing to more user-friendly name
  select(-dataset_name_soc, -starts_with('soc')) #removing ISCN flags that say nothing was done for the SOC fill
```


The `Heckman/Swanston Biscuit Burn` data set in ISCN3 contains `r nrow(dataset_layer)` layer-level information and `r nrow(dataset_profile)` profile-level information after cleaning for ISCN3.5.
This data set compared data for burned and unburned soils at the site of the 2002 Biscuit Fire.
In ISCN3, the country was not set to United States despite the location of the collected data being identified as the state of Oregon. ISCN3.5 now identifies this as `'United States'`.
ISCN3 filled soil carbon stock computations that created duplicated data entries, which were then removed in ISCN3.5.

```{r}
knitr::kable(t(dataset_study))
```


There are the following factors in the profile:
```{r}
knitr::kable(summary(dataset_profile %>% select_if(is.factor)))
```

And the following factors in the layers:
```{r}
knitr::kable(summary(dataset_layer %>% select_if(is.factor)))
```


## Location

```{r}
##using location data to display what state/region of the United States was sampled
ggplot(data =  map_data('state') %>%  #identifying Oregon as state of interest in this data set
         full_join(dataset_profile %>% 
                     mutate(region = 
                              tolower(`state (state_province)`), by = 'region') %>%
                     group_by(region) %>% 
                     tally, by='region')) + #making state of interest a different color than the rest
  geom_polygon(aes(x=long, y = lat, group=group, fill = n), 
               color = 'black') + #making map outline black
  coord_fixed(1.3) + #fixing ratio  between y and x axes
  theme_nothing() #makes blank ggplot theme

##using location data to plot sites on a state map
ggplot(data =  map_data('state') %>% #pull out relevant info, found in state column
         filter(region %in% c("oregon", "washington"))) + # selecting which states to include on second map
  geom_polygon(aes(x=long, y = lat, group = group), # creating the shape of the map; ‘group’ connects adjacent points with lines to outline shape
               fill = 'grey', color = 'black') +  
  geom_point(data=dataset_profile, # creating scatter plot of latitudes and longitudes from dataset_profile 
             aes(x = `long (dec. deg)`, y = `lat (dec. deg)`), #assigning latitude and longitude to respective axes
             shape = 'x', color = 'red') + # locations will be marked by red x
  coord_fixed(1.3) + #fixing ratio  between y and x axes
  theme_nothing() 
```


```{r}
#this is useful to see for the analysis but we don't want it in the report
#create a table that displays the names of the number and factor columns in the dataset, how many values there are in each, and how many unique values there are
dataset_layer %>% #pipe dataset_layer
  pivot_longer(cols = intersect(names(.), type_cols$num_cols), values_drop_na = TRUE) %>% #pivot the names of columns from dataset_layer with number values into a column called names and the values associated with the column into a column called values (All NA values are dropped)
  group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) %>% #find how many total values and unique values each number column has
  bind_rows( #combine the number and factor column summaries into one dataframe
    dataset_layer %>% #pipe dataset_layer
      pivot_longer(cols = intersect(names(.), type_cols$factor_cols), values_drop_na = TRUE) %>% #pivot the names of columns from dataset_layer with factor values into a column called names and the values associated with the column into a column called values (All NA values are dropped)
      group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) ) %>% #find how many total values and unique values each factor column has
  arrange(n) %>%
  knitr::kable() #create the table
```

## Profile histograms

```{r}
##using profile data to create a histogram for each measured variable in this data set
ggplot(dataset_profile %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), #selecting what columns to pull
                      values_to = 'measurement', names_to = 'type')) + #naming x values and the variables to make histograms of
  geom_histogram(aes(x=measurement)) +  #setting x and y axes
  facet_wrap(~type, scales='free') +
  theme_bw() #using a compatible theme with histogram function
```

## Depth plots

```{r}
##using layer data to create a plot for each measured variable at corresponding depths in this data set
ggplot(dataset_layer %>% 
         pivot_longer( cols=c('layer_top (cm)', 'layer_bot (cm)'),
                       values_to='depth') %>% 
         pivot_longer(cols = c(`oc (percent)`,`13c (‰)`, `14c (‰)`), 
                      values_to = 'measurement', names_to = 'type')) + #naming x values and the variables to make plots of
  geom_line(aes(x=depth, y= measurement, group = profile_name)) + # assigning x and y axis names
  facet_wrap(~type, scales='free')
```

## TODO

 - Follow up with Heckman to see if control at same lat/lon as Burn 7
 - Follow up with Heckman about curator name discrepancy (101 says Kate but this document, 102, says Katherine)


## Citations

Please see @Heckman2013 for additional details and if you are using ISCN3 please cite.
