# Lu_PIMA

In this section, we will pull out the 'Lu_Pima' from ISCN3 and clean it.
Removing duplicate entries/columns that are all NA. 
Ensure all tables created have values, and the NA columns are not included.

    
```{r warning=FALSE, message=FALSe}
##set the dataset name gotten from

#unique(layer_raw$dataset_name_sub)
datasetName <- 'Lu_PIMA'


dataset_study <- citation_raw %>% 
  filter(dataset_name == datasetName) %>% #pull out the info LuPima here
  select(where(function(xx){!(is.na(xx))})) %>% #pull columns that are not all NA values
  full_join(dataset_raw %>% #put the citation and dataset information together 
              filter(dataset_name == datasetName) %>% #subset the dataset table
              select(where(function(xx){!(is.na(xx))})),  #only pull columns that are not NA values entirely
              suffix = c('_citation', '_dataset'))%>% #flag common columns that are not id columns between the two tables
  standardCast() #go see index.Rmd for this function, mostly recasts the column to the right type


dataset_profile <- profile_raw  %>%
  filter(dataset_name_sub == datasetName) %>%
  standardCast() %>% #comment out pipe here and take a look at the profile information to figure out how to remove ISCN soc computations
  filter(dataset_name_soc != 'ISCN SOC stock computation') %>% #in this case just remove the whole row, no other data is lost
  select(where(function(xx){!all(is.na(xx))})) #remove the extra soc column lingering from ISCN soc computation


dataset_layer <- layer_raw %>%
  filter(dataset_name_sub == datasetName) %>%
  standardCast() %>%
  mutate(dataset_name_soc = recode(dataset_name_soc, 'ISCN SOC stock computation' = '_ISCN SOC stock computation')) %>% #recode ISCN so that when we sort it it's on top to fill in the values (see next two lines of code)
  arrange(dataset_name_sub, site_name, profile_name, layer_name, dataset_name_soc) %>% #sort the rows for the fill on the next line
  tidyr::fill(-starts_with('soc'), .direction='down') %>% #fill in the data that was not repeated for the rows not associated with ISCN SOC computations; be careful when reusing this code that you have the right direction set for the fill!
  filter(dataset_name_soc != '_ISCN SOC stock computation') %>% #remove the ISCN SOC stocks
  select(-dataset_name_soc) %>% #we don't care about the flag any more
  select(where(function(xx){!all(is.na(xx))}))
```


## Tables

The `LuPima` data set in ISCN3 contains `r nrow(dataset_layer)` layer-level information and `r nrow(dataset_profile)` profile-level information after cleaning for ISCN3.5.
ISCN3 also filled in the soil carbon stocks where bulk density and organic carbon was provided, this was removed for ISCN3.5.
The 'r datasetname' is located in Alaska, based out of the University of Alaska - Fairbanks on Black Spruce Pines.

```{r message=FALSE, warning=FALSE}
knitr::kable(t(dataset_study))
```

There are the following factors in the profile:

```{r}
knitr::kable(summary(dataset_profile %>% select_if(is.factor)))
```

And the following factors in the layers:
```{r}
knitr::kable(summary(dataset_layer %>% select_if(is.factor)))
```


## Figures

```{r}

#world map with focus on Pacific Ocean/Alaska
ggplot(data = map_data('world2') %>%
            full_join(dataset_profile %>% 
                     mutate(region = 
                              tolower(`state (state_province)`), by = 'region') %>%
                     group_by(region) %>% 
                     tally)) + 
  geom_polygon(aes(x=long, y = lat, group=group, fill = n), 
               color = 'black') +
  coord_fixed(1.3) + 
  theme_nothing() +
  geom_point(data = dataset_layer,
             aes(x = `lat (dec. deg)`, y = `long (dec. deg)`), colour = "purple", size = 1)


#visuals on data from table
ggplot(dataset_layer %>% 
         standardCast() %>%
         pivot_longer( cols=c('layer_top (cm)', 'layer_bot (cm)'),
                       values_to='depth') %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
         geom_line(aes(x=depth, y= measurement, group = profile_name)) +
  facet_wrap(~type, scales='free')
```

## Citations
Please see @Ping2010 for additional details and if you are using ISCN3 please cite.
