Oak Ridge National Lab_Loblolly_DWJ

In this section, we will pull out the `Oak Ridge National Lab_Loblolly_DWJ` from ISCN3 and clean it up.
This dataset requires the country to be set to US. Additionally, the hzn in layer and profile needs to be reset to "NA" when the symbol "?" is present.
In addition, gapfilled bd with sampling was set to "NA" based on method noted.
ISCN SOC stock computation's were removed reducing the duplicated data entries.
Dataset reference noted in a cryptic message, needs to be formly cited using bibtext. (https://doi.org/10.2172/921773)
   
    
```{r message=FALSE, warning=FALSE} 

## set the raw dataset to a variable
datasetName <- 'Oak Ridge National Lab_Loblolly_DWJ'

## subset the citation and dataset description tables and merge them into a study table
dataset_study <- citation_raw %>% 
  filter(dataset_name == datasetName) %>% #pull out the info on the dataset we want here
  select(where(function(xx){!all(is.na(xx))})) %>% #pull out columns where values are not all "NA"
  full_join( #join the selected citation and the dataset data
    dataset_raw %>%
              filter(dataset_name == datasetName) %>% #subset the dataset table
              select(where(function(xx){!all(is.na(xx))})), #only pull columns that are not enitrely NA values
    suffix = c('_citation', '_dataset'))%>% #flag common columns that are not the id columns between the two tables
  standardCast() #reference index.Rmd for this function, mostly recasts the column to the right data types

## subset the profile and dataset description tables and merge them into a single table
dataset_profile <- profile_raw %>%
  filter(dataset_name_sub ==  datasetName) %>% #pull out info from desired dataset
  mutate(`country (country)` = 'United States') %>% #set country name to United States
  standardCast() %>%
  filter(dataset_name_soc != 'ISCN SOC stock computation') %>% #remove gapfilled values
  mutate(`site_note` = case_when(
    site_note == 'see Parr and Hughes 2006' ~ 'Parr, P., & Hughes, J.F. (2006). OAK RIDGE RESERVATION PHYSICAL CHARACTERISTICS AND NATURAL RESOURCES.',
    TRUE ~ site_note))

dataset_layer <- layer_raw %>%
  filter(dataset_name_sub == datasetName) %>% #pull out info from desired dataset
  standardCast() %>%
 mutate(`country (country)` = 'United States', #set country name to United States
         `hzn` = case_when(
           hzn == "?" ~ NA_character_,
           TRUE ~ as.character(hzn)
           ), #replace '?' in hzn with 'NA'
         `soc (g cm-2)` = if_else(dataset_name_soc != 'ISCN SOC stock computation', `soc (g cm-2)`, NA_real_), #changing values in column "soc (g cm-2)" for if a row in "dataset_name_soc" contains "ISCN"....
         `soc_carbon_flag` = if_else(dataset_name_soc != 'ISCN SOC stock computation', `soc_carbon_flag`, NA_character_), #changing values in "soc_carbon_flag" when row contains "ISCN"
         `soc_method` = if_else(dataset_name_soc != 'ISCN SOC stock computation', `soc_method`, NA_character_)) %>% #changing values in "soc_method" when row contains "ISCN"
  group_by(`dataset_name_sub`, `site_name`, `profile_name`, layer_name) %>% #combines existing tables into one table with seperate components
  fill(starts_with('soc')) %>% #select needed data
  filter(dataset_name_soc == 'ISCN SOC stock computation') %>%
 select(-dataset_name_soc) #removes dataset_name_soc
  

```



```{r message=FALSE, warning=FALSE} 

knitr::kable(t(dataset_study))
```


There are the following factors in the profile:

```{r}
knitr::kable(summary(dataset_profile %>% select_if(is.factor)))
```

And the following factors in the layers:
```{r}
knitr::kable(summary(dataset_layer %>% select_if(is.factor)))
```


#Location
Creates a map showing the location of sample taking within specificed state(s) and map of specified state(s) on a United States map.

```{r message=FALSE, warning=FALSE} 

dataset_study 
summary(dataset_profile %>% select_if(is.factor)) #selecting data of type factor
summary(dataset_layer %>% select_if(is.factor)) #selecting data of type factor

ggplot(data =  map_data('state') %>% #specifies map data
         full_join(dataset_profile %>% #selects states the lat/long points are located in by using data_profile data
                     mutate(region = tolower(`state (state_province)`), by = 'region') %>%
                     group_by(region) %>% 
                     tally)) + 
  geom_polygon(aes(x=long, y = lat, group=group, fill = n), #sets attributes/visualization of states on map
               color = 'black') +
  coord_fixed(1.3) + #sets scale for coordinates, making them visible by increasing ratio
  theme_nothing() #sets ggplot theme

ggplot(data =  map_data('state') %>% #selects map data
         filter(region %in% c("tennessee"))) + #filters out the state(s) to be represented 
  geom_polygon(aes(x=long, y = lat, group = group), #defines attributes of state outline/visualization
               fill = 'grey', color = 'black') + 
  geom_point(data=dataset_profile, #specifying location of lat/long data
             aes(x = `long (dec. deg)`, y = `lat (dec. deg)`), #defines attributes and puts lat/long points on graph
             shape = 'x', color = 'red') +
  coord_fixed(1.3) + #sets scale for coordinates, making them visible by increasing ratio
  theme_nothing() #sets ggplot theme
```
#Graphs
Graphs dataset_layer data

```{r message=FALSE, warning=FALSE} 
#Creates long table of the unique features in dataset_layer
dataset_layer %>% 
  pivot_longer(cols = intersect(names(.), type_cols$num_cols), values_drop_na = TRUE) %>% 
  group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) %>%
  bind_rows(
    dataset_layer %>%
      pivot_longer(cols = intersect(names(.), type_cols$factor_cols), values_drop_na = TRUE) %>%
      group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) ) %>%
  arrange(n) %>%
  knitr::kable()

#plots graphs
ggplot(dataset_layer %>% 
         pivot_longer( cols=c('layer_top (cm)', 'layer_bot (cm)'), #pivot data into long table format, specifies column names
                       values_to='depth') %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), #pivot data into long table format
                      values_to = 'measurement', names_to = 'type')) +
         geom_line(aes(x=depth, y= measurement, group = profile_name)) + #set attributes for graphs
  facet_wrap(~type, scales='free') #Wrap a 1d ribbon of panels into 2d
```

#Need to Do 
-reformat graphs and rest of document

## Citations

Please see @Parr2006 for additional details and if you are using ISCN3 please cite.
