# Boby_Mack

```{r warning=FALSE, message=FALSE}
datasetName <- "Boby_Mack" #fix this

##### Extract the study information ####
dataset_study <- citation_raw %>% 
  filter(dataset_name == datasetName) %>%
  select(where(function(xx){!all(is.na(xx))})) %>%
  full_join(dataset_raw %>% 
              filter(dataset_name == datasetName) %>%
              select(where(function(xx){!all(is.na(xx))})), suffix = c('_citation', '_dataset'))%>%
  standardCast()
```

```{r warning=FALSE, message=FALSE}
##### Extract the profile information ####

#comparison for pre ISCN soc stock correction
dataset_profile_org <- profile_raw  %>%
   filter(dataset_name_sub == datasetName) %>%
   standardCast()

dataset_profile <- profile_raw  %>%
  filter(dataset_name_sub == datasetName) 

if(any(grepl('ISCN', dataset_profile$dataset_name_soc))){
  #reassign rows where the ISCN tried to fill in SOC values
  dataset_profile <- dataset_profile %>%
    group_by(dataset_name_soc) %>%
    mutate(`soc_depth (cm)` = if_else(grepl('ISCN', dataset_name_soc),
                                      rep(NA_character_, length(`soc_depth (cm)`)), `soc_depth (cm)`),
           `soc (g cm-2)` = if_else(grepl('ISCN', dataset_name_soc),
                                    rep(NA_character_, length(`soc (g cm-2)`)), `soc (g cm-2)`),
           soc_carbon_flag = if_else(grepl('ISCN', dataset_name_soc),
                                     rep(NA_character_, length(soc_carbon_flag)), soc_carbon_flag),
           soc_spatial_flag = if_else(grepl('ISCN', dataset_name_soc),
                                      rep(NA_character_, length(soc_spatial_flag)), soc_spatial_flag),
           soc_method = if_else(grepl('ISCN', dataset_name_soc), 
                                rep(NA_character_, length(soc_method)), soc_method)) %>%
    ungroup()
  
}

#remove the soc dataset since we've taken care of the ISCN notation
dataset_profile <- select(dataset_profile, -dataset_name_soc)  

if(any(count(dataset_profile, dataset_name_sub, site_name, profile_name)$n > 1)){
  #if the rows are duplicated then fill in missing values by group
  dataset_profile <- dataset_profile %>%
    group_by(dataset_name_sub, site_name, profile_name) %>%
    mutate_at(vars(-group_cols()), 
              function(xx){ifelse(sum(!is.na(xx)) == 1, rep(xx[!is.na(xx)], length(xx)),xx)}) %>% #if there is one value that isn't na then populate the rest of the entry, this fills in the
    ungroup() %>%
    unique() #collapase rows that are non-unique
}

dataset_profile <- standardCast(dataset_profile)
```

```{r warning = FALSE, message = FALSE}
##### Extract the layer infromation ####

#comparison before SOC correction
dataset_layer_org <- layer_raw %>%
  filter(dataset_name_sub == datasetName) %>%
  standardCast()

dataset_layer <- layer_raw %>%
  filter(dataset_name_sub == datasetName) 

if(any(grepl('ISCN', dataset_layer$dataset_name_soc))){
  #reassign rows where the ISCN tried to fill in SOC values
  dataset_layer <- dataset_layer %>%
    group_by(dataset_name_soc) %>%
    mutate(`soc (g cm-2)` = if_else(grepl('ISCN', dataset_name_soc),
                                    rep(NA_character_, length(`soc (g cm-2)`)), `soc (g cm-2)`),
           soc_carbon_flag = if_else(grepl('ISCN', dataset_name_soc),
                                     rep(NA_character_, length(soc_carbon_flag)), soc_carbon_flag),
           soc_method = if_else(grepl('ISCN', dataset_name_soc), 
                                rep(NA_character_, length(soc_method)), soc_method)) %>%
    ungroup()
  
}

#remove the soc dataset since we've taken care of the ISCN notation
dataset_layer <- select(dataset_layer, -dataset_name_soc) 

if(any(count(dataset_layer, dataset_name_sub, site_name, profile_name, layer_name)$n > 1)){
  #if the rows are duplicated then fill in missing values by group
  dataset_layer <- dataset_layer %>%
    group_by(dataset_name_sub, site_name, profile_name, layer_name) %>%
    mutate_at(vars(-group_cols()), 
              function(xx){ifelse(sum(!is.na(xx)) == 1, rep(xx[!is.na(xx)], length(xx)),xx)}) %>% #if there is one value that isn't na then populate the rest of the entry, this fills in the
    ungroup() %>%
    unique() #collapase rows that are non-unique
}

dataset_layer <- standardCast(dataset_layer)

```

The `r datasetName` data set in ISCN3 contains `r nrow(dataset_layer)` layer-level information and `r nrow(dataset_profile)` profile-level information after cleaning for ISCN3.5.

```{r message=FALSE, warning=FALSE}
knitr::kable(t(dataset_study))
```

```{r viz_LP}

#world map
ggplot(data = map_data('world2') %>%
            full_join(dataset_profile %>% 
                     mutate(region = 
                              tolower(`state (state_province)`), by = 'region') %>%
                     group_by(region) %>% 
                     tally)) + 
  geom_polygon(aes(x=long, y = lat, group=group, fill = n), 
               color = 'black') +
  coord_fixed(1.3) + 
  theme_nothing() 


#visuals on data from table
ggplot(dataset_layer %>% 
         standardCast() %>%
         pivot_longer( cols=c('layer_top (cm)', 'layer_bot (cm)'),
                       values_to='depth') %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
         geom_line(aes(x=depth, y= measurement, group = profile_name)) +
  facet_wrap(~type, scales='free')
```

## TODO

- Fix map of Alaska and have points show up on world map
- check table visualizations
- Correct ? in hzm in dataset_layer to NA
- Correct O and unknowns to NA in dataset_layer
- Pull citation and add it to the bib file, check that citation matches the acknowledgement for study
- Summarize what the modifications were for this particular dataset
- Create an issue on github with the template below
- Commit and link the commit to the issue

#### Github issue template

- [ ]  Evaluate dataset
- [ ]  Complete template checklist
- [ ]  Write up summary report narrative
- [ ]  Good commenting
- [ ]  External review by someone else

## Citations

Please see [bibtex citation here] for additional details and if you are using ISCN3 please cite.