# Myers Smith

The Myers Smith data set in ISCN3 contains `r nrow(dataset_layer)` layer-level rows and `r nrow(dataset_profile)` profile-level information rows after cleaning for ISCN3.5.
ISCN3 removed repeated information in citations, so these column values were corrected and filled for ISCN3.5.
ISCN3 filled in the soil carbon stocks where bulk density and organic carbon was provided, so these columns were removed for ISCN3.5.
This data set was set in Alaska, so the Alaska Location code from the Template was used instead of the generic.


```{r warning=FALSE, message=FALSE}
datasetName <- "Myers-Smith" 

##### Extract the study information ####
dataset_study <- citation_raw %>% 
  filter(dataset_name == datasetName) %>%
  select(where(function(xx){!all(is.na(xx))})) %>%
  full_join(dataset_raw %>% 
              filter(dataset_name == datasetName) %>%
              select(where(function(xx){!all(is.na(xx))})), suffix = c('_citation', '_dataset'))%>%
  mutate (`citation`= case_when(`citation`== "Myers-Smith, Isla. 2005. Soil data for cores from a transect from the center of the BBC collapse scar into the surrounding burn, Bonanza Creek LTER - University of Alaska Fairbanks. BNZ:192, http://www.lter.uaf.edu/data_detail.cfm?datafile_pkey=192." ~ "Myers-Smith, Isla. 2005. Soil data for cores from a transect from the center of the BBC collapse scar into the surrounding burn, Bonanza Creek LTER - University of Alaska Fairbanks. BNZ:192, http://www.lter.uaf.edu/data/data-detail/id/192",
            TRUE~ as.character(citation))) %>%
  standardCast() %>%
  fill(curator_name, curator_organization, curator_email, dataset_description, .direction = "downup" )


```

```{r warning=FALSE, message=FALSE}
##### Extract the profile information ####

#comparison for pre ISCN soc stock correction
dataset_profile_org <- profile_raw  %>%
   filter(dataset_name_sub == datasetName) %>%
   standardCast()

dataset_profile <- profile_raw  %>%
  filter(dataset_name_sub == datasetName) 

if(any(grepl('ISCN', dataset_profile$dataset_name_soc))){
  #reassign rows where the ISCN tried to fill in SOC values
  dataset_profile <- dataset_profile %>%
    group_by(dataset_name_soc) %>%
    mutate(`soc_depth (cm)` = if_else(grepl('ISCN', dataset_name_soc),
                                      rep(NA_character_, length(`soc_depth (cm)`)), `soc_depth (cm)`),
           `soc (g cm-2)` = if_else(grepl('ISCN', dataset_name_soc),
                                    rep(NA_character_, length(`soc (g cm-2)`)), `soc (g cm-2)`),
           soc_carbon_flag = if_else(grepl('ISCN', dataset_name_soc),
                                     rep(NA_character_, length(soc_carbon_flag)), soc_carbon_flag),
           soc_spatial_flag = if_else(grepl('ISCN', dataset_name_soc),
                                      rep(NA_character_, length(soc_spatial_flag)), soc_spatial_flag),
           soc_method = if_else(grepl('ISCN', dataset_name_soc), 
                                rep(NA_character_, length(soc_method)), soc_method)) %>%
    ungroup()
  
}

#remove the soc dataset since we've taken care of the ISCN notation
dataset_profile <- select(dataset_profile, -dataset_name_soc)  

if(any(count(dataset_profile, dataset_name_sub, site_name, profile_name)$n > 1)){
  #if the rows are duplicated then fill in missing values by group
  dataset_profile <- dataset_profile %>%
    group_by(dataset_name_sub, site_name, profile_name) %>%
    mutate_at(vars(-group_cols()), 
              function(xx){ifelse(sum(!is.na(xx)) == 1, rep(xx[!is.na(xx)], length(xx)),xx)}) %>% #if there is one value that isn't na then populate the rest of the entry, this fills in the
    ungroup() %>%
    unique() #collapase rows that are non-unique
}

dataset_profile <- standardCast(dataset_profile)
```

```{r warning = FALSE, message = FALSE}
##### Extract the layer infromation ####

#comparison before SOC correction
dataset_layer_org <- layer_raw %>%
  filter(dataset_name_sub == datasetName) %>%
  standardCast()

dataset_layer <- layer_raw %>%
  filter(dataset_name_sub == datasetName) 

if(any(grepl('ISCN', dataset_layer$dataset_name_soc))){
  #reassign rows where the ISCN tried to fill in SOC values
  dataset_layer <- dataset_layer %>%
    group_by(dataset_name_soc) %>%
    mutate(`soc (g cm-2)` = if_else(grepl('ISCN', dataset_name_soc),
                                    rep(NA_character_, length(`soc (g cm-2)`)), `soc (g cm-2)`),
           soc_carbon_flag = if_else(grepl('ISCN', dataset_name_soc),
                                     rep(NA_character_, length(soc_carbon_flag)), soc_carbon_flag),
           soc_method = if_else(grepl('ISCN', dataset_name_soc), 
                                rep(NA_character_, length(soc_method)), soc_method)) %>%
    ungroup()
  
}

#remove the soc dataset since we've taken care of the ISCN notation
dataset_layer <- select(dataset_layer, -dataset_name_soc) 

if(any(count(dataset_layer, dataset_name_sub, site_name, profile_name, layer_name)$n > 1)){
  #if the rows are duplicated then fill in missing values by group
  dataset_layer <- dataset_layer %>%
    group_by(dataset_name_sub, site_name, profile_name, layer_name) %>%
    mutate_at(vars(-group_cols()), 
              function(xx){ifelse(sum(!is.na(xx)) == 1, rep(xx[!is.na(xx)], length(xx)),xx)}) %>% #if there is one value that isn't na then populate the rest of the entry, this fills in the
    ungroup() %>%
    unique() #collapase rows that are non-unique
}

dataset_layer <- standardCast(dataset_layer)

```


```{r message=FALSE, warning=FALSE}
knitr::kable(t(dataset_study))
```


There are the following factors in the profile:

```{r}
knitr::kable(summary(dataset_profile %>% select_if(is.factor)))
```

And the following factors in the layers:
```{r}
knitr::kable(summary(dataset_layer %>% select_if(is.factor)))
```


## Location
```{r}
country <- ggplot2::map_data('world2', 'usa')
avgLat <- dataset_layer %>% 
  pull('lat (dec. deg)') %>%
  mean()

avgLong <- dataset_layer %>% 
  pull('long (dec. deg)') %>%
  mean() + 360

ggplot(data =  map_data("world2")) +
  geom_polygon( aes(x = long, y = lat, group = group),
                fill = 'grey', color = "black") +
  geom_polygon(data = country, aes(x = long, y = lat, group = group),
                fill = 'lightblue', color = "black") +
  geom_point(data = dataset_profile, aes(x = avgLong, y = avgLat), shape = 'x', color = 'red', size = 5) +
  coord_cartesian(xlim=c(avgLong - 65, avgLong + 65), ylim = c(avgLat - 25, avgLat + 25)) +
  theme_nothing() +
  labs(title = 'Profile data')

ggplot(data =  map_data("world2")) + 
  geom_polygon(aes(x=long, y = lat, group = group), 
               fill = 'grey', color = 'black') + 
  geom_polygon(data = country, aes(x = long, y = lat, group = group),
                fill = 'lightblue', color = "black") +
  geom_point(data = dataset_layer, aes(x = `long (dec. deg)` + 360, y = `lat (dec. deg)`),
             shape = 'x', color = 'red', size = 2.5) +
  coord_cartesian(xlim=c(avgLong - 9.1, avgLong + 9.1), ylim = c(avgLat - 3.5, avgLat + 3.5)) +
  theme_nothing() +
  labs(title = 'Layer data')
```

```{r eval=FALSE}
#this is useful to see for the analysis but we don't want it in the report
dataset_layer %>%
  pivot_longer(cols = intersect(names(.), type_cols$num_cols), values_drop_na = TRUE) %>%
  group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) %>%
  bind_rows(
    dataset_layer %>%
      pivot_longer(cols = intersect(names(.), type_cols$factor_cols), values_drop_na = TRUE) %>%
      group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) ) %>%
  arrange(n) %>%
  knitr::kable()
```

## Profile histograms

```{r}
ggplot(dataset_profile %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
  geom_histogram(aes(x=measurement)) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```

## Depth plots

```{r}
ggplot(dataset_layer %>% 
         pivot_longer(cols=c('layer_top (cm)', 'layer_bot (cm)'),
                       values_to='depth') %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
         geom_line(aes(x=depth, y= measurement, group = profile_name), alpha = 0.5) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```

## TODO

- This dataset contains an actual data citation to be re-ingested in next ISCN update

## Citations

Please see @MyersSmithHarden2007 @MyersSmithMcGuire2007 for additional details and if you are using ISCN3 please cite.
