#Bonanza_LTER

```{r warning=FALSE, message=FALSE}
datasetName <- "Bonanza LTER"

##### Extract the study information ####
dataset_study <- citation_raw %>% 
  filter(dataset_name == datasetName) %>%
  select(where(function(xx){!all(is.na(xx))})) %>%
  full_join(dataset_raw %>% 
              filter(dataset_name == datasetName) %>%
              select(where(function(xx){!all(is.na(xx))})), suffix = c('_citation', '_dataset'))%>%
  standardCast()
```

```{r warning=FALSE, message=FALSE}
##### Extract the profile information ####

#comparison for pre ISCN soc stock correction
dataset_profile_org <- profile_raw  %>%
   filter(dataset_name_sub == datasetName) %>%
   standardCast()

dataset_profile <- profile_raw  %>%
  filter(dataset_name_sub == datasetName) 

if(any(grepl('ISCN', dataset_profile$dataset_name_soc))){
  #reassign rows where the ISCN tried to fill in SOC values
  dataset_profile <- dataset_profile %>%
    group_by(dataset_name_soc) %>%
    mutate(`soc_depth (cm)` = if_else(grepl('ISCN', dataset_name_soc),
                                      rep(NA_character_, length(`soc_depth (cm)`)), `soc_depth (cm)`),
           `soc (g cm-2)` = if_else(grepl('ISCN', dataset_name_soc),
                                    rep(NA_character_, length(`soc (g cm-2)`)), `soc (g cm-2)`),
           soc_carbon_flag = if_else(grepl('ISCN', dataset_name_soc),
                                     rep(NA_character_, length(soc_carbon_flag)), soc_carbon_flag),
           soc_spatial_flag = if_else(grepl('ISCN', dataset_name_soc),
                                      rep(NA_character_, length(soc_spatial_flag)), soc_spatial_flag),
           soc_method = if_else(grepl('ISCN', dataset_name_soc), 
                                rep(NA_character_, length(soc_method)), soc_method)) %>%
    ungroup()
  
}

#remove the soc dataset since we've taken care of the ISCN notation
dataset_profile <- select(dataset_profile, -dataset_name_soc)  

if(any(count(dataset_profile, dataset_name_sub, site_name, profile_name)$n > 1)){
  #if the rows are duplicated then fill in missing values by group
  dataset_profile <- dataset_profile %>%
    group_by(dataset_name_sub, site_name, profile_name) %>%
    mutate_at(vars(-group_cols()), 
              function(xx){ifelse(sum(!is.na(xx)) == 1, rep(xx[!is.na(xx)], length(xx)),xx)}) %>% #if there is one value that isn't na then populate the rest of the entry, this fills in the
    ungroup() %>%
    unique() #collapse rows that are non-unique
}

dataset_profile <- standardCast(dataset_profile)
```

```{r warning = FALSE, message = FALSE}
##### Extract the layer information ####

#comparison before SOC correction
dataset_layer_org <- layer_raw %>%
  filter(dataset_name_sub == datasetName) %>%
  standardCast()

dataset_layer <- layer_raw %>%
  filter(dataset_name_sub == datasetName) 

if(any(grepl('ISCN', dataset_layer$dataset_name_soc))){
  #reassign rows where the ISCN tried to fill in SOC values
  dataset_layer <- dataset_layer %>%
    group_by(dataset_name_soc) %>%
    mutate( `hzn` = case_when(
           `hzn` == "?" ~ NA_character_,
           TRUE ~ as.character(hzn)), #hard-coding special character "?" to be NA
           `hzn_desgn` = case_when(
            `hzn_desgn` == "unknown" ~ NA_character_ ,
            TRUE ~ as.character(hzn_desgn)), #hardcoding "unknown" to be NA
      `soc (g cm-2)` = if_else(grepl('ISCN', dataset_name_soc),
                                    rep(NA_character_, length(`soc (g cm-2)`)), `soc (g cm-2)`),
           soc_carbon_flag = if_else(grepl('ISCN', dataset_name_soc),
                                     rep(NA_character_, length(soc_carbon_flag)), soc_carbon_flag),
           soc_method = if_else(grepl('ISCN', dataset_name_soc), 
                                rep(NA_character_, length(soc_method)), soc_method)) %>%
    ungroup()
}

#remove the soc dataset since we've taken care of the ISCN notation
dataset_layer <- select(dataset_layer, -dataset_name_soc) 

if(any(count(dataset_layer, dataset_name_sub, site_name, profile_name, layer_name)$n > 1)){
  #if the rows are duplicated then fill in missing values by group
  dataset_layer <- dataset_layer %>%
    group_by(dataset_name_sub, site_name, profile_name, layer_name) %>%
    mutate_at(vars(-group_cols()), 
              function(xx){ifelse(sum(!is.na(xx)) == 1, rep(xx[!is.na(xx)], length(xx)),xx)}) %>% #if there is one value that isn't na then populate the rest of the entry, this fills in the
    ungroup() %>%
    unique() #collapse rows that are non-unique
}

dataset_layer <- standardCast(dataset_layer)

```

The `Bonanza LTER` data set in ISCN3 contains `r nrow(dataset_layer)` layer-level information and `r nrow(dataset_profile)` profile-level information after cleaning for ISCN3.5. In the dataset_profile, SOC computed values were removed and duplicate entries were removed. In the dataset_layer, SOC computed values were removed.  

```{r message=FALSE, warning=FALSE}
knitr::kable(t(dataset_study))
```


There are the following factors in the profile:

```{r}
knitr::kable(summary(dataset_profile %>% select_if(is.factor)))
```

And the following factors in the layers:
```{r}
knitr::kable(summary(dataset_layer %>% select_if(is.factor)))
```


## Location

```{r}

#ggplot(data =  map_data("world")) + 
#  geom_polygon(aes(x=long, y = lat, group = group), 
#               fill = 'grey', color = 'black') + 
#  geom_point(data= dataset_profile, 
#             aes(x = `long (dec. deg)`, y = `lat (dec. deg)`),
#             shape = 'x', color = 'red') +
#  coord_fixed(1.3) +
#  theme_nothing() +
#  labs(title = 'Profile data') #+
  #coord_map("ortho", orientation = c(90, 0, 0))

#ggplot(data =  map_data("world")) + 
#  geom_polygon(aes(x=long, y = lat, group = group), 
#               fill = 'grey', color = 'black') + 
#  geom_point(data= dataset_layer %>% select(`long (dec. deg)`, `lat (dec. deg)`) %>% unique(), 
#             aes(x = `long (dec. deg)`, y = `lat (dec. deg)`),
#             shape = 'x', color = 'red') +
#  coord_fixed(1.3) +
#  theme_nothing() +
#  labs(title = 'Layer data')

```

## Use for Location instead if dataset is in Alaska
```{r}
#May need to adjust coordinates accordingly (maintain 1.3 ratio for x:y / long:lat)

##reference rectangle
locs<-data.frame(long=c(209,213,213,209),lat=c(64,64,66,66))

country <- ggplot2::map_data('world2', 'usa')

ggplot(data =  map_data("world2")) +
  geom_polygon( aes(x = long, y = lat, group = group),
                fill = 'grey', color = "black") +
  geom_polygon(data = country, aes(x = long, y = lat, group = group),
                fill = 'lightblue', color = "black") +
  coord_cartesian(xlim=c(145, 275), ylim = c(40, 90)) +
  theme_nothing() +
  geom_point(data=locs,aes(long,lat),
             size=0,color= NA) +
  geom_polygon(aes(x = long, y = lat),
               data = locs, color = "darkblue", alpha = .5)+
  labs(title = 'Profile data')

ggplot(data =  map_data("world2")) + 
  geom_polygon(aes(x=long, y = lat, group = group), 
               fill = 'grey', color = 'black') + 
  geom_polygon(data = country, aes(x = long, y = lat, group = group),
                fill = 'lightblue', color = "black") +
  geom_point(data = dataset_layer, aes(x = `long (dec. deg)` + 360, y = `lat (dec. deg)`),
             shape = 'x', color = 'red', size = 2) +
  coord_cartesian(xlim=c(196, 222), ylim = c(60, 70)) +
  theme_nothing() +
geom_point(data=locs,aes(long,lat),
             size=0,color= NA) +
  geom_polygon(aes(x = long, y = lat),
               data = locs, color = "darkblue", alpha = .5)+
  labs(title = 'Layer data')
  
```

```{r eval=FALSE}
#this is useful to see for the analysis but we don't want it in the report
dataset_layer %>%
  pivot_longer(cols = intersect(names(.), type_cols$num_cols), values_drop_na = TRUE) %>%
  group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) %>%
  bind_rows(
    dataset_layer %>%
      pivot_longer(cols = intersect(names(.), type_cols$factor_cols), values_drop_na = TRUE) %>%
      group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) ) %>%
  arrange(n) %>%
  knitr::kable()
```

## Profile histograms

```{r}
ggplot(dataset_profile %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
  geom_histogram(aes(x=measurement)) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```

## Depth plots

```{r}
ggplot(dataset_layer %>% 
         pivot_longer(cols=c('layer_top (cm)', 'layer_bot (cm)'),
                       values_to='depth') %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
         geom_line(aes(x=depth, y= measurement, group = profile_name), alpha = 0.5) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```

## TODO

- Data citation present, import directly from EDI for ISCN4

## Citations


Please see @Cleve1993 and @Yarie2013 for additional details and if you are using ISCN3 please cite.

