# UMBS_FASET

The UMBS_Faset data set in ISCN3 contains 405 layer-level rows and 127 profile-level information rows after cleaning for ISCN3.5.
Rows were removed from ISCN3 because of repeated entries and ISCN SOC computation.
Columns were removed from ISCN3 because of ISCN SOC computation.
Values in the "hzn" column of the dataset_layer table were reassigned to an NA character from "?", a nonspecific character.
Values in column  country (country) were reassigned to "united States" in the dataset_layer and dataset_profile tables because it was the known location of the collection state.
Ggplot maps were modified to ensure datasets accurately focused on Michigan, United States.


```{r warning=FALSE, message=FALSE}
datasetName <- "UMBS_FASET" 

##### Extract the study information ####
dataset_study <- citation_raw %>% 
  filter(dataset_name == datasetName) %>%
  select(where(function(xx){!all(is.na(xx))})) %>%
  full_join(dataset_raw %>% 
              filter(dataset_name == datasetName) %>%
              select(where(function(xx){!all(is.na(xx))})), suffix = c('_citation', '_dataset'))%>%
  standardCast()
```

```{r warning=FALSE, message=FALSE}
##### Extract the profile information ####

#comparison for pre ISCN soc stock correction
dataset_profile_org <- profile_raw  %>%
   filter(dataset_name_sub == datasetName) %>%
   standardCast()

dataset_profile <- profile_raw  %>%
  filter(dataset_name_sub == datasetName) 

if(any(grepl('ISCN', dataset_profile$dataset_name_soc))){
  #reassign rows where the ISCN tried to fill in SOC values
  dataset_profile <- dataset_profile %>%
    group_by(dataset_name_soc) %>%
    mutate(`country (country)` = 'United States', #hard-coding country name as United States
      `soc_depth (cm)` = if_else(grepl('ISCN', dataset_name_soc),
                                      rep(NA_character_, length(`soc_depth (cm)`)), `soc_depth (cm)`),
           `soc (g cm-2)` = if_else(grepl('ISCN', dataset_name_soc),
                                    rep(NA_character_, length(`soc (g cm-2)`)), `soc (g cm-2)`),
           soc_carbon_flag = if_else(grepl('ISCN', dataset_name_soc),
                                     rep(NA_character_, length(soc_carbon_flag)), soc_carbon_flag),
           soc_spatial_flag = if_else(grepl('ISCN', dataset_name_soc),
                                      rep(NA_character_, length(soc_spatial_flag)), soc_spatial_flag),
           soc_method = if_else(grepl('ISCN', dataset_name_soc), 
                                rep(NA_character_, length(soc_method)), soc_method)) %>%
    ungroup()
  
}

#remove the soc dataset since we've taken care of the ISCN notation
dataset_profile <- select(dataset_profile, -dataset_name_soc)  

if(any(count(dataset_profile, dataset_name_sub, site_name, profile_name)$n > 1)){
  #if the rows are duplicated then fill in missing values by group
  dataset_profile <- dataset_profile %>%
    group_by(dataset_name_sub, site_name, profile_name) %>%
    mutate_at(vars(-group_cols()), 
              function(xx){ifelse(sum(!is.na(xx)) == 1, rep(xx[!is.na(xx)], length(xx)),xx)}) %>% #if there is one value that isn't na then populate the rest of the entry, this fills in the
    ungroup() %>%
    unique() #collapase rows that are non-unique
}

dataset_profile <- standardCast(dataset_profile)
```

```{r warning = FALSE, message = FALSE}
##### Extract the layer infromation ####

#comparison before SOC correction
dataset_layer_org <- layer_raw %>%
  filter(dataset_name_sub == datasetName) %>%
  standardCast()

dataset_layer <- layer_raw %>%
  filter(dataset_name_sub == datasetName) 

if(any(grepl('ISCN', dataset_layer$dataset_name_soc))){
  #reassign rows where the ISCN tried to fill in SOC values
  dataset_layer <- dataset_layer %>%
    group_by(dataset_name_soc) %>%
    mutate(`country (country)` = 'United States', #hard-coding country name as United States
          `hzn` = case_when(
           `hzn` == "?" ~ NA_character_,
           TRUE ~ as.character(hzn)), #hard-coding special character "?" to be NA
      `soc (g cm-2)` = if_else(grepl('ISCN', dataset_name_soc),
                                    rep(NA_character_, length(`soc (g cm-2)`)), `soc (g cm-2)`),
           soc_carbon_flag = if_else(grepl('ISCN', dataset_name_soc),
                                     rep(NA_character_, length(soc_carbon_flag)), soc_carbon_flag),
           soc_method = if_else(grepl('ISCN', dataset_name_soc), 
                                rep(NA_character_, length(soc_method)), soc_method)) %>%
    ungroup()
  
}

#remove the soc dataset since we've taken care of the ISCN notation
dataset_layer <- select(dataset_layer, -dataset_name_soc) 

if(any(count(dataset_layer, dataset_name_sub, site_name, profile_name, layer_name)$n > 1)){
  #if the rows are duplicated then fill in missing values by group
  dataset_layer <- dataset_layer %>%
    group_by(dataset_name_sub, site_name, profile_name, layer_name) %>%
    mutate_at(vars(-group_cols()), 
              function(xx){ifelse(sum(!is.na(xx)) == 1, rep(xx[!is.na(xx)], length(xx)),xx)}) %>% #if there is one value that isn't na then populate the rest of the entry, this fills in the
    ungroup() %>%
    unique() #collapase rows that are non-unique
}

dataset_layer <- standardCast(dataset_layer)

```


```{r message=FALSE, warning=FALSE}
knitr::kable(t(dataset_study))
```


There are the following factors in the profile:

```{r}
knitr::kable(summary(dataset_profile %>% select_if(is.factor)))
```

And the following factors in the layers:
```{r}
knitr::kable(summary(dataset_layer %>% select_if(is.factor)))
```


## Location

```{r}

ggplot(data =  map_data("world")) + 
  geom_polygon(aes(x=long, y = lat, group = group), 
               fill = 'grey', color = 'black') + 
  geom_point(data= dataset_profile, 
             aes(x = `long (dec. deg)`, y = `lat (dec. deg)`),
             shape = 'x', color = 'red') +
  coord_fixed(1.3) +
  theme_nothing() +
  labs(title = 'Profile data') #+
  #coord_map("ortho", orientation = c(90, 0, 0))

ggplot(data =  map_data("world")) + 
  geom_polygon(aes(x=long, y = lat, group = group), 
               fill = 'grey', color = 'black') + 
  geom_point(data= dataset_layer %>% select(`long (dec. deg)`, `lat (dec. deg)`) %>% unique(), 
             aes(x = `long (dec. deg)`, y = `lat (dec. deg)`),
             shape = 'x', color = 'red', size = 3) + #increasing size of x's on the map
  coord_cartesian(xlim=c(-100,-70), ylim=c(35,55)) + #shifting second map location to center on Michigan
  theme_nothing() +
  labs(title = 'Layer data')

```

## Use if map is not centered around the specific region of dataset
```{r}

#country <- ggplot2::map_data('world2', 'usa')
#avgLat <- dataset_layer %>% 
#  pull('lat (dec. deg)') %>%
#  mean()

#avgLong <- dataset_layer %>% 
#  pull('long (dec. deg)') %>%
#  mean() + 360

#ggplot(data =  map_data("world2")) +
#  geom_polygon( aes(x = long, y = lat, group = group),
#                fill = 'grey', color = "black") +
#  geom_polygon(data = country, aes(x = long, y = lat, group = group),
#                fill = 'lightblue', color = "black") +
#  geom_point(data = dataset_profile, aes(x = avgLong, y = avgLat), shape = 'x', color = 'red', size = 5) +
#  coord_cartesian(xlim=c(avgLong - 65, avgLong + 65), ylim = c(avgLat - 25, avgLat + 25)) +
#  theme_nothing() +
#  labs(title = 'Profile data')

#ggplot(data =  map_data("world2")) + 
#  geom_polygon(aes(x=long, y = lat, group = group), 
#               fill = 'grey', color = 'black') + 
#  geom_polygon(data = country, aes(x = long, y = lat, group = group),
#                fill = 'lightblue', color = "black") +
#  geom_point(data = dataset_layer, aes(x = `long (dec. deg)` + 360, y = `lat (dec. deg)`),
#             shape = 'x', color = 'red', size = 2.5) +
#  coord_cartesian(xlim=c(avgLong - 9.1, avgLong + 9.1), ylim = c(avgLat - 3.5, avgLat + 3.5)) +
#  theme_nothing() +
#  labs(title = 'Layer data')
  
```

```{r eval=FALSE}
#this is useful to see for the analysis but we don't want it in the report
dataset_layer %>%
  pivot_longer(cols = intersect(names(.), type_cols$num_cols), values_drop_na = TRUE) %>%
  group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) %>%
  bind_rows(
    dataset_layer %>%
      pivot_longer(cols = intersect(names(.), type_cols$factor_cols), values_drop_na = TRUE) %>%
      group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) ) %>%
  arrange(n) %>%
  knitr::kable()
```

## Profile histograms

```{r}
ggplot(dataset_profile %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
  geom_histogram(aes(x=measurement)) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```

## Depth plots

```{r}
ggplot(dataset_layer %>% 
         pivot_longer(cols=c('layer_top (cm)', 'layer_bot (cm)'),
                       values_to='depth') %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
         geom_line(aes(x=depth, y= measurement, group = profile_name), alpha = 0.5) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```

## TODO

- Pull citation and add it to the bib file, check that citation matches the acknowledgement for study
- Commit and link the commit to the issue

#### Github issue template

- [X]  Evaluate dataset
- [ ]  Complete template checklist
- [X]  Write up summary report narrative
- [x]  Good commenting
- [ ]  External review by someone else

## Citations

Please see [bibtex citation here] for additional details and if you are using ISCN3 please cite.
