#Oak Ridge National Lab_TDE

The Oak Ridge National Lab_TDE data set in ISCN3 contains 1176 layer-level rows and 294 profile-level information rows after cleaning for ISCN3.5.
Rows weren't removed from ISCN3 because there were no duplicated entries within the layer-level and profile-level information.
Columns were removed from ISCN3 in places where information was marked as 'NA' throughout the entirety of the column.
A reference and citation in the `dataset_study` information were replaced with a more expanded version. Additionally, cryptic `site_note` information in the profile level was also replaced with a more formal citation.
Ggplot maps were modified to display a map centered around the United States and a second map focusing on the data collection site in Tennessee. 


```{r warning=FALSE, message=FALSE}
datasetName <- "Oak Ridge National Lab_TDE" #fix this

##### Extract the study information ####
dataset_study <- citation_raw %>% 
  filter(dataset_name == datasetName) %>%
  mutate(`reference` = 'Fröberg, M., Hanson, P.J., Todd, D.E. et al. Evaluation of effects of sustained decadal precipitation manipulations on soil carbon stocks. Biogeochemistry 89, 151–161 (2008).' )%>%
  mutate(`citation` = 'Fröberg, M., Hanson, P.J., Todd, D.E. et al. Evaluation of effects of sustained decadal precipitation manipulations on soil carbon stocks. Biogeochemistry 89, 151–161 (2008).' )%>%
  select(where(function(xx){!all(is.na(xx))})) %>%
  full_join(dataset_raw %>% 
              filter(dataset_name == datasetName) %>%
              select(where(function(xx){!all(is.na(xx))})), suffix = c('_citation', '_dataset'))%>%
  standardCast()
```

```{r warning=FALSE, message=FALSE}
##### Extract the profile information ####

#comparison for pre ISCN soc stock correction
dataset_profile_org <- profile_raw  %>%
   filter(dataset_name_sub == datasetName) %>%
   standardCast()

dataset_profile <- profile_raw  %>%
  filter(dataset_name_sub == datasetName) %>%
  mutate(`site_note` = 'Fröberg, M., Hanson, P.J., Todd, D.E. et al. Evaluation of effects of sustained decadal precipitation manipulations on soil carbon stocks. Biogeochemistry 89, 151–161 (2008).')

if(any(grepl('ISCN', dataset_profile$dataset_name_soc))){
  #reassign rows where the ISCN tried to fill in SOC values
  dataset_profile <- dataset_profile %>%
    group_by(dataset_name_soc) %>%
    mutate(`soc_depth (cm)` = if_else(grepl('ISCN', dataset_name_soc),
                                      rep(NA_character_, length(`soc_depth (cm)`)), `soc_depth (cm)`),
           `soc (g cm-2)` = if_else(grepl('ISCN', dataset_name_soc),
                                    rep(NA_character_, length(`soc (g cm-2)`)), `soc (g cm-2)`),
           soc_carbon_flag = if_else(grepl('ISCN', dataset_name_soc),
                                     rep(NA_character_, length(soc_carbon_flag)), soc_carbon_flag),
           soc_spatial_flag = if_else(grepl('ISCN', dataset_name_soc),
                                      rep(NA_character_, length(soc_spatial_flag)), soc_spatial_flag),
           soc_method = if_else(grepl('ISCN', dataset_name_soc), 
                                rep(NA_character_, length(soc_method)), soc_method)) %>%
    ungroup()
  
}

#remove the soc dataset since we've taken care of the ISCN notation
dataset_profile <- select(dataset_profile, -dataset_name_soc)  

if(any(count(dataset_profile, dataset_name_sub, site_name, profile_name)$n > 1)){
  #if the rows are duplicated then fill in missing values by group
  dataset_profile <- dataset_profile %>%
    group_by(dataset_name_sub, site_name, profile_name) %>%
    mutate_at(vars(-group_cols()), 
              function(xx){ifelse(sum(!is.na(xx)) == 1, rep(xx[!is.na(xx)], length(xx)),xx)}) %>% #if there is one value that isn't na then populate the rest of the entry, this fills in the
    ungroup() %>%
    unique() #collapase rows that are non-unique
}

dataset_profile <- standardCast(dataset_profile)
```

```{r warning = FALSE, message = FALSE}
##### Extract the layer infromation ####

#comparison before SOC correction
dataset_layer_org <- layer_raw %>%
  filter(dataset_name_sub == datasetName) %>%
  standardCast()

dataset_layer <- layer_raw %>%
  filter(dataset_name_sub == datasetName) 

if(any(grepl('ISCN', dataset_layer$dataset_name_soc))){
  #reassign rows where the ISCN tried to fill in SOC values
  dataset_layer <- dataset_layer %>%
    group_by(dataset_name_soc) %>%
    mutate(`soc (g cm-2)` = if_else(grepl('ISCN', dataset_name_soc),
                                    rep(NA_character_, length(`soc (g cm-2)`)), `soc (g cm-2)`),
           soc_carbon_flag = if_else(grepl('ISCN', dataset_name_soc),
                                     rep(NA_character_, length(soc_carbon_flag)), soc_carbon_flag),
           soc_method = if_else(grepl('ISCN', dataset_name_soc), 
                                rep(NA_character_, length(soc_method)), soc_method)) %>%
    ungroup()
  
}

#remove the soc dataset since we've taken care of the ISCN notation
dataset_layer <- select(dataset_layer, -dataset_name_soc) 

if(any(count(dataset_layer, dataset_name_sub, site_name, profile_name, layer_name)$n > 1)){
  #if the rows are duplicated then fill in missing values by group
  dataset_layer <- dataset_layer %>%
    group_by(dataset_name_sub, site_name, profile_name, layer_name) %>%
    mutate_at(vars(-group_cols()), 
              function(xx){ifelse(sum(!is.na(xx)) == 1, rep(xx[!is.na(xx)], length(xx)),xx)}) %>% #if there is one value that isn't na then populate the rest of the entry, this fills in the
    ungroup() %>%
    unique() #collapase rows that are non-unique
}

dataset_layer <- standardCast(dataset_layer)

```


```{r message=FALSE, warning=FALSE}
knitr::kable(t(dataset_study))
```


There are the following factors in the profile:

```{r}
knitr::kable(summary(dataset_profile %>% select_if(is.factor)))
```

And the following factors in the layers:
```{r}
knitr::kable(summary(dataset_layer %>% select_if(is.factor)))
```


## Location

```{r}
ggplot(data =  map_data('state')) + 
  geom_polygon(aes(x=long, y = lat, group=group),
              fill = 'lightblue', color = "black") +
geom_point(data=dataset_profile, #specifying location of lat/long data
             aes(x = `long (dec. deg)`, y = `lat (dec. deg)`), #defines attributes and puts lat/long points on graph
             shape = 'x', color = 'red') +
  coord_fixed(1.3) + 
  theme_nothing() 

ggplot(data =  map_data('state') %>% #selects map data
         filter(region %in% c("tennessee"))) + #filters out the state(s) to be represented 
  geom_polygon(aes(x=long, y = lat, group = group), #defines attributes of state outline/visualization
               fill = 'lightblue', color = 'black') + 
  geom_point(data=dataset_profile, #specifying location of lat/long data
             aes(x = `long (dec. deg)`, y = `lat (dec. deg)`), #defines attributes and puts lat/long points on graph
             shape = 'x', color = 'red') +
  coord_fixed(1.3) + #sets scale for coordinates, making them visible by increasing ratio
  theme_nothing() #sets ggplot theme

```

```{r eval=FALSE}
#this is useful to see for the analysis but we don't want it in the report
dataset_layer %>%
  pivot_longer(cols = intersect(names(.), type_cols$num_cols), values_drop_na = TRUE) %>%
  group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) %>%
  bind_rows(
    dataset_layer %>%
      pivot_longer(cols = intersect(names(.), type_cols$factor_cols), values_drop_na = TRUE) %>%
      group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) ) %>%
  arrange(n) %>%
  knitr::kable()
```

## Profile histograms

```{r}
ggplot(dataset_profile %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
  geom_histogram(aes(x=measurement)) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```

## Depth plots

```{r}
ggplot(dataset_layer %>% 
         pivot_longer(cols=c('layer_top (cm)', 'layer_bot (cm)'),
                       values_to='depth') %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
         geom_line(aes(x=depth, y= measurement, group = profile_name), alpha = 0.5) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```

## Citations

Please see @Froberg2008, https://search.dataone.org/view/doi%3A10.3334%2FCDIAC%2FVRC.002, and 
http://cdiac.ess-dive.lbl.gov/epubs/ndp/ndp078a/ndp078a.html for additional details and if you are using ISCN3 please cite.
