# USGS Harden

The USGS Harden data set in ISCN3 contains 1203 layer-level rows and 186 profile-level information rows after cleaning for ISCN3.5.
Rows weren't removed the dataset_layer table from ISCN3 because of there were no duplicated gap-filled entries.
As it is such a large data set, the lack of duplicates was verified using a length(summary()) function.
4 columns were removed from the dataset_layer table from ISCN3 because these columns contained values from ISCN soc stock correction.
10 rows were removed from the dataset_profile table from ISCN3 because of ISCN soc gap-filling.
5 columns were removed from the dataset_profile table from ISCN3 because these columns contained values from ISCN soc stock correction.
Ggplot maps were adjusted to be centered upon Alaska and narrowed in on location of specific data points.

```{r warning=FALSE, message=FALSE}
datasetName <- "USGS Harden" 

##### Extract the study information ####
dataset_study <- citation_raw %>% 
  filter(dataset_name == datasetName) %>%
  select(where(function(xx){!all(is.na(xx))})) %>%
  full_join(dataset_raw %>% 
              filter(dataset_name == datasetName) %>%
              select(where(function(xx){!all(is.na(xx))})), suffix = c('_citation', '_dataset'))%>%
  standardCast()
```

```{r warning=FALSE, message=FALSE}
##### Extract the profile information ####

#comparison for pre ISCN soc stock correction
dataset_profile_org <- profile_raw  %>%
   filter(dataset_name_sub == datasetName) %>%
   standardCast()

dataset_profile <- profile_raw  %>%
  filter(dataset_name_sub == datasetName) 

if(any(grepl('ISCN', dataset_profile$dataset_name_soc))){
  #reassign rows where the ISCN tried to fill in SOC values
  dataset_profile <- dataset_profile %>%
    group_by(dataset_name_soc) %>%
    mutate(`soc_depth (cm)` = if_else(grepl('ISCN', dataset_name_soc),
                                      rep(NA_character_, length(`soc_depth (cm)`)), `soc_depth (cm)`),
           `soc (g cm-2)` = if_else(grepl('ISCN', dataset_name_soc),
                                    rep(NA_character_, length(`soc (g cm-2)`)), `soc (g cm-2)`),
           soc_carbon_flag = if_else(grepl('ISCN', dataset_name_soc),
                                     rep(NA_character_, length(soc_carbon_flag)), soc_carbon_flag),
           soc_spatial_flag = if_else(grepl('ISCN', dataset_name_soc),
                                      rep(NA_character_, length(soc_spatial_flag)), soc_spatial_flag),
           soc_method = if_else(grepl('ISCN', dataset_name_soc), 
                                rep(NA_character_, length(soc_method)), soc_method)) %>%
    ungroup()
  
}

#remove the soc dataset since we've taken care of the ISCN notation
dataset_profile <- select(dataset_profile, -dataset_name_soc)  

if(any(count(dataset_profile, dataset_name_sub, site_name, profile_name)$n > 1)){
  #if the rows are duplicated then fill in missing values by group
  dataset_profile <- dataset_profile %>%
    group_by(dataset_name_sub, site_name, profile_name) %>%
    mutate_at(vars(-group_cols()), 
              function(xx){ifelse(sum(!is.na(xx)) == 1, rep(xx[!is.na(xx)], length(xx)),xx)}) %>% #if there is one value that isn't na then populate the rest of the entry, this fills in the
    ungroup() %>%
    unique() #collapase rows that are non-unique
}

dataset_profile <- standardCast(dataset_profile)
```

```{r warning = FALSE, message = FALSE}
##### Extract the layer infromation ####

#comparison before SOC correction
dataset_layer_org <- layer_raw %>%
  filter(dataset_name_sub == datasetName) %>%
  standardCast()

dataset_layer <- layer_raw %>%
  filter(dataset_name_sub == datasetName) 

if(any(grepl('ISCN', dataset_layer$dataset_name_soc))){
  #reassign rows where the ISCN tried to fill in SOC values
  dataset_layer <- dataset_layer %>%
    group_by(dataset_name_soc) %>%
    mutate(`soc (g cm-2)` = if_else(grepl('ISCN', dataset_name_soc),
                                    rep(NA_character_, length(`soc (g cm-2)`)), `soc (g cm-2)`),
           soc_carbon_flag = if_else(grepl('ISCN', dataset_name_soc),
                                     rep(NA_character_, length(soc_carbon_flag)), soc_carbon_flag),
           soc_method = if_else(grepl('ISCN', dataset_name_soc), 
                                rep(NA_character_, length(soc_method)), soc_method)) %>%
    ungroup()
  
}

#remove the soc dataset since we've taken care of the ISCN notation
dataset_layer <- select(dataset_layer, -dataset_name_soc) 

if(any(count(dataset_layer, dataset_name_sub, site_name, profile_name, layer_name)$n > 1)){
  #if the rows are duplicated then fill in missing values by group
  dataset_layer <- dataset_layer %>%
    group_by(dataset_name_sub, site_name, profile_name, layer_name) %>%
    mutate_at(vars(-group_cols()), 
              function(xx){ifelse(sum(!is.na(xx)) == 1, rep(xx[!is.na(xx)], length(xx)),xx)}) %>% #if there is one value that isn't na then populate the rest of the entry, this fills in the
    ungroup() %>%
    unique() #collapase rows that are non-unique
}

dataset_layer <- standardCast(dataset_layer)

```

## Running Summaries to Make Sure Only Duplicates Were Removed

```{r}
length(summary(as.factor(as.character(dataset_profile_org$profile_name)), maxsum = 22000))
#To guarantee that only duplicates were removed, the value of unique dataset_profile_org$profile_name should be the same amount of objects as dataset_profile
#4357 = 4357
length(summary(as.factor(as.character(dataset_layer_org$layer_name)), maxsum = 22000))
#To guarantee that only duplicates were removed, the value of unique dataset_layer_org$layer_name should be the same amount of objects as dataset_layer
#21736 != 21739
```


```{r message=FALSE, warning=FALSE}
knitr::kable(t(dataset_study))
```


There are the following factors in the profile:

```{r}
knitr::kable(summary(dataset_profile %>% select_if(is.factor)))
```

And the following factors in the layers:
```{r}
knitr::kable(summary(dataset_layer %>% select_if(is.factor)))
```


## Location

## Use if map is not centered around the specific region of dataset
```{r}

country <- ggplot2::map_data('world2', 'usa')
avgLat <- dataset_layer %>% 
  pull('lat (dec. deg)') %>%
  mean()

avgLong <- dataset_layer %>% 
  pull('long (dec. deg)') %>%
  mean() + 360

ggplot(data =  map_data("world2")) +
  geom_polygon( aes(x = long, y = lat, group = group),
                fill = 'grey', color = "black") +
  geom_polygon(data = country, aes(x = long, y = lat, group = group),
                fill = 'lightblue', color = "black") +
  geom_point(data = dataset_profile, aes(x = avgLong, y = avgLat), shape = 'x', color = 'red', size = 5) +
  coord_cartesian(xlim=c(avgLong - 65, avgLong + 65), ylim = c(avgLat - 25, avgLat + 25)) +
  theme_nothing() +
  labs(title = 'Profile data') #centers first map around Alaska

ggplot(data =  map_data("world2")) + 
  geom_polygon(aes(x=long, y = lat, group = group), 
               fill = 'grey', color = 'black') + 
  geom_polygon(data = country, aes(x = long, y = lat, group = group),
                fill = 'lightblue', color = "black") +
  geom_point(data = dataset_layer, aes(x = `long (dec. deg)` + 360, y = `lat (dec. deg)`),
             shape = 'x', color = 'red', size = 2.5) +
  coord_cartesian(xlim=c(avgLong - 9.1, avgLong + 9.1), ylim = c(avgLat - 3.5, avgLat + 3.5)) +
  theme_nothing() +
  labs(title = 'Layer data') #centers on specific data points in Alaska, providing a more narrow window
  
```

```{r eval=FALSE}
#this is useful to see for the analysis but we don't want it in the report
dataset_layer %>%
  pivot_longer(cols = intersect(names(.), type_cols$num_cols), values_drop_na = TRUE) %>%
  group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) %>%
  bind_rows(
    dataset_layer %>%
      pivot_longer(cols = intersect(names(.), type_cols$factor_cols), values_drop_na = TRUE) %>%
      group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) ) %>%
  arrange(n) %>%
  knitr::kable()
```

## Profile histograms

```{r}
ggplot(dataset_profile %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
  geom_histogram(aes(x=measurement)) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```

## Depth plots

```{r}
ggplot(dataset_layer %>% 
         pivot_longer(cols=c('layer_top (cm)', 'layer_bot (cm)'),
                       values_to='depth') %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
         geom_line(aes(x=depth, y= measurement, group = profile_name), alpha = 0.5) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```

## TODO

- Contains data citations (@Harden2008a @Harden2008b) that need to be revisited in ISCN 4
    Both of the URL's provided lead to 404 error pages.
      [2] URL for dataone of Harden2008a: https://search.dataone.org/view/knb-lter-bnz.336.19
            actual download URL leads to an error: http://www.lter.uaf.edu/404.php
              Contacting of researchers is necessary.
      [4] URL for dataone of Harden2008b: https://search.dataone.org/view/https%3A%2F%2Fpasta.lternet.edu%2Fpackage%2Fmetadata%2Feml%2Fknb-lter-bnz%2F334%2F18
            actual download URL: http://www.lter.uaf.edu/data/data-detail/id/334

## Citations

Please see @Harden2008a, @ODonnell2011, @Manies204, @Harden2008b, @Harden2006 and @Neff2005 for additional details and if you are using ISCN3 please cite.
