# AK DSC Project SOC stock computation

The AK DSC Project SOC Stock computation data set in ISCN3 contains 9002 layer-level rows and 1605 profile-level information rows after cleaning for ISCN3.5.
Rows were removed from ISCN3 from the dataset_layer table because they were generated using ISCN soc stock correction.
Columns were removed from ISCN3 from the dataset_layer table because they were generated using ISCN soc stock correction.
Rows were removed from ISCN3 from the dataset_profile layer table because they were generated using ISCN soc stock correction.
Columns were removed from ISCN3 from the dataset_profile layer table because of duplicated entries and using ISCN soc stock correction.
Ggplot maps were adjusted to focus on the location of Alaska, United States.
Longitude and bulk density values were deemed to be appropriate.
Dataset_layer values were checked and compared to original to ensure proper cleaning.


```{r warning=FALSE, message=FALSE}
datasetName <- "AK DSC Project SOC stock computation"

##### Extract the study information ####
dataset_study <- citation_raw %>% 
  filter(dataset_name == datasetName) %>%
  select(where(function(xx){!(is.na(xx))})) %>%
  full_join(dataset_raw %>% 
              filter(dataset_name == datasetName) %>%
              select(where(function(xx){!(is.na(xx))})), suffix = c('_citation', '_dataset'))%>%
  standardCast()
```

```{r warning=FALSE, message=FALSE}
##### Extract the profile information ####

#comparison for pre ISCN soc stock correction
dataset_profile_org <- profile_raw  %>%
   filter(dataset_name_sub == datasetName) %>%
   standardCast()

dataset_profile <- profile_raw  %>%
  filter(dataset_name_sub == datasetName)

if(any(grepl('ISCN', dataset_profile$dataset_name_soc))){
  #reassign rows where the ISCN tried to fill in SOC values
  dataset_profile <- dataset_profile %>%
    group_by(dataset_name_soc) %>%
    mutate(`soc_depth (cm)` = if_else(grepl('ISCN', dataset_name_soc),
                                      rep(NA_character_, length(`soc_depth (cm)`)), `soc_depth (cm)`),
           `soc (g cm-2)` = if_else(grepl('ISCN', dataset_name_soc),
                                    rep(NA_character_, length(`soc (g cm-2)`)), `soc (g cm-2)`),
           soc_carbon_flag = if_else(grepl('ISCN', dataset_name_soc),
                                     rep(NA_character_, length(soc_carbon_flag)), soc_carbon_flag),
           soc_spatial_flag = if_else(grepl('ISCN', dataset_name_soc),
                                      rep(NA_character_, length(soc_spatial_flag)), soc_spatial_flag),
           soc_method = if_else(grepl('ISCN', dataset_name_soc), 
                                rep(NA_character_, length(soc_method)), soc_method)) %>%
    ungroup()
  
}

#remove the soc dataset since we've taken care of the ISCN notation
dataset_profile <- select(dataset_profile, -dataset_name_soc) 

if(any(count(dataset_profile, dataset_name_sub, site_name, profile_name)$n > 1)){
  #if the rows are duplicated then fill in missing values by group
  dataset_profile <- dataset_profile %>%
    group_by(dataset_name_sub, site_name, profile_name) %>%
    mutate_at(vars(-group_cols()), 
              function(xx){ifelse(sum(!is.na(xx)) == 1, rep(xx[!is.na(xx)], length(xx)),xx)}) %>% #if there is one value that isn't na then populate the rest of the entry, this fills in the
    ungroup() %>%
    unique() #collapase rows that are non-unique
}

dataset_profile <- standardCast(dataset_profile)
```

```{r warning= FALSE, message=FALSE}
##### Extract the layer infromation ####

#comparison before SOC correction
dataset_layer_org <- layer_raw %>%
  filter(dataset_name_sub == datasetName) %>%
  standardCast()

dataset_layer <- layer_raw %>%
  filter(dataset_name_sub == datasetName) 

if(any(grepl('ISCN', dataset_layer$dataset_name_soc))){
  #reassign rows where the ISCN tried to fill in SOC values
  dataset_layer <- dataset_layer %>%
    group_by(dataset_name_soc) %>%
    mutate(`soc (g cm-2)` = if_else(grepl('ISCN', dataset_name_soc),
                                    rep(NA_character_, length(`soc (g cm-2)`)), `soc (g cm-2)`),
           soc_carbon_flag = if_else(grepl('ISCN', dataset_name_soc),
                                     rep(NA_character_, length(soc_carbon_flag)), soc_carbon_flag),
           soc_method = if_else(grepl('ISCN', dataset_name_soc), 
                                rep(NA_character_, length(soc_method)), soc_method)) %>%
    ungroup()
  
}

#remove the soc dataset since we've taken care of the ISCN notation
dataset_layer <- select(dataset_layer, -dataset_name_soc) 

if(any(count(dataset_layer, dataset_name_sub, site_name, profile_name, layer_name)$n > 1)){
  #if the rows are duplicated then fill in missing values by group
  dataset_layer <- dataset_layer %>%
    group_by(dataset_name_sub, site_name, profile_name, layer_name) %>%
    mutate_at(vars(-group_cols()), 
              function(xx){ifelse(sum(!is.na(xx)) == 1, rep(xx[!is.na(xx)], length(xx)),xx)}) %>% #if there is one value that isn't na then populate the rest of the entry, this fills in the
    ungroup() %>%
    unique() #collapase rows that are non-unique
}

dataset_layer <- standardCast(dataset_layer)

```

```{r message=FALSE, warning=FALSE}
knitr::kable(t(dataset_study))
```


There are the following factors in the profile:

```{r}
knitr::kable(summary(dataset_profile %>% select_if(is.factor)))
```

And the following factors in the layers:
```{r}
knitr::kable(summary(dataset_layer %>% select_if(is.factor)))
```

```{r}
###determining if there are extraneous values in column
summary(dataset_layer$`bd_tot (g cm-3)`)#shows max and min values for column
```
```{r}
###determining if there are extraneous values in column
summary(dataset_layer$`long (dec. deg)`)
sum(dataset_layer$`bd_tot (g cm-3)` >0, na.rm = TRUE) #returns number of values greater than acceptable range
```

```{r}
###determining if there are extraneous values in column
summary(dataset_profile$`long (dec. deg)`)
sum(dataset_layer$`bd_tot (g cm-3)` >0, na.rm = TRUE) #returns number of values greater than acceptable range
```

## Location

```{r}

ggplot(data =  map_data("world2")) + 
  geom_polygon(aes(x=long, y = lat, group = group), 
               fill = 'grey', color = 'black') + 
  geom_point(data= dataset_profile, 
             aes(x = `long (dec. deg)` +360, y = `lat (dec. deg)`),
             shape = 'x', color = 'red') +
  coord_cartesian(xlim=c(20,340), ylim=c(-80,80)) +
  theme_nothing() +
  labs(title = 'Profile data')

ggplot(data =  map_data("world2")) + 
  geom_polygon(aes(x=long, y = lat, group = group), 
               fill = 'grey', color = 'black') + 
  geom_point(data= dataset_layer %>% select(`long (dec. deg)`, `lat (dec. deg)`) %>% unique(), 
             aes(x = `long (dec. deg)` +360, y = `lat (dec. deg)`),
             shape = 'x', color = 'red', size = 2) +
  coord_cartesian(xlim=c(150,250), ylim=c(50,75)) + #shifting second map location to center on Alaska
  theme_nothing() +
  labs(title = 'Layer data')

  
```

## Profile histograms

```{r}
ggplot(dataset_profile %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
  geom_histogram(aes(x=measurement)) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```

## Depth plots

```{r}
ggplot(dataset_layer %>% 
         pivot_longer(cols=c('layer_top (cm)', 'layer_bot (cm)'),
                       values_to='depth') %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
         geom_line(aes(x=depth, y= measurement, group = profile_name), alpha = 0.5) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```

## TODO

- download and re-ingest from 10.3334/CDIAC/lue.ndp018

## Citations

Please see @Johnson2011 for additional details and if you are using ISCN3 please cite.
