# USGS_S3C

```{r warning=FALSE, message=FALSE}
datasetName <- "USGS_S3C"

dataset_study <- citation_raw %>% 
  filter(dataset_name == datasetName) %>%
  select(where(function(xx){!(is.na(xx))})) %>%
  full_join(dataset_raw %>% 
              filter(dataset_name == datasetName) %>%
              select(where(function(xx){!(is.na(xx))})), suffix = c('_citation', '_dataset'))%>%
  standardCast()

#comparison for pre ISCN soc stock correction
#dataset_profile_temp <- profile_raw  %>%
#   filter(dataset_name_sub == datasetName) %>%
#   standardCast()

dataset_profile <- profile_raw  %>%
  filter(dataset_name_sub == datasetName) %>%
  select(-dataset_name_soc) %>%
  standardCast() #comment out pipe here and take a look at the profile information to figure out how to remove ISCN soc computations
  #in this case we don't need to do anything

#comparison before SOC correction
#dataset_layer_temp <- layer_raw %>%
#  filter(dataset_name_sub == datasetName) %>%
#  standardCast()

dataset_layer <- layer_raw %>%
  filter(dataset_name_sub == datasetName)  %>%
  group_by(dataset_name_sub, site_name, profile_name, layer_name) %>%
  #mutate_at(vars(-group_cols(), -'dataset_name_soc'), 
  #          function(xx){ifelse(sum(!is.na(xx)) == 1, rep(xx[!is.na(xx)], length(xx)),xx)}) %>% #if there is one value that isn't na then populate the rest of the entry, this fills in the 
  #filter(dataset_name_soc == dataset_name_sub) %>% #only take the soc values provided by the submitter
  select(-starts_with('soc')) %>% # all SOC columns are ISCN calculated regardless of dataset_name_soc apparently
  select(-dataset_name_soc) %>%
  standardCast() #we don't care about the flag any more

```

The `USGS_S3C` data set in ISCN3 contains `r nrow(dataset_layer)` layer-level information and `r nrow(dataset_profile)` profile-level information after cleaning for ISCN3.5.
All layer level soil carbon calculations were ISCN3 calculated according to soc_method

```{r message=FALSE, warning=FALSE}
knitr::kable(t(dataset_study))
```


There are the following factors in the profile:

```{r}
knitr::kable(summary(dataset_profile %>% select_if(is.factor)))
```

And the following factors in the layers:
```{r}
knitr::kable(summary(dataset_layer %>% select_if(is.factor)))
```


## Location

```{r}

ggplot(data =  map_data("usa")) + 
  geom_polygon(aes(x=long, y = lat, group = group), 
               fill = 'grey', color = 'black') + 
  geom_point(data= dataset_profile, 
             aes(x = `long (dec. deg)`, y = `lat (dec. deg)`),
             shape = 'x', color = 'red') +
  coord_fixed(1.3) +
  theme_nothing() +
  labs(title = 'Profile data')

ggplot(data =  map_data("usa")) + 
  geom_polygon(aes(x=long, y = lat, group = group), 
               fill = 'grey', color = 'black') + 
  geom_point(data= dataset_layer %>% select(`long (dec. deg)`, `lat (dec. deg)`) %>% unique(), 
             aes(x = `long (dec. deg)`, y = `lat (dec. deg)`),
             shape = 'x', color = 'red') +
  coord_fixed(1.3) +
  theme_nothing() +
  labs(title = 'Layer data')

```

```{r eval=FALSE}
#this is useful to see for the analysis but we don't want it in the report
dataset_layer %>%
  pivot_longer(cols = intersect(names(.), type_cols$num_cols), values_drop_na = TRUE) %>%
  group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) %>%
  bind_rows(
    dataset_layer %>%
      pivot_longer(cols = intersect(names(.), type_cols$factor_cols), values_drop_na = TRUE) %>%
      group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) ) %>%
  arrange(n) %>%
  knitr::kable()
```


## Depth plots

```{r}
ggplot(dataset_layer %>% 
         pivot_longer(cols=c('layer_top (cm)', 'layer_bot (cm)'),
                       values_to='depth') %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
         geom_line(aes(x=depth, y= measurement, group = profile_name), alpha = 0.5) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```

## TODO

- download and re-ingest from @Buell2004
- Double check profile soc calculations, do we just remove all the columns?
- Clean up pH values, check clay-sand-silt percent, clean up bulk density total, check bounds for caco3 percent

## Citations

Please see @Buell2004 for additional details and if you are using ISCN3 please cite.
