# Worldwide soil carbon and nitrogen data

```{r warning=FALSE, message=FALSE}
datasetName <- "Worldwide soil carbon and nitrogen data"

dataset_study <- citation_raw %>% 
  filter(dataset_name == datasetName) %>%
  select(where(function(xx){!(is.na(xx))})) %>%
  full_join(dataset_raw %>% 
              filter(dataset_name == datasetName) %>%
              select(where(function(xx){!(is.na(xx))})), suffix = c('_citation', '_dataset'))%>%
  standardCast()

#comparison for pre ISCN soc stock correction
# dataset_profile_temp <- profile_raw  %>%
#   filter(dataset_name_sub == datasetName) %>%
#   standardCast()

dataset_profile <- profile_raw  %>%
  filter(dataset_name_sub == datasetName) %>%
  standardCast() %>% #comment out pipe here and take a look at the profile information to figure out how to remove ISCN soc computations
  filter(dataset_name_soc == datasetName) %>% #in this case just remove the whole row, no other data is lost
  select(where(function(xx){!all(is.na(xx))})) %>% #remove the extra soc column lingering from ISCN soc computation
  select(-dataset_name_soc)

#comparison before SOC correction
# dataset_layer_temp <- layer_raw %>%
#   filter(dataset_name_sub == datasetName) %>%
#   standardCast()

dataset_layer <- layer_raw %>%
  filter(dataset_name_sub == datasetName) %>%
  standardCast() %>%
  select(-starts_with('soc')) %>% #all dataset_name_soc values are ISCN soc stock computation flags
  select(-dataset_name_soc) %>% #we don't care about the flag any more
  select(where(function(xx){!all(is.na(xx))}))

```

The `Worldwide soil carbon and nitrogen data` data set in ISCN3 contains `r nrow(dataset_layer)` layer-level information and `r nrow(dataset_profile)` profile-level information after cleaning for ISCN3.5.
ISCN3 all of the soil carbon stocks where bulk density and organic carbon were calculated by ISCN and thus these columns were removed for ISCN3.5.

```{r message=FALSE, warning=FALSE}
knitr::kable(t(dataset_study))
```


There are the following factors in the profile:

```{r}
knitr::kable(summary(dataset_profile %>% select_if(is.factor)))
```

And the following factors in the layers:
```{r}
knitr::kable(summary(dataset_layer %>% select_if(is.factor)))
```


## Location

```{r}

ggplot(data =  map_data("world")) + 
  geom_polygon(aes(x=long, y = lat, group = group), 
               fill = 'grey', color = 'black') + 
  geom_point(data= dataset_profile, 
             aes(x = `long (dec. deg)`, y = `lat (dec. deg)`),
             shape = 'x', color = 'red') +
  coord_fixed(1.3) +
  theme_nothing() +
  labs(title = 'Profile data')

ggplot(data =  map_data("world")) + 
  geom_polygon(aes(x=long, y = lat, group = group), 
               fill = 'grey', color = 'black') + 
  geom_point(data= dataset_layer %>% select(`long (dec. deg)`, `lat (dec. deg)`) %>% unique(), 
             aes(x = `long (dec. deg)`, y = `lat (dec. deg)`),
             shape = 'x', color = 'red') +
  coord_fixed(1.3) +
  theme_nothing() +
  labs(title = 'Layer data')

```

```{r eval=FALSE}
#this is useful to see for the analysis but we don't want it in the report
dataset_layer %>%
  pivot_longer(cols = intersect(names(.), type_cols$num_cols), values_drop_na = TRUE) %>%
  group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) %>%
  bind_rows(
    dataset_layer %>%
      pivot_longer(cols = intersect(names(.), type_cols$factor_cols), values_drop_na = TRUE) %>%
      group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) ) %>%
  arrange(n) %>%
  knitr::kable()
```

## Profile histograms

```{r}
ggplot(dataset_profile %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
  geom_histogram(aes(x=measurement)) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```

## Depth plots

```{r}
ggplot(dataset_layer %>% 
         pivot_longer(cols=c('layer_top (cm)', 'layer_bot (cm)'),
                       values_to='depth') %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
         geom_line(aes(x=depth, y= measurement, group = profile_name), alpha = 0.5) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```

## TODO

- download and re-ingest from 10.3334/CDIAC/lue.ndp018

## Citations

Please see @Zinke1986 for additional details and if you are using ISCN3 please cite.
