---
title: "ISCN 3_1"
author: "Dr. Katherine Todd-Brown, Julie Bielecki, Anne Davin, Rita Hippe, Urmi Thorat"
date: "5/25/2021"
output: html_document
---

# Introduction

International Soil Carbon Network 3 (ISCN3) consists of serveral data sets. This R Markdown cleans and reorganizes ISCN3 datasets into a study level table, a profile level table, and a layer level table using the ISCN3_1 function. Additionally, maps for each data set are produced within this R Markdown.

The `study` table produced by the ISCN3_1 function describe information about each data contribution including any citations practices and contact information for the contributor. The `profile` table produced by the ISCN3_1 function describes core summary information that is 2D resolved.In contrast, the `layer` table produced the ISCN3_1 function holds 3D resolved information.


```{r warning=FALSE, message=FALSE}
library(data.table)
library(SOCDRaH2)
library(tidyverse)
library(lubridate)
library(tibble)
library(ggmap)
library(maps)
library(mapdata)
library(knitr)
library(tidyr)

data_dir <- '../ISCN3'

ISCN3 <- ISCN3_1(data_dir)

```


```{r message=FALSE, warning=FALSE}
knitr::kable(t(dataset_study))
```


There are the following factors in the profile:

```{r}
knitr::kable(summary(dataset_profile %>% select_if(is.factor)))
```

And the following factors in the layers:
```{r}
knitr::kable(summary(dataset_layer %>% select_if(is.factor)))
```


## Location
```{r}

ggplot(data =  map_data("world")) + 
  geom_polygon(aes(x=long, y = lat, group = group), 
               fill = 'grey', color = 'black') + 
  geom_point(data= dataset_profile, 
             aes(x = `long (dec. deg)`, y = `lat (dec. deg)`),
             shape = 'x', color = 'red') +
  coord_fixed(1.3) +
  theme_nothing() +
  labs(title = 'Profile data') #+
  #coord_map("ortho", orientation = c(90, 0, 0))

ggplot(data =  map_data("world")) + 
  geom_polygon(aes(x=long, y = lat, group = group), 
               fill = 'grey', color = 'black') + 
  geom_point(data= dataset_layer %>% select(`long (dec. deg)`, `lat (dec. deg)`) %>% unique(), 
             aes(x = `long (dec. deg)`, y = `lat (dec. deg)`),
             shape = 'x', color = 'red') +
  coord_fixed(1.3) +
  theme_nothing() +
  labs(title = 'Layer data')

```


## Use if map is not centered around the specific region of dataset
```{r}

#country <- ggplot2::map_data('world2', 'usa')
#avgLat <- dataset_layer %>% 
#  pull('lat (dec. deg)') %>%
#  mean()

#avgLong <- dataset_layer %>% 
#  pull('long (dec. deg)') %>%
#  mean() + 360

#ggplot(data =  map_data("world2")) +
#  geom_polygon( aes(x = long, y = lat, group = group),
#                fill = 'grey', color = "black") +
#  geom_polygon(data = country, aes(x = long, y = lat, group = group),
#                fill = 'lightblue', color = "black") +
#  geom_point(data = dataset_profile, aes(x = avgLong, y = avgLat), shape = 'x', color = 'red', size = 5) +
#  coord_cartesian(xlim=c(avgLong - 65, avgLong + 65), ylim = c(avgLat - 25, avgLat + 25)) +
#  theme_nothing() +
#  labs(title = 'Profile data')

#ggplot(data =  map_data("world2")) + 
#  geom_polygon(aes(x=long, y = lat, group = group), 
#               fill = 'grey', color = 'black') + 
#  geom_polygon(data = country, aes(x = long, y = lat, group = group),
#                fill = 'lightblue', color = "black") +
#  geom_point(data = dataset_layer, aes(x = `long (dec. deg)` + 360, y = `lat (dec. deg)`),
#             shape = 'x', color = 'red', size = 2.5) +
#  coord_cartesian(xlim=c(avgLong - 9.1, avgLong + 9.1), ylim = c(avgLat - 3.5, avgLat + 3.5)) +
#  theme_nothing() +
#  labs(title = 'Layer data')
  
```

```{r eval=FALSE}
#this is useful to see for the analysis but we don't want it in the report
dataset_layer %>%
  pivot_longer(cols = intersect(names(.), type_cols$num_cols), values_drop_na = TRUE) %>%
  group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) %>%
  bind_rows(
    dataset_layer %>%
      pivot_longer(cols = intersect(names(.), type_cols$factor_cols), values_drop_na = TRUE) %>%
      group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) ) %>%
  arrange(n) %>%
  knitr::kable()
```

## Profile histograms

```{r}
ggplot(dataset_profile %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
  geom_histogram(aes(x=measurement)) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```

## Depth plots

```{r}
ggplot(dataset_layer %>% 
         pivot_longer(cols=c('layer_top (cm)', 'layer_bot (cm)'),
                       values_to='depth') %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
         geom_line(aes(x=depth, y= measurement, group = profile_name), alpha = 0.5) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```


## Individualized tables, maps, histograms, and depth plots by dataset



## Plot-generating function 
```{r}

# TODO Implement special case code
# TODO Average coordinates for base map code
# TODO if/else call for US

plotGenerate <- function(ISCN3, datasetName){

#### TABLES ####  
   studyTable <- knitr::kable(t(ISCN3$study %>%
                        filter(dataset_name == datasetName) %>%
                        dplyr::select(where(function(xx){!all(is.na(xx))}))),
                              col.names = "",
                              caption = paste("A summary of [", datasetName,
                                        "] study and contact information.", sep = ""))
   
   profileTable <- knitr::kable(summary(ISCN3$profile %>%
                        filter(dataset_name_sub == datasetName) %>%
                        dplyr::select(where(function(xx){!all(is.na(xx))})) %>%
                        select_if(is.factor) %>%
                        mutate_all(droplevels)),
                        caption = paste("A summary of [", datasetName,
                                        "] profile data,  \nwhich is 2D resolved information.", sep = ""))
   
      
   layerTable <- knitr::kable(summary(ISCN3$layer %>%
                        filter(dataset_name_sub == datasetName) %>%
                        dplyr::select(where(function(xx){!all(is.na(xx))})) %>%
                        select_if(is.factor) %>%
                        mutate_all(droplevels)),
                        caption = paste("A summary of [", datasetName,
                                        "] layer data,  \ncontaining 3D resolved information.", sep = ""))  
 #### MAP CODE ####  
  country <- ggplot2::map_data('world2', 'usa')
  

  baseLat <- ISCN3$profile %>%
                  filter(dataset_name_sub == datasetName) %>%
                  dplyr::pull('lat (dec. deg)')
  
  baseLong <- ISCN3$profile %>%
                  filter(dataset_name_sub == datasetName) %>%
                  dplyr::pull('long (dec. deg)')
   
  rangeLat <- if (diff(range(baseLong, na.rm = TRUE)) <= 5.0){
                  diff(range(baseLong, na.rm = TRUE)) * 25
              } else {
                  diff(range(baseLong, na.rm = TRUE)) * 30
              }
                    
  calcLat <- rangeLat * 2.6
  
   profileMap <- ggplot(data =  map_data("world2")) + 
                            geom_polygon(aes(x=long, y = lat, group = group), 
                                         fill = 'grey', color = 'black') + 
                            geom_polygon(data = country, aes(x = long, y = lat, group = group),
                                         fill = 'lightblue', color = 'black') +
                            geom_point(data= ISCN3$profile %>%
                                filter(dataset_name_sub == datasetName),
                                        aes(x = `long (dec. deg)` + 360,
                                            y = `lat (dec. deg)`),
                                        shape = 'x', color = 'red') +
                          coord_cartesian(xlim = c(mean(baseLong, na.rm = TRUE) + 360 - calcLat,
                                                   mean(baseLong, na.rm = TRUE) + 360 + calcLat),
                                          ylim = c(mean(baseLat, na.rm = TRUE) - rangeLat,
                                                   mean(baseLat, na.rm = TRUE) + rangeLat)) +
                          theme_grey() +
                          xlab("Longitude") + ylab("Latitude") +
                          labs(title = 'Profile Data Map',
                               caption = paste("A map of [", datasetName,
                                        "] sampling sites,  \nusing coordinates found within the profile data.", sep = ""))
                            #coord_map("ortho", orientation = c(90, 0, 0))
   
  baseLat <- ISCN3$layer %>%
                  filter(dataset_name_sub == datasetName) %>%
                  dplyr::pull('lat (dec. deg)')
  
  baseLong <- ISCN3$layer %>%
                  filter(dataset_name_sub == datasetName) %>%
                  dplyr::pull('long (dec. deg)')
   
  rangeLat <- if (diff(range(baseLong, na.rm = TRUE)) <= 5.0){
                  diff(range(baseLong, na.rm = TRUE)) * 13.5
              } else {
                  diff(range(baseLong, na.rm = TRUE)) * 16
              }
  
  calcLat <- (rangeLat * 2.6)
  
   layerMap <- ggplot(data =  map_data('world2')) + 
                           geom_polygon(aes(x=long, y = lat, group = group), 
                                        color = 'black', fill = 'grey') +
                           geom_polygon(data = ggplot2::map_data('world2', 'usa'),
                                        aes(x = long, y = lat, group = group),
                                        fill = 'lightblue', color = 'black') +
                           geom_point(data= ISCN3$layer %>%
                                          filter(dataset_name_sub == datasetName) %>%
                                          dplyr::select(where(function(xx){!all(is.na(xx))})) %>% 
                                          select(`long (dec. deg)`, `lat (dec. deg)`) %>%
                                          unique(), 
                                        aes(x = `long (dec. deg)` + 360, #shift y-coordinate 360 degrees into the correct orientation
                                            y = `lat (dec. deg)`),
                                        shape = 'x', color = 'red') +
                          coord_cartesian(xlim = c(mean(baseLong, na.rm = TRUE) + 360 - calcLat,
                                                   mean(baseLong, na.rm = TRUE) + 360 + calcLat),
                                          ylim = c(mean(baseLat, na.rm = TRUE) - rangeLat,
                                                   mean(baseLat, na.rm = TRUE) + rangeLat)) +
                          theme_grey() +
                          xlab("Longitude") + ylab("Latitude") +
                          labs(title = 'Layer Data Map',
                               caption = paste("A map of [", datasetName,
                                        "] sampling sites,  \nusing coordinates found within the layer data.",
                                         sep = ""))
#### PLOTS ####   
   histograms <- ggplot(ISCN3$profile %>%
                        filter(dataset_name_sub == datasetName) %>%
                        dplyr::select(where(function(xx){!all(is.na(xx))})) %>%
                              pivot_longer(cols = intersect(names(.), ISCN3$type_columns$num_cols), 
                                  values_to = 'measurement', names_to = 'type')) +
                              geom_histogram(aes(x=measurement)) +
                              facet_wrap(~type, scales='free') +
                          theme_grey() +
                          labs(title = 'Profile Value Histograms',
                               caption = paste("Histograms detailing measurements in [", datasetName,
                                        "] profile data.", sep = ""))
   
   depthPlots <- ggplot(ISCN3$layer %>%
                        filter(dataset_name_sub == datasetName) %>%
                        dplyr::select(where(function(xx){!all(is.na(xx))})) %>% 
                              pivot_longer(cols=c('layer_top (cm)', 'layer_bot (cm)'),
                                      values_to='depth') %>%
                              pivot_longer(cols = intersect(names(.), ISCN3$type_columns$num_cols), 
                                      values_to = 'measurement', names_to = 'type')) +
                              geom_line(aes(x=depth, y= measurement, group = profile_name), alpha = 0.5) +
                          facet_wrap(~type, scales='free') +
                          theme_grey() +
                          labs(title = 'Layer Data Depth Plots',
                               caption = paste("Plots of [", datasetName,
                                        "] layer data by depth.", sep = ""))
  
#### RETURN ####   
return(list(summaryStudyTable = studyTable,
            profileTable = profileTable,
            layerTable = layerTable,
            profileMap = profileMap,
            layerMap = layerMap,
            profileValueHistograms = histograms,
            depthValuePlots = depthPlots))  
}

# HeckmanVisuals <- plotGenerate(ISCN3, datasetName = 'Heckman/Swanston Biscuit Burn')
# 
# print(HeckmanVisuals)

```

## Heckman/Swanston Biscuit Burn
```{r}

print(plotGenerate(ISCN3, datasetName = 'Heckman/Swanston Biscuit Burn'))

```

## Lu_PIMA
```{r}

```

## Heckman lithosequence
```{r}

```

## Oak Ridge National Lab_Loblolly_DWJ
```{r}

```

## Lu_LTER
```{r}

```

## Lehmann Soil C&BC #1
```{r}

```

## Schuur
```{r}

```

## Lehmann NE US soils
```{r}

```

## Vogel
```{r}

```

## Myers-Smith
```{r}

```

## USGS Muhs
```{r}

```

## USGS Harden Yazoo
```{r}

```

## Boby_Mack
```{r}

```

## Kane
```{r}

```

## Bonanza LTER
```{r}

```

## Jorgensen_YKDE
```{r}

```

## UMBS_FASET
```{r}

```

## Jorgensen_ARCN
```{r}

```

## Oak Ridge National Lab_TDE
```{r}

```

## USGS Harden
```{r}

```

## USDA-FS NRS Landscape Carbon Inventory
```{r}

```

## Bockheim
```{r}

```

## Jorgensen_NPS
```{r}

```

## Permafrost_RCN
```{r}

```

## Northern Circumpolar Soil Carbon Database (NCSCD)
```{r}

```

## USGS_S3C
```{r}

```

## AK DSC Project SOC stock computation
```{r}

```

## Worldwide soil carbon and nitrogen data
```{r}

```

## NRCS Sept/2014
```{r}

```


## TODO

- Rename header and file name to reflect dataset_name
- Pull citation and add it to the bib file, check that citation matches the acknowledgement for study
- Check that the clean profile is correct given the orginal information
- Check that the clean layer is correct given the orginal information
- Check that maps are reasonable for the location
- Summerize what the modifications were for this particular dataset
- Create an issue on github with the template below
- Commit and link the commit to the issue

#### Github issue template

- [ ]  Evaluate dataset
- [ ]  Complete template checklist
- [ ]  Write up summary report narrative
- [ ]  Good commenting
- [ ]  External review by someone else

## Citations

Please see [bibtex citation here] for additional details and if you are using ISCN3 please cite.