---
title: "ISCN 3_1"
author: "Dr. Katherine Todd-Brown, Julie Bielecki, Anne Davin, Rita Hippe, Urmi Thorat"
date: "5/25/2021"
output: html_document
---

# Introduction

International Soil Carbon Network 3 (ISCN3) consists of serveral data sets. This R Markdown cleans and reorganizes ISCN3 datasets into a study level table, a profile level table, and a layer level table using the ISCN3_1 function. Additionally, maps for each dataset are produced within this R Markdown.

The `study` tables describe information about each data contribution including any citations practices and contact information for the contributor.
The `profile` table describes core summary information that is 2D resolved.
The `layer` table, in contrasts, holds 3D resolved information.


```{r warning=FALSE, message=FALSE}
library(data.table)
#library(SOCDRaH2)
library(tidyverse)
library(lubridate)
library(tibble)
library(ggmap)
library(maps)
library(mapdata)
library(knitr)
library(tidyr)

data_dir <- '~/Documents/Todd-Brown Lab/ISCN3'

ISCN3 <- ISCN3_1(data_dir)

```


```{r message=FALSE, warning=FALSE}
knitr::kable(t(dataset_study))
```


There are the following factors in the profile:

```{r}
knitr::kable(summary(dataset_profile %>% select_if(is.factor)))
```

And the following factors in the layers:
```{r}
knitr::kable(summary(dataset_layer %>% select_if(is.factor)))
```


## Location

```{r}

ggplot(data =  map_data("world")) + 
  geom_polygon(aes(x=long, y = lat, group = group), 
               fill = 'grey', color = 'black') + 
  geom_point(data= dataset_profile, 
             aes(x = `long (dec. deg)`, y = `lat (dec. deg)`),
             shape = 'x', color = 'red') +
  coord_fixed(1.3) +
  theme_nothing() +
  labs(title = 'Profile data') #+
  #coord_map("ortho", orientation = c(90, 0, 0))

ggplot(data =  map_data("world")) + 
  geom_polygon(aes(x=long, y = lat, group = group), 
               fill = 'grey', color = 'black') + 
  geom_point(data= dataset_layer %>% select(`long (dec. deg)`, `lat (dec. deg)`) %>% unique(), 
             aes(x = `long (dec. deg)`, y = `lat (dec. deg)`),
             shape = 'x', color = 'red') +
  coord_fixed(1.3) +
  theme_nothing() +
  labs(title = 'Layer data')

```


## Use if map is not centered around the specific region of dataset
```{r}

#country <- ggplot2::map_data('world2', 'usa')
#avgLat <- dataset_layer %>% 
#  pull('lat (dec. deg)') %>%
#  mean()

#avgLong <- dataset_layer %>% 
#  pull('long (dec. deg)') %>%
#  mean() + 360

#ggplot(data =  map_data("world2")) +
#  geom_polygon( aes(x = long, y = lat, group = group),
#                fill = 'grey', color = "black") +
#  geom_polygon(data = country, aes(x = long, y = lat, group = group),
#                fill = 'lightblue', color = "black") +
#  geom_point(data = dataset_profile, aes(x = avgLong, y = avgLat), shape = 'x', color = 'red', size = 5) +
#  coord_cartesian(xlim=c(avgLong - 65, avgLong + 65), ylim = c(avgLat - 25, avgLat + 25)) +
#  theme_nothing() +
#  labs(title = 'Profile data')

#ggplot(data =  map_data("world2")) + 
#  geom_polygon(aes(x=long, y = lat, group = group), 
#               fill = 'grey', color = 'black') + 
#  geom_polygon(data = country, aes(x = long, y = lat, group = group),
#                fill = 'lightblue', color = "black") +
#  geom_point(data = dataset_layer, aes(x = `long (dec. deg)` + 360, y = `lat (dec. deg)`),
#             shape = 'x', color = 'red', size = 2.5) +
#  coord_cartesian(xlim=c(avgLong - 9.1, avgLong + 9.1), ylim = c(avgLat - 3.5, avgLat + 3.5)) +
#  theme_nothing() +
#  labs(title = 'Layer data')
  
```

```{r eval=FALSE}
#this is useful to see for the analysis but we don't want it in the report
dataset_layer %>%
  pivot_longer(cols = intersect(names(.), type_cols$num_cols), values_drop_na = TRUE) %>%
  group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) %>%
  bind_rows(
    dataset_layer %>%
      pivot_longer(cols = intersect(names(.), type_cols$factor_cols), values_drop_na = TRUE) %>%
      group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) ) %>%
  arrange(n) %>%
  knitr::kable()
```

## Profile histograms

```{r}
ggplot(dataset_profile %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
  geom_histogram(aes(x=measurement)) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```

## Depth plots

```{r}
ggplot(dataset_layer %>% 
         pivot_longer(cols=c('layer_top (cm)', 'layer_bot (cm)'),
                       values_to='depth') %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
         geom_line(aes(x=depth, y= measurement, group = profile_name), alpha = 0.5) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```

## TODO

- Rename header and file name to reflect dataset_name
- Pull citation and add it to the bib file, check that citation matches the acknowledgement for study
- Check that the clean profile is correct given the orginal information
- Check that the clean layer is correct given the orginal information
- Check that maps are reasonable for the location
- Summerize what the modifications were for this particular dataset
- Create an issue on github with the template below
- Commit and link the commit to the issue

#### Github issue template

- [ ]  Evaluate dataset
- [ ]  Complete template checklist
- [ ]  Write up summary report narrative
- [ ]  Good commenting
- [ ]  External review by someone else

## Citations

Please see [bibtex citation here] for additional details and if you are using ISCN3 please cite.
