---
title: "ISCN 3_1"
author: "Dr. Katherine Todd-Brown, Julie Bielecki, Anne Davin, Rita Hippe, Urmi Thorat"
date: "5/25/2021"
output: html_document
---

# Introduction

International Soil Carbon Network 3 (ISCN3) consists of serveral data sets. This R Markdown cleans and reorganizes ISCN3 datasets into a study level table, a profile level table, and a layer level table using the ISCN3_1 function. Additionally, maps for each data set are produced within this R Markdown.

The `study` table produced by the ISCN3_1 function describe information about each data contribution including any citations practices and contact information for the contributor. The `profile` table produced by the ISCN3_1 function describes core summary information that is 2D resolved.In contrast, the `layer` table produced the ISCN3_1 function holds 3D resolved information.


```{r warning=FALSE, message=FALSE}
library(data.table)
library(SOCDRaH2)
library(tidyverse)
library(lubridate)
library(tibble)
library(ggmap)
library(maps)
library(mapdata)
library(knitr)
library(tidyr)

data_dir <- '../ISCN3'

ISCN3 <- ISCN3_1(data_dir)
```


```{r message=FALSE, warning=FALSE}
knitr::kable(t(dataset_study))
```


There are the following factors in the profile:

```{r}
knitr::kable(summary(dataset_profile %>% select_if(is.factor)))
```

And the following factors in the layers:
```{r}
knitr::kable(summary(dataset_layer %>% select_if(is.factor)))
```


## Location
```{r}
ggplot(data =  map_data("world")) + 
  geom_polygon(aes(x=long, y = lat, group = group), 
               fill = 'grey', color = 'black') + 
  geom_point(data= dataset_profile, 
             aes(x = `long (dec. deg)`, y = `lat (dec. deg)`),
             shape = 'x', color = 'red') +
  coord_fixed(1.3) +
  theme_nothing() +
  labs(title = 'Profile data') #+
  #coord_map("ortho", orientation = c(90, 0, 0))
ggplot(data =  map_data("world")) + 
  geom_polygon(aes(x=long, y = lat, group = group), 
               fill = 'grey', color = 'black') + 
  geom_point(data= dataset_layer %>% select(`long (dec. deg)`, `lat (dec. deg)`) %>% unique(), 
             aes(x = `long (dec. deg)`, y = `lat (dec. deg)`),
             shape = 'x', color = 'red') +
  coord_fixed(1.3) +
  theme_nothing() +
  labs(title = 'Layer data')
```


## Use if map is not centered around the specific region of dataset
```{r}
#country <- ggplot2::map_data('world2', 'usa')
#avgLat <- dataset_layer %>% 
#  pull('lat (dec. deg)') %>%
#  mean()
#avgLong <- dataset_layer %>% 
#  pull('long (dec. deg)') %>%
#  mean() + 360
#ggplot(data =  map_data("world2")) +
#  geom_polygon( aes(x = long, y = lat, group = group),
#                fill = 'grey', color = "black") +
#  geom_polygon(data = country, aes(x = long, y = lat, group = group),
#                fill = 'lightblue', color = "black") +
#  geom_point(data = dataset_profile, aes(x = avgLong, y = avgLat), shape = 'x', color = 'red', size = 5) +
#  coord_cartesian(xlim=c(avgLong - 65, avgLong + 65), ylim = c(avgLat - 25, avgLat + 25)) +
#  theme_nothing() +
#  labs(title = 'Profile data')
#ggplot(data =  map_data("world2")) + 
#  geom_polygon(aes(x=long, y = lat, group = group), 
#               fill = 'grey', color = 'black') + 
#  geom_polygon(data = country, aes(x = long, y = lat, group = group),
#                fill = 'lightblue', color = "black") +
#  geom_point(data = dataset_layer, aes(x = `long (dec. deg)` + 360, y = `lat (dec. deg)`),
#             shape = 'x', color = 'red', size = 2.5) +
#  coord_cartesian(xlim=c(avgLong - 9.1, avgLong + 9.1), ylim = c(avgLat - 3.5, avgLat + 3.5)) +
#  theme_nothing() +
#  labs(title = 'Layer data')
  
```

```{r eval=FALSE}
#this is useful to see for the analysis but we don't want it in the report
dataset_layer %>%
  pivot_longer(cols = intersect(names(.), type_cols$num_cols), values_drop_na = TRUE) %>%
  group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) %>%
  bind_rows(
    dataset_layer %>%
      pivot_longer(cols = intersect(names(.), type_cols$factor_cols), values_drop_na = TRUE) %>%
      group_by(name) %>% summarize(n = length(value), unique_n = length(unique(value))) ) %>%
  arrange(n) %>%
  knitr::kable()
```

## Profile histograms

```{r}
ggplot(dataset_profile %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
  geom_histogram(aes(x=measurement)) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```

## Depth plots

```{r}
ggplot(dataset_layer %>% 
         pivot_longer(cols=c('layer_top (cm)', 'layer_bot (cm)'),
                       values_to='depth') %>%
         pivot_longer(cols = intersect(names(.), type_cols$num_cols), 
                      values_to = 'measurement', names_to = 'type')) +
         geom_line(aes(x=depth, y= measurement, group = profile_name), alpha = 0.5) +
  facet_wrap(~type, scales='free') +
  theme_bw()
```


## Individualized tables, maps, histograms, and depth plots by dataset



## Plot-generating function 
```{r}
# TODO See why layer map is printing the same as profile map (probably need to clear environment and run again)

plotGenerate <- function(ISCN3, datasetName){
#### VARIABLE INITIALIZATION ####
# dataset tables and visualizations
  datasetStudy <- ISCN3$study %>%
    filter(dataset_name == datasetName)
  datasetProfile <- ISCN3$profile %>%
    filter(dataset_name_sub == datasetName)
  datasetLayer <- ISCN3$layer %>%
    filter(dataset_name_sub == datasetName)
  options(knitr.kable.NA = '')
  profileTable <- paste("[", datasetName, "] does not contain profile data.", sep = "")
  profileMap <- NA
  histograms <- NA
# auto-scaling variables 
  baseLat <- datasetProfile %>%
    dplyr::pull('lat (dec. deg)')
  baseLong <- datasetProfile %>%
    dplyr::pull('long (dec. deg)')
  if(diff(range(baseLong, na.rm = TRUE)) <= 10){
    rangeLat <- diff(range(baseLong, na.rm = TRUE)) * 12
  } else {rangeLat <- diff(range(baseLong, na.rm = TRUE)) * 8
  }
  calcLat <- rangeLat * 2.6
  adjustedCoords <- coord_cartesian(xlim = c(mean(baseLong, na.rm = TRUE) + 360 - calcLat,
                                             mean(baseLong, na.rm = TRUE) + 360 + calcLat),
                                    ylim = c(mean(baseLat, na.rm = TRUE) - rangeLat,
                                             mean(baseLat, na.rm = TRUE) + rangeLat))
# general mapping variables
shift360 <- 0 #can shift y-coordinate 360 degrees into the correct orientation if needed
alaskaCoords <- list(coord_cartesian(xlim = c(184.6,231.4),
                           ylim = c(53,71)))
datasestState <- tolower(unique(datasetLayer$`state (state_province)`))   #use layer info to set state, and highlight state of interest on US map
mapStateData <- map_data('state') %>% #pull out relevant info, found in state column
                      filter(region %in% datasestState)
ggMapBase <- list(geom_polygon(aes(x=long, y = lat, group = group), 
                  fill = 'grey', color = 'black'),
                  coord_fixed(1.3),
                  xlab("Longitude"), ylab("Latitude"))
ggUSAFillBlue <- list(geom_polygon(data = ggplot2::map_data('world2', 'usa'),
               aes(x = long, y = lat, group = group),
               fill = 'lightblue', color = 'black'))
ggStateFillBlue <- geom_polygon(data = map_data('state') %>% #pull out relevant info, found in state column
                      filter(region %in% datasestState),
                    aes(x=long, y = lat, group = group),
                    fill = 'light blue', color = 'black')
mapTitleProfile <- labs(title = 'Profile Data Map',
                   caption = paste("A map of [", datasetName, "] sampling sites,  \nusing coordinates found within the profile data.", sep = ""))
mapTitleLayer <- labs(title = 'Layer Data Map',
                 caption = paste("A map of [", datasetName, "] sampling sites,  \nusing coordinates found within the layer data.", sep = ""))

#### TABLES ####  
   studyTable <- knitr::kable(t(datasetStudy %>%
                        dplyr::select(where(function(xx){!all(is.na(xx))}))),
                              col.names = "",
                              caption = paste("A summary of [", datasetName,
                                        "] study and contact information.", sep = ""))

  if(datasetName %in% ISCN3$profile$dataset_name_sub){
     profileTable <- knitr::kable(summary(datasetProfile %>%
                             dplyr::select(where(function(xx){!all(is.na(xx))})) %>%
                             select_if(is.factor) %>%
                             mutate_all(droplevels)),
                  caption = paste("A summary of [", datasetName,
                                  "] profile data,  \nwhich is 2D resolved information.", sep = ""))
  } else {}

   layerTable <- knitr::kable(summary(datasetLayer %>%
                        dplyr::select(where(function(xx){!all(is.na(xx))})) %>%
                        select_if(is.factor) %>%
                        mutate_all(droplevels)),
                        caption = paste("A summary of [", datasetName,
                                        "] layer data,  \ncontaining 3D resolved information.", sep = ""))
   
#### PROFILE MAP ####
   datasetInfo <- datasetProfile
   ggRedX <- list(geom_point(data= datasetInfo, 
             aes(x = `long (dec. deg)` + shift360, y = `lat (dec. deg)`),
             shape = 'x', color = 'red', size = 3))
   if(datasetName %in% ISCN3$profile$dataset_name_sub){    #checking if dataset has profile data
     if(nrow(unique(datasetProfile %>%  select(`country (country)`))) > 1 ){    #checking if dataset has data in more than one country
       profileMap <- ggplot(data =  map_data("world")) +   #using world map
         ggMapBase + ggRedX + mapTitleProfile
     } else if(unique(datasetLayer %>% select(`country (country)`)) ==  "United States"){   #checking if in USA, then in Alaska or not
       if(unique(datasetLayer %>% select(`state (state_province)`)) == "Alaska"){
         shift360 <- 360
         profileMap <- ggplot(data =  map_data("world2")) +  # Alaska-specific mapping
           ggMapBase + ggUSAFillBlue + ggRedX + alaskaCoords +
           mapTitleProfile
       } else { profileMap <- ggplot(data =  map_data('state')) + #any other US state mapping 
         ggMapBase + ggStateFillBlue + ggRedX + mapTitleProfile  
       }
     } else { shift360 <-  360 #use auto-scaling map code (for datasets in one country that is not the US)  
     profileMap <- ggplot(data =  map_data("world2")) + 
       ggMapBase + ggRedX + adjustedCoords + mapTitleProfile   #coord_map("ortho", orientation = c(90, 0, 0))
     }
   } else if(unique(datasetLayer %>% select(`country (country)`)) ==  "United States" &
             unique(datasetLayer %>% select(`state (state_province)`)) != "Alaska"){  
     profileMap <- ggplot(data =  map_data('state')) +
       ggMapBase + ggStateFillBlue +
       labs(title = 'Highlighted Layer Data Map',
            caption = paste("A U.S. map of where [", datasetName,
                            "] sampling occured,  \nusing layer information since this dataset does not have profile information.", sep = ""))
   } else {}
   
#### LAYER MAP ####
   datasetInfo <- datasetLayer %>%
                    dplyr::select(where(function(xx){!all(is.na(xx))})) %>% 
                    select(`long (dec. deg)`, `lat (dec. deg)`) %>%
                    unique()
   ggRedX <- list(geom_point(data= datasetInfo, 
             aes(x = `long (dec. deg)` + shift360, y = `lat (dec. deg)`),
             shape = 'x', color = 'red', size = 3))
   if(nrow(unique(datasetLayer %>% select(`country (country)`))) > 1 ){
     layerMap <-ggplot(data =  map_data('world')) +   #using world map if sites span multiple countries
       ggMapBase + ggRedX + mapTitleLayer
   } else if(unique(datasetLayer %>% select(`country (country)`)) ==  "United States" &
             unique(datasetLayer %>% select(`state (state_province)`)) != "Alaska"){   #US state mapping
     layerMap <- ggplot(data =  map_data('state') %>% filter(region %in% datasestState)) +
       ggMapBase + ggStateFillBlue + ggRedX + mapTitleLayer
   } else { shift360 <- 360 #datasets in Alaska
     layerMap <- ggplot(data =  map_data('world2')) +
       ggMapBase + ggUSAFillBlue + ggRedX + alaskaCoords + mapTitleLayer
     }
   
#### PLOTS ####   
   if(datasetName %in% ISCN3$profile$dataset_name_sub){
      histograms <-  ggplot(datasetProfile %>%
                         dplyr::select(where(function(xx){!all(is.na(xx))})) %>%
                         pivot_longer(cols =  intersect(names(.), ISCN3$type_columns$num_cols), 
                                      values_to = 'measurement', names_to = 'type')) +
                     geom_histogram(aes(x=measurement)) +
                     facet_wrap(~type, scales='free') +
                     labs(title = 'Profile Value Histograms',
                          caption = paste("Histograms detailing measurements in [", datasetName,
                                          "] profile data.", sep = ""))
   } else {}
     
   depthPlots <- ggplot(datasetLayer %>%
                        dplyr::select(where(function(xx){!all(is.na(xx))})) %>% 
                              pivot_longer(cols=c('layer_top (cm)', 'layer_bot (cm)'),
                                      values_to='depth') %>%
                              pivot_longer(cols = intersect(names(.), ISCN3$type_columns$num_cols), 
                                      values_to = 'measurement', names_to = 'type')) +
                              geom_line(aes(x=depth, y= measurement, group = profile_name), alpha = 0.5) +
                          facet_wrap(~type, scales='free') +
                          labs(title = 'Layer Data Depth Plots',
                               caption = paste("Plots of [", datasetName,
                                        "] layer data by depth.", sep = ""))
   
   theme_set(theme_grey())  #setting ggplot theme

#### RETURN ####   
return(list(summaryStudyTable = studyTable,
            profileTable = profileTable,
            layerTable = layerTable,
            profileMap = profileMap,
            layerMap = layerMap,
            profileValueHistograms = histograms,
            depthValuePlots = depthPlots))  
}

#Visuals <- plotGenerate(ISCN3, datasetName = 'Heckman lithosequence')
#Visuals <- plotGenerate(ISCN3, datasetName = 'Heckman/Swanston Biscuit Burn')
#Visuals <- plotGenerate(ISCN3, datasetName = 'Jorgensen_NPS')
Visuals <- plotGenerate(ISCN3, datasetName = "Worldwide soil carbon and nitrogen data")


print(Visuals)

```

## Heckman/Swanston Biscuit Burn
```{r}
#### dataset variables for debugging ####
# datasetName = 'Lu_PIMA'
# datasetName = 'Jorgensen_NPS'
# datasetName = 'Heckman/Swanston Biscuit Burn'
# datasetName = 'Heckman lithosequence'
datasetName = "Vogel"

datasetStudy <- ISCN3$study %>%
    filter(dataset_name == datasetName)
  datasetProfile <- ISCN3$profile %>%
    filter(dataset_name_sub == datasetName)
  datasetLayer <- ISCN3$layer %>%
    filter(dataset_name_sub == datasetName)

```

## Lu_PIMA
```{r}




```

## Heckman lithosequence
```{r}
```

## Oak Ridge National Lab_Loblolly_DWJ
```{r}
```

## Lu_LTER
```{r}
```

## Lehmann Soil C&BC #1
```{r}
```

## Schuur
```{r}
```

## Lehmann NE US soils
```{r}
```

## Vogel
```{r}
# FIX MOD DATE IN ISCN3_1

```

## Myers-Smith
```{r}
```

## USGS Muhs
```{r}
```

## USGS Harden Yazoo
```{r}
```

## Boby_Mack
```{r}
```

## Kane
```{r}
```

## Bonanza LTER
```{r}
```

## Jorgensen_YKDE
```{r}
```

## UMBS_FASET
```{r}
```

## Jorgensen_ARCN
```{r}
```

## Oak Ridge National Lab_TDE
```{r}
```

## USGS Harden
```{r}
```

## USDA-FS NRS Landscape Carbon Inventory
```{r}
```

## Bockheim
```{r}
```

## Jorgensen_NPS
```{r}
```

## Permafrost_RCN
```{r}
```

## Northern Circumpolar Soil Carbon Database (NCSCD)
```{r}
```

## USGS_S3C
```{r}
```

## AK DSC Project SOC stock computation
```{r}
```

## Worldwide soil carbon and nitrogen data
```{r}
```

## NRCS Sept/2014
```{r}
```


## TODO

- Rename header and file name to reflect dataset_name
- Pull citation and add it to the bib file, check that citation matches the acknowledgement for study
- Check that the clean profile is correct given the orginal information
- Check that the clean layer is correct given the orginal information
- Check that maps are reasonable for the location
- Summerize what the modifications were for this particular dataset
- Create an issue on github with the template below
- Commit and link the commit to the issue

#### Github issue template

- [ ]  Evaluate dataset
- [ ]  Complete template checklist
- [ ]  Write up summary report narrative
- [ ]  Good commenting
- [ ]  External review by someone else

## Citations

Please see [bibtex citation here] for additional details and if you are using ISCN3 please cite.