---
title: "ISCN contributing datasets"
author: "K Todd-Brown (ktoddbrown@ufl.edu)"
site: bookdown::bookdown_site
documentclass: book
output:
  bookdown::gitbook: default
  #bookdown::pdf_book: default
bibliography: data_reports/ISCN3_bibliography.bib
---

<!-- This index.Rmd is the contributing datasets report out -->
<!-- Compile with the below command -->
<!-- bookdown::render_book(input="index.Rmd", output_format = "bookdown::gitbook") -->

# Introduction

ISCNv3 is composed of several contributing data sets. 
This report considers each one separately, and reports on the location of the data, as well as basic summary statistics.

```{r message=FALSE, echo=FALSE}
library(data.table)
#library(SOCDRaH2)
library(tidyverse)
library(lubridate)
library(tibble)
library(ggmap)
library(maps)
library(mapdata)
library(knitr)
library(tidyr)

data_dir <- '~/Documents/Datasets/ISCN' #comment out and set to your directory to run this code

#data_dir <- 'ISCN3' #set to your directory to run this code

knitr::opts_chunk$set(echo = FALSE)
```

```{r eval=FALSE}
ISCN3 <- SOCDRaH2::ISCN3(orginalFormat=TRUE)
citation_raw <- ISCN3$citation
dataset_raw <- ISCN3$dataset
profile_raw <- ISCN3$profile
layer_raw <- ISCN3$layer
rm(ISCN3) #memory management, ISCN3 is big
```

```{r }
#Data currently available via UF lab Dropbox
#Run previous chunk if you need access to the data
#csv data should be archived by summer 2021

citation_raw <- read_delim(file.path(data_dir, 'ISCN3_citation.csv'), delim = ';', col_types = strrep('c', times = 12)) %>% 
  #round all modification dates to their nearest day (ie whole number)
  mutate(`modification_date (YYYY-MM-DD)` = as.character(round(as.numeric(`modification_date (YYYY-MM-DD)`))))

dataset_raw <- read_delim(file.path(data_dir, 'ISCN3_dataset.csv'), delim = ';', col_types = strrep('c', times = 19)) %>% 
  #round all modification dates to their nearest day (ie whole number)
  mutate(`modification_date (YYYY-MM-DD)` = as.character(round(as.numeric(`modification_date (YYYY-MM-DD)`))))

profile_raw <-  vroom::vroom(file.path(data_dir, 'ISCN3_profile.csv'), col_types = strrep('c', times = 44))

layer_raw <- vroom::vroom(file.path(data_dir, 'ISCN3_layer.csv'), col_types = strrep('c', times = 95))
```


## ISCN3 structure

ISCN3 consists of four tables: citation, dataset, profile, and layer. 
The `citation` and `dataset` tables describe information about each data contribution including any citations practices and contact information for the contributor.
The `profile` table describes core summary information that is 2D resolved.
The `layer` table, in contrasts, holds 3D resolved information.

The headers for these tables are:

**`citation`**: `r names(citation_raw)`

**`dataset`**: `r names(dataset_raw)`

**`profile`**: `r names(profile_raw)`

**`layer`**: `r names(layer_raw)`

```{r}
#when were ISCN SOC stocks caluclated?

#temp <- unique(ISCN_raw$layer[,c('dataset_name_sub', 'dataset_name_soc')]) %>%
#  mutate(value = TRUE) %>%
#  pivot_wider(id_cols = 'dataset_name_sub', names_from='dataset_name_soc')

#looking at temp, the dataset_name_soc doesnt' make any sense to us. So we are going dataset by dataset to reconstruct for ISCN4.

#remove NRCS to be replaced with new database
#as.data.frame(ISCN_raw$layer[dataset_name_sub != 'NRCS Sept/2014', -c('ISCN 1-1 (2015-12-10)')]) %>%
knitr::kable(layer_raw %>%
  group_by(dataset_name_sub) %>% 
  tally(name = 'layer_rows') %>%
  full_join( profile_raw %>%
      group_by(dataset_name_sub) %>% 
      tally(name = 'profile_rows'),
    by = 'dataset_name_sub') %>% 
      arrange(layer_rows),
  caption = 'Row counts for each contributing dataset submision (dataname_sub) in the layer and profile data tables.')
```

Column types are set as follows:

```{r}

type_cols <- list(num_cols  = c("lat (dec. deg)", "long (dec. deg)",
                                "layer_top (cm)", "layer_bot (cm)",
                                "oc (percent)", 'c_tot (percent)', 'loi (percent)',
                                'bd_samp (g cm-3)',  'bd_tot (g cm-3)', "bd_whole (g cm-3)", 'bd_other (g cm-3)',
                                'soc (g cm-2)', "soc_depth (cm)",
                                'wpg2 (percent)',
                                'caco3 (percent)',
                                'sand_tot_psa (percent)', 'silt_tot_psa (percent)', 'clay_tot_psa (percent)', 
                                'n_tot (percent)', "c_to_n (mass ratio)",
                                'cat_exch (cmol H+ kg-1)',
                                "al_dith (specified by al_fe_units)", "al_ox (specified by al_fe_units)", "al_other (specified by al_fe_units)",
                                "fe_dith (specified by al_fe_units)", "fe_ox (specified by al_fe_units)", "fe_other (specified by al_fe_units)",
                                "mn_dith (specified by al_fe_units)", "mn_ox (specified by al_fe_units)", "mn_other (specified by al_fe_units)",
                                "ca_ext (specified by bc_units)", "k_ext (specified by bc_units)", "mg_ext (specified by bc_units)", 
                                "na_ext (specified by bc_units)",
                                "ca_al (specified by bc_units)",
                                'ph_h2o', 'ph_cacl', 'ph_other',
                                "p_bray (specified by p_units)", "p_ox (specified by p_units)", "p_meh (specified by p_units)",
                                "p_other (specified by p_units)", 
                                "base_sum (specified by cec_h_units)", "bs (percent)", "bs_sum (percent)",
                                "h_ext (specified by metal_ext_units)", "zn_ext (specified by metal_ext_units)",
                                "cec_sum (specified by cec_h_units)", "ecec (specified by cec_h_units)", "cec_h_units (extract_units)", 
                                "13c (‰)", "14c (‰)", '15n (‰)',
                                "root_weight (g)", "14c_sigma (‰)", "14c_age (BP)", "14c_age_sigma (BP)", 
                                "fraction_modern", "fraction_modern_sigma",
                                "elevation (m)", 
                                "aspect_deg (degree)", "slope (percent)",
                                 "thaw_depth_profile (cm)",
                                'map (mm)', 'mat (°C)'), 
                  factor_cols = c('dataset_name_sub', "datum (datum)", 
                                  "country (country)", "state (state_province)",
                                  "hzn", "hzn_desgn", "hzn_desgn_other",
                                  "site_perm (site_perm)", "runoff (runoff)",
                                  "soil_series", 'color', 'soil_taxon',
                                  'textureClass',
                                  "profile_zero_ref (profile_zero_ref)", 
                                  "ecoregion", "surface_veg",
                                  'vegclass_local', 
                                  "landuse (landsat)", 'landform (landform)', 'landscape (landscape)', 
                                  '2d_position (2d_position)', 
                                  'drainagecl (drainage)',
                                  'root_quant_size',
                                  "al_fe_units (extract_units)", "metal_ext_units (extract_units)", "p_units (extract_units)",
                                  "bc_units (extract_units)", 
                                  "bdNRCS_prep_code", "cNRCS_prep_code", 
                                   'dataset_type (dataset_type)'), 
                  date_cols = c("observation_date (YYYY-MM-DD)", 
                                "modification_date (YYYY-MM-DD)"),
                  char_cols = c("dataset_name", 
                                'site_name', 'profile_name', 'layer_name',
                                "curator_name", "curator_organization",
                                "curator_email",
                                "contact_name", "contact_email",
                                "reference","dataset_description", 
                                "c_method", 'soc_method', "soc_spatial_flag", 'soc_carbon_flag',
                                'bd_method', 'ph_method', 
                                'wpg2_method', "p_method",
                                "al_fe_method", "bc_method", "metal_ext_method", 
                                'site_note', 'landform_note', 'layer_note',
                                "citation", "citation_usage", "acknowledgement", "acknowledgement_usage"),
                  discard_cols = c("total_lcount", "carbon_lcount", "soc_lcount", "soc_lcount_ISCN",
                                   "total_pcount", "soc_pcount", "soc_pcount_ISCN",
                                   'total_scount', "soc_scount", "soc_scount_ISCN"),
                  id_cols = c("dataset_name", 
                              'site_name', 'profile_name', 'layer_name',
                              'dataset_name_sub'),
                  noAction_cols = c("ISCN 1-1 (2015-12-10)", "locator_alias", "locator_parent_alias", "dataset_name_soc"))

missingCols <- setdiff(unique(c(names(citation_raw), names(dataset_raw), names(profile_raw), names(layer_raw))), unlist(type_cols))
if(length(missingCols) > 0){
  cat(paste('Column names unspecified:', paste(missingCols, collapse = '", "')))
}
```

**numerical** [`r length(type_cols$num_cols)`] `r type_cols$num_cols`

**factor** [`r length(type_cols$factor_cols)`] `r type_cols$factor_cols`

**date** [`r length(type_cols$date_cols)`] `r type_cols$date_cols`

**id** [`r length(type_cols$id_cols)`] `r type_cols$id_cols`

**discard** [`r length(type_cols$discard_cols)`] `r type_cols$discard_cols`

**noAction** [`r length(type_cols$noAction_cols)`] `r type_cols$noAction_cols`

```{r}


#' Cast the columns in a standard way
#'
#' This function removes the columns that are entirely NA then goes through and uses a list of columns names to cast these to either numeric factor or date. Note that character columns are left alone since we assume all the columns were characters coming into the function.
#'
#' @param data A data frame with column names that match the columns identified in the list `type_cols`
#' @param column_types A list with four character vectors named `num_cols`, `factor_cols`, `date_cols`, `discard_cols` that correspond to columns in `data` that are cast as numeric, factor, lubridate::Date, or discarded from the data frame.
#' 
#' @return a data frame that matches the `data` argument with the column types modified or dropped if specified as discarded or discarded because they started as being NA columns.
standardCast <- function(data, column_types = type_cols){
  return(data %>%
           select(where(function(xx){!all(is.na(xx))})) %>%
           mutate_at(intersect(c(column_types$num_cols, column_types$date_cols),
                               names(.)), as.numeric) %>%
           mutate_at(intersect(column_types$factor_cols, names(.)), as.factor) %>%
           mutate_at(intersect(column_types$date_cols, names(.)), function(xx){
             ##Both conditions will be run but things throw warnings for the wrong conditional... supressing this function
             suppressWarnings(
               ans <- case_when(is.na(xx) ~ NA_Date_,
                                as.numeric(xx) < 2020 ~ lubridate::ymd(paste0(xx, '-01-01')),
                                as.numeric(xx) >= 2020 ~ lubridate::as_date(as.numeric(xx), 
                                                              origin = lubridate::ymd('1899-12-31')),
                                TRUE ~ NA_Date_)
             )
             return(ans)
           }) %>%
           select(-any_of(column_types$discard_cols)))
}
```
